### 2022-11-12 paper/paper-CableRoute.tex
add first (emtpy) versions of the paper.

\documentclass[acmsmall]{acmart}

\AtBeginDocument{%
	\providecommand\BibTeX{{%
			Bib\TeX}}}

\setcopyright{acmcopyright}
\copyrightyear{2022}
\acmYear{2022}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{111}
\acmMonth{8}

\begin{document}
	\title{Web Processing - Standardised GIS Analyses for Cable Route Planning}

	\author{Sebastian Heiden}
	\email{u38XXX@hs-harz.de}
	\affiliation{%
		\institution{Harz University of Applied Sciences}
		\streetaddress{Friedrichstrasse 57-59}
		\city{Wernigerode}
		\state{Saxony-Anhalt}
		\country{Germany}
		\postcode{38855}
	}

	\renewcommand{\shortauthors}{Heiden}

	\begin{abstract}
		add as last part
	\end{abstract}

	%%
	%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
	%% Please copy and paste the code instead of the example below.
	%%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>10010520.10010553.10010562</concept_id>
		<concept_desc>Computer systems organization~Embedded systems</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>10010520.10010575.10010755</concept_id>
		<concept_desc>Computer systems organization~Redundancy</concept_desc>
		<concept_significance>300</concept_significance>
		</concept>
		<concept>
		<concept_id>10010520.10010553.10010554</concept_id>
		<concept_desc>Computer systems organization~Robotics</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		<concept>
		<concept_id>10003033.10003083.10003095</concept_id>
		<concept_desc>Networks~Network reliability</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}

	\ccsdesc[500]{Computer systems organization~Embedded systems}
	\ccsdesc[300]{Computer systems organization~Redundancy}
	\ccsdesc{Computer systems organization~Robotics}
	\ccsdesc[100]{Networks~Network reliability}

	%%
	%% Keywords. The author(s) should pick words that accurately describe
	%% the work being presented. Separate the keywords with commas.
	\keywords{datasets, neural networks, gaze detection, text tagging}

	\received{20 February 2007}
	\received[revised]{12 March 2009}
	\received[accepted]{5 June 2009}

	%%
	%% This command processes the author and affiliation and title
	%% information and builds the first part of the formatted document.
	\maketitle

	\section{Introduction}

	\section{Template Overview}

	\subsection{Template Styles}

	\subsection{Template Parameters}

	\section{Modifications}

	\section{Typefaces}

	\section{Title Information}

	\section{Authors and Affiliations}

	\section{Rights Information}

	\section{CCS Concepts and User-Defined Keywords}

	\section{Sectioning Commands}

	\section{Tables}

	\section{Math Equations}

	\subsection{Inline (In-text) Equations}

	\subsection{Display Equations}

	\section{Figures}

	\section{Citations and Bibliographies}

	\section{Acknowledgments}

\section{Multi-language papers}

\begin{acks}
	bla
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

\subsection{Part Two}

\section{Online Resources}

\end{document}
\endinput
%%
%% End of file `sample-acmsmall.tex'.

### 2022-11-12 code/rules2weights.py
updated requirements.txt
rules2weights.py:
 - save partial results in a sub folder

@@ -137,7 +137,7 @@ def process_base_rule(df: DataFrame, _resolution: float | tuple[float, list], _c
                       _save_dir: Path, _all_touched: bool) -> DataArray:
     named_tuple = list((df.copy().loc[df['Description'] == 'Base']).itertuples())[0]
     vec = __vector_processing(named_tuple, _crs)
-    save_path = _save_dir / f"rule_{named_tuple.LineNum}_{named_tuple.Description}.tif"
+    save_path = _save_dir / f"rule_{named_tuple.LineNum}_{named_tuple.Description}_res_{resolution}_all_touched_{all_touched}.tif"
     raster = make_geocube(vec, ['Weights', ], resolution=resolution, fill=999,
                           rasterize_function=partial(rasterize_image, all_touched=_all_touched))
     raster = raster['Weights']
@@ -149,11 +149,12 @@ def process_rule(named_tuple, example_da: DataArray, _crs: int,
                  _save_dir: Path, _all_touched: bool, bbox) -> list[str | Path]:
     vec = __vector_processing(named_tuple, _crs, bbox)
 
-    save_path = _save_dir / f"rule_{named_tuple.LineNum}_{named_tuple.Description}.tif"
+    save_path = _save_dir / f"res_{resolution}/all_touched_{all_touched}/rule_{named_tuple.LineNum}_{named_tuple.Description}.tif"
     raster = make_geocube(vec, ['Weights'], like=example_da, fill=999,
                           rasterize_function=partial(rasterize_image, all_touched=_all_touched))
     raster = raster['Weights']
     raster.data[example_da == example_da.rio.nodata] = raster.rio.nodata
+    save_path.parent.mkdir(parents=True, exist_ok=True)
     write_compress(raster, save_path)
     return [save_path, ]
 

### 2022-11-25 paper/paper-CableRoute.tex
update and bug fixes for rules2weights.py:  -  added parallel processing for buffering larger files  - bug fix apply filtering by bbox after reprojection
requirements.txt:
 - add ray for multiprocessing

paper-CableRoute.tex
 - add bibtex
 - changed document class
 - add chapters
 -  add first text for chapter `Methods`

@@
-\documentclass[acmsmall]{acmart}
+% !TeX spellcheck = en_US
+
+\documentclass[acmtog]{acmart}
+\usepackage[acronym]{glossaries}
 
@@
 
+\newacronym{wfs}{wfs}{Web Feature Service}
+
 \begin{document}
@@
 	\author{Sebastian Heiden}
-	\email{u38XXX@hs-harz.de}
+	\email{u38439@hs-harz.de}
 	\affiliation{%
@@
 
-	\section{Template Overview}
-
-	\subsection{Template Styles}
-
-	\subsection{Template Parameters}
-
-	\section{Modifications}
-
-	\section{Typefaces}
-
-	\section{Title Information}
-
-	\section{Authors and Affiliations}
-
-	\section{Rights Information}
-
-	\section{CCS Concepts and User-Defined Keywords}
-
-	\section{Sectioning Commands}
-
-	\section{Tables}
-
-	\section{Math Equations}
-
-	\subsection{Inline (In-text) Equations}
-
-	\subsection{Display Equations}
-
-	\section{Figures}
-
-	\section{Citations and Bibliographies}
-
-	\section{Acknowledgments}
-
-\section{Multi-language papers}
+	\section{Methods}
+	We retrieve a set of different spacial data-sets from  public sources as a basis to create the cost raster.
+Field of study are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
+	Areas protected by different European and National conservation laws are provided by the German Environment Agency as \acrfull{wfs}~\cite{noauthor_schutzgebiete_nodate}.
+	\section{Results}
+	\section{Discussion}
+	\section{Related Works}
+	\section{Conclusion}
 
@@
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-base}
+\bibliography{Web_Processing.bib}
 

### 2022-11-25 code/rules2weights.py

@@ -1,5 +1,6 @@
 from argparse import ArgumentParser
 from functools import partial
+from os import cpu_count
 from pathlib import Path
 from pprint import pprint
 
@@ -9,8 +10,9 @@ from brackettree import Node
 from brackettree.nodes import TextNode, RoundNode
 from geocube.api.core import make_geocube
 from geocube.rasterize import rasterize_image
-from geopandas import read_file, GeoDataFrame
+from geopandas import read_file, GeoDataFrame, GeoSeries
 from pandas import read_excel, DataFrame, concat, merge, Series
+from ray.util.multiprocessing import Pool
 from rioxarray import open_rasterio
 from rioxarray.merge import merge_arrays
 from tqdm import tqdm
@@ -110,11 +112,24 @@ def buffer(_gdf: GeoDataFrame, distance: float | str) -> GeoDataFrame:
 
     if isinstance(distance, str):
         distance = __estimate_buffer_size(_gdf, Series(np.zeros(len(_gdf)), index=_gdf.index), Node(distance))
+        distance = distance.values
     else:
-        distance = distance if not np.isnan(distance) else 0
+        distance = np.full(len(_gdf), distance if not np.isnan(distance) else 0,)
 
-    # TODO ray multi processing
-    _gdf['geometry'] = _gdf['geometry'].buffer(distance)
+    if len(_gdf) > 5_000:  # ray multiprocessing the buffer
+        geometry_arrays = np.array_split(_gdf['geometry'].values, cpu_count() - 1)
+        distances = np.array_split(distance, cpu_count() - 1)
+
+        def single_buffer(geometry_array, distance_array):
+            s = GeoSeries(geometry_array)
+            return s.buffer(distance_array)
+        with Pool(cpu_count()-1) as p:
+            geometries = p.starmap(single_buffer, zip(geometry_arrays, distances))
+        p.close()
+        _gdf['geometry'] = concat(geometries).values
+        p.terminate()
+    else:
+        _gdf['geometry'] = _gdf['geometry'].buffer(distance)
     return _gdf
 
 
@@ -123,7 +138,7 @@ def __estimate_buffer_size(data: GeoDataFrame, _buffer: Series, _rule: Node, ) -
     estimate the size of the Buffer for every GeoObject in the GeoDataFrame
     :param data: GeoDataFrame
     :param _buffer: Series with the current buffer distances
-    :param _rule: rule to estimate the size of the buffer as Node
+    :param _rule: rule to estimate the size of the buffer as a Node
     :return: Series
     """
[…]

### 2022-12-11 paper/paper-CableRoute.tex
add requirements
added a sub package for implementing wps and test it with ows.

Paper:
    Beginning to add some content from the midterm presentation to the paper. Layers on their sources and the usage/implementation of the layer -> cost. Added todo notes.

@@
 \usepackage[acronym]{glossaries}
+\usepackage{todonotes}
 
@@
 \newacronym{wfs}{wfs}{Web Feature Service}
+\newacronym{wps}{wps}{Web Processing Service}
 
@@
 
+	Sometimes, finding the shortest path is not sufficient. Additional parameters play also have to be taken into consideration. As the steepness of a road or the soil,
+	play an important role for the building cost of a road or pipeline \todo{source}.
+For planing the additional routes for a power grid, additional aspects as legal regulations and acceptance by the local population have to be taken in consideration.
+Also technical aspects, as the effects on the grid stability might be further points to consider.\cite{schafer_understanding_2022}
+	\todo{what is the least cost path}
+
 	\section{Methods}
@@
 Field of study are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
-	Areas protected by different European and National conservation laws are provided by the German Environment Agency as \acrfull{wfs}~\cite{noauthor_schutzgebiete_nodate}.
+	Areas protected by different European and National conservation laws are provided by the German Environment Agency as \acrfull{wfs}~\cite{noauthor_schutzgebiete_2015}.
+The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography and Geodesy\cite{noauthor_digitales_2021}.
+The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap.\cite{boeing_osmnx_2017}
+	Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of Lower Saxony\cite{noauthor_opengeodatani_2022}.
+In addition local planning geodata for the land usage are taken from
+	from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)\cite{noauthor_metropolplaner_2022}
+
+	PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
+	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin 'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
+	\todo{layers -> aggregation -> costs}
+
+	The files form above are optionally filtered and buffered and than converted into rasters. Filtering the layers of the files for special attributes \ldots
+	\todo{add some text} Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the correct physical width, or to add a minimum distance from an existing of planed area to the new power grid.
+
+	The costs has been grouped into different levels~(see table \ref{table:1}) starting from preferential areas with very low costs, via no restrictions, which is the default, used when no other layers are covering the local area, to restricted, strongly restricted and prohibited areas with high costs.
+These higher costs resemble the degree how much a local area should be avoided, while routing the path.
+The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels the algorithm is willing to go, for avoiding an area of high costs.
+	Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum, the weight that resembles the costs for these types of areas, has to be especially high.
+
+	\begin{table}[h!]
+		\caption{Used levels of costs, the applied numerical equivalent and and example layer this cost have been used for.}
+		\label{table:1}
+		\centering
+		\begin{tabular}{ l  r  l }
+			Cost Level & Cost &  Example\\
+			\hline
+			Prohibited & 500					& National Parks, Buildings \\
+			strongly Restricted & 10 	& Bird Reserve \\
+			Restricted & 5	& Industrial Areas \\
+			No Restriction & 0.5					& Default\\
+			Preferential & 0.1					& Power Grid, Motorway Buffers\\
+		\end{tabular}
+
+	\end{table}
+
 	\section{Results}
+
 	\section{Discussion}

### 2022-12-11 code/wps/ows_test.py

@@ -1,0 +1,19 @@
+from owslib.wps import WebProcessingService, GMLMultiPolygonFeatureCollection, printInputOutput
+
+client = WebProcessingService('http://127.0.0.1:5000/wps')
+client.getcapabilities()
+print(client.processes)
+
+process_id = 'buffer'
+
+buffer_process_description = client.describeprocess(process_id)
+for i in buffer_process_description.dataInputs:
+    printInputOutput(i)
+
+polygon = [(-102.8184, 39.5273), (-102.8184, 37.418), (-101.2363, 37.418), (-101.2363, 39.5273), (-102.8184, 39.5273)]
+featureCollection = GMLMultiPolygonFeatureCollection([polygon])
+
+inputs = [("poly_in", featureCollection),
+          ("buffer", 10.0), ]
+
+execution = client.execute(process_id, inputs, output="OUTPUT")

### 2022-12-11 code/wps/pywps_test.py

@@ -1,0 +1,11 @@
+import flask
+import pywps
+
+from .buffer import Buffer
+from .centroids import Centroids
+
+
+service = pywps.Service([Buffer(), Centroids()])
+app = flask.Flask(__name__)
+app.route('/wps', methods=['GET', 'POST'])(lambda: service)
+app.run()

### 2022-12-17 paper/paper-CableRoute.tex
add requirements
renamed sub folder code to src

Paper:
    minor additions

add line2line_distance.ipynb
 - notebook to estimate costs length, etc. for the least cost path

add least_cost_path_stat.ipynb
 - stat of the res vs execution time

@@
 	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin 'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
-	\todo{layers -> aggregation -> costs}
 
-	The files form above are optionally filtered and buffered and than converted into rasters. Filtering the layers of the files for special attributes \ldots
-	\todo{add some text} Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the correct physical width, or to add a minimum distance from an existing of planed area to the new power grid.
-
+	The different layers from the different entities are optionally filtered, buffered and than converted into rasters.
+Filtering the layers of the files for special attributes enables to further differentiate further.
+For examples makes it possible to differentiate between heath and uncultivated land.
+	 Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
+ 	Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
+ 	The costs of all layers of all layers is aggregated with the maximum function.
+Thus, an area covered by multiple layers is uniformly used with the highest costs.
+For every area in the study area.
+that is not covered by any layer, is given the default cost.
 	The costs has been grouped into different levels~(see table \ref{table:1}) starting from preferential areas with very low costs, via no restrictions, which is the default, used when no other layers are covering the local area, to restricted, strongly restricted and prohibited areas with high costs.
@@
 	Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum, the weight that resembles the costs for these types of areas, has to be especially high.
+	\todo{all touched}
 
 	\begin{table}[h!]
-		\caption{Used levels of costs, the applied numerical equivalent and and example layer this cost have been used for.}
+		\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
 		\label{table:1}
@@
 		\end{tabular}
-
 	\end{table}
+	The completed list of layers and the processing applied to them, can be found in Supplement S1. \todo{create the Supplement from the processing rules}
 
 	\section{Results}
+	In this chapter we want to show the different cost rasters, that were created from the very same set of layers, but computed for different resolutions.
+From this different rasters the cost paths are calculated and compared.
+In the last step the rasters with lower resolution were used to calculate in a way, that they shall result in similar paths as if the paths were computed from a high resolution raster.
+	\subsection{Cost Rasters}
+	\subsection{Least Cost Paths}
+	\subsection{Faster Processing of the Cost Path Algorithm}
 
 	\section{Discussion}
+	\todo{discuss aggrgation with max vs sum}
 	\section{Related Works}
@@
 \begin{acks}
-	bla
+	\ldots
 \end{acks}

### 2022-12-17 code/lod1_load/download_lod1.py

@@ -1,100 +1,0 @@
-from argparse import ArgumentParser
-from concurrent.futures import ThreadPoolExecutor
-from functools import partial
-from io import BytesIO
-from json import load
-from pathlib import Path
-from zipfile import ZipFile
-
-import numpy as np
-from fiona.errors import UnsupportedGeometryTypeError
-from geopandas import read_file, GeoDataFrame
-from pandas import concat, date_range
-from requests import get, codes
-from requests.exceptions import ReadTimeout, ConnectionError
-from shapely.geometry import Point
-from tqdm import tqdm
-from warnings import warn
-
-
-def download_zips(_url: str, _save_path: Path) -> bool:
-    try:
-        response = get(_url, timeout=(2, 5))
-    except (ConnectionError, ReadTimeout):
-        return False
-    if isinstance(response.status_code, int) and response.status_code == codes.ok:
-        with ZipFile(BytesIO(response.content)) as source_zip:
-            source_zip.extractall(_save_path)
-        return True
-    else:
-        return False
-
-
-def concat_vector_files(in_dir: Path) -> GeoDataFrame:
-    lod1_vectors = []
-    shp_files = [str(x) for x in in_dir.iterdir() if x.suffix == '.shp']
-    error_counter = 0
-    for f in shp_files:
-        try:
-            lod1_vectors.append(read_file(f))
-        except UnsupportedGeometryTypeError as e:
-            error_counter = error_counter + 1
-    if error_counter > 0:
-        warn(f'{error_counter} could not be concatenated.')
-
-    return concat(lod1_vectors)
-
-
-if __name__ == '__main__':
-    parser = ArgumentParser(description='')
-    parser.add_argument('json', type=str, help='Full path to the json config file.')
-
-    args = parser.parse_args()
-
-    with open(args.json) as j:
-        config = load(j)
-
-    geb_path = config['areas_vector_path']
-
-    geb_vector = read_file(geb_path)
-    geb_vector = geb_vector.loc[geb_vector.BEZ_LAN == 'Niedersachsen']
-    geb_vector = geb_vector.loc[~geb_vector.BEZ_KRS.isna()]
-    geb_vector = geb_vector.dissolve(by='BEZ_LAN')
-
-    x_range = list(range(2342, 2672, 2))
-    y_range = range(5684, 5970, 2)
-
-    unzip_path = Path(config['unzip_lod1_path'])
-    dates = date_range(*config['dates'], freq='1D', inclusive='both')
-
[…]

### 2022-12-17 code/lod1_load/mass_download_lod1.py

@@ -1,38 +1,0 @@
-from argparse import ArgumentParser
-from concurrent.futures import ThreadPoolExecutor
-from functools import partial
-from io import BytesIO
-from json import load
-from pathlib import Path
-
-from geopandas import read_file
-from requests import get
-
-from download_lod1 import download_zips, concat_vector_files
-
-if __name__ == '__main__':
-    parser = ArgumentParser(description='Download all files ')
-    parser.add_argument('json', type=str, help='Full path to the json config file.')
-
-    args = parser.parse_args()
-
-    with open(args.json) as j:
-        config = load(j)
-
-        geb_path = config['areas_vector_path']
-        response = get(config['mass_geojson_url']).content
-    tmp = BytesIO(response)
-    all_files_gdf = read_file(tmp)
-    unzip_path = Path(config['unzip_lod1_path'])
-
-    p_download = partial(download_zips, _save_path=unzip_path)
-
-    existing_files = [x for x in unzip_path.iterdir() if x.suffix == ".shp"]
-    existing_tiles_names = [x.stem.split('_')[1] for x in existing_files]
-
-    to_download_gdf = all_files_gdf.loc[~all_files_gdf['tile_id'].isin(existing_tiles_names)]
-    with ThreadPoolExecutor() as p:
-        is_downloaded = p.map(p_download, to_download_gdf['3DShape'])
-
-    concat_vector_files(unzip_path).to_file(config['save_result_path'])
-

### 2022-12-17 code/wps/ows_test.py

@@ -1,19 +1,0 @@
-from owslib.wps import WebProcessingService, GMLMultiPolygonFeatureCollection, printInputOutput
-
-client = WebProcessingService('http://127.0.0.1:5000/wps')
-client.getcapabilities()
-print(client.processes)
-
-process_id = 'buffer'
-
-buffer_process_description = client.describeprocess(process_id)
-for i in buffer_process_description.dataInputs:
-    printInputOutput(i)
-
-polygon = [(-102.8184, 39.5273), (-102.8184, 37.418), (-101.2363, 37.418), (-101.2363, 39.5273), (-102.8184, 39.5273)]
-featureCollection = GMLMultiPolygonFeatureCollection([polygon])
-
-inputs = [("poly_in", featureCollection),
-          ("buffer", 10.0), ]
-
-execution = client.execute(process_id, inputs, output="OUTPUT")

### 2022-12-17 code/wps/pywps_test.py

@@ -1,11 +1,0 @@
-import flask
-import pywps
-
-from .buffer import Buffer
-from .centroids import Centroids
-
-
-service = pywps.Service([Buffer(), Centroids()])
-app = flask.Flask(__name__)
-app.route('/wps', methods=['GET', 'POST'])(lambda: service)
-app.run()

### 2022-12-18 src/least_cost_path/dijkstra.py
added some code that was removed by renaming the code folder to src
Testing least cost path with wps (something does not work, sending files?)

@@ -1,0 +1,124 @@
+from math import sqrt
+import queue
+import collections
+
+from shapely.geometry import Point
+
+
+def dijkstra(start_tuple: tuple[tuple[int, int], Point, int], end_tuples: tuple[tuple[tuple[int, int], Point, int]],
+             block: list[list[None | float]], find_nearest: bool):
+    class Grid:
+        def __init__(self, matrix: list[list[None | float]]):
+            self.map = matrix
+            self.h = len(matrix)
+            self.w = len(matrix[0])
+            self.manhattan_boundary = None
+            self.curr_boundary = None
+
+        def _in_bounds(self, _id: tuple[int, int]):
+            x, y = _id
+            return 0 <= x < self.h and 0 <= y < self.w
+
+        def _passable(self, _id: tuple[int, int]):
+            x, y = _id
+            return self.map[x][y] is not None
+
+        def is_valid(self, _id: tuple[int, int]):
+            return self._in_bounds(_id) and self._passable(_id)
+
+        def neighbors(self, _id: tuple[int, int]):
+            x, y = _id
+            results = [(x + 1, y), (x, y - 1), (x - 1, y), (x, y + 1),
+                       (x + 1, y - 1), (x + 1, y + 1), (x - 1, y - 1), (x - 1, y + 1)]
+            results = list(filter(self.is_valid, results))
+            return results
+
+        @staticmethod
+        def manhattan_distance(id1: tuple[int, int], id2: tuple[int, int]):
+            x1, y1 = id1
+            x2, y2 = id2
+            return abs(x1 - x2) + abs(y1 - y2)
+
+        @staticmethod
+        def min_manhattan(curr_node, end_nodes):
+            return min(map(lambda node: Grid.manhattan_distance(curr_node, node), end_nodes))
+
+        @staticmethod
+        def max_manhattan(curr_node, end_nodes):
+            return max(map(lambda node: Grid.manhattan_distance(curr_node, node), end_nodes))
+
+        @staticmethod
+        def all_manhattan(curr_node, end_nodes):
+            return {end_node: Grid.manhattan_distance(curr_node, end_node) for end_node in end_nodes}
+
+        def simple_cost(self, current: tuple[int, int], _next: tuple[int, int]):
+            cx, cy = current
+            nx, ny = _next
+            current_value = self.map[cx][cy]
+            offset_value = self.map[nx][ny]
+            if cx == nx or cy == ny:
+                return (current_value + offset_value) / 2
[…]

### 2022-12-18 src/least_cost_path/find_least_cost_path.py

@@ -1,0 +1,138 @@
+from math import floor
+
+from geopandas import read_file, GeoSeries, GeoDataFrame
+from pandas import concat
+from rioxarray import open_rasterio
+from shapely.geometry import Point, LineString, MultiPoint, box
+from xarray import DataArray
+
+from src.least_cost_path.dijkstra import dijkstra
+from argparse import ArgumentParser
+
+
+# TODO comment for function descriptions
+def _point_to_row_col(point_xy: Point, raster_layer: DataArray):
+    affine = raster_layer.rio.transform()
+    x_res = affine.a
+    y_res = abs(affine.e)
+    x_min, _, _, y_max = raster_layer.rio.bounds()
+
+    col = floor((point_xy.x - x_min) / x_res)
+    row = floor((y_max - point_xy.y) / y_res)
+
+    return row, col
+
+
+def features_to_tuples(point_features: GeoSeries, raster_layer: DataArray) -> list[tuple]:
+    # TODO check data type
+    row_cols = []
+
+    extent = box(*raster_layer.rio.bounds())
+
+    for idx, point_feature in enumerate(point_features):
+        if point_feature.is_valid:
+
+            if isinstance(point_feature, MultiPoint):
+                for point_xy in point_feature:
+                    if extent.contains(point_xy):
+                        row_col = _point_to_row_col(point_xy, raster_layer)
+                        row_cols.append((row_col, point_xy, idx))
+
+            elif isinstance(point_feature, Point):
+                if extent.contains(point_feature):
+                    row_col = _point_to_row_col(point_feature, raster_layer)
+                    row_cols.append((row_col, point_feature, idx))
+
+    return row_cols
+
+
+def _row_col_to_point(row_col, raster_layer: DataArray):
+    affine = raster_layer.rio.transform()
+    x_res = affine.a
+    y_res = abs(affine.e)
+    x_min, y_min, _, y_max = raster_layer.rio.bounds()
+
+    x = (row_col[1] + 0.5) * x_res + x_min
+    y = y_max - (row_col[0] + 0.5) * y_res
+    return Point(x, y)
+
+
+def raster2matrix(block: DataArray) -> tuple[list[list[None | float]], bool]:
+    _contains_negative = False
+    _matrix = [[None if block.data[i, j] == block.rio.nodata else block.data[i, j] for j in range(block.rio.width)]
+               for i in range(block.rio.height)]
+
+    for line in _matrix:
+        for value in line:
+            if value is not None:
[…]

### 2022-12-18 src/lod1_load/download_lod1.py

@@ -1,0 +1,100 @@
+from argparse import ArgumentParser
+from concurrent.futures import ThreadPoolExecutor
+from functools import partial
+from io import BytesIO
+from json import load
+from pathlib import Path
+from zipfile import ZipFile
+
+import numpy as np
+from fiona.errors import UnsupportedGeometryTypeError
+from geopandas import read_file, GeoDataFrame
+from pandas import concat, date_range
+from requests import get, codes
+from requests.exceptions import ReadTimeout, ConnectionError
+from shapely.geometry import Point
+from tqdm import tqdm
+from warnings import warn
+
+
+def download_zips(_url: str, _save_path: Path) -> bool:
+    try:
+        response = get(_url, timeout=(2, 5))
+    except (ConnectionError, ReadTimeout):
+        return False
+    if isinstance(response.status_code, int) and response.status_code == codes.ok:
+        with ZipFile(BytesIO(response.content)) as source_zip:
+            source_zip.extractall(_save_path)
+        return True
+    else:
+        return False
+
+
+def concat_vector_files(in_dir: Path) -> GeoDataFrame:
+    lod1_vectors = []
+    shp_files = [str(x) for x in in_dir.iterdir() if x.suffix == '.shp']
+    error_counter = 0
+    for f in shp_files:
+        try:
+            lod1_vectors.append(read_file(f))
+        except UnsupportedGeometryTypeError as e:
+            error_counter = error_counter + 1
+    if error_counter > 0:
+        warn(f'{error_counter} could not be concatenated.')
+
+    return concat(lod1_vectors)
+
+
+if __name__ == '__main__':
+    parser = ArgumentParser(description='')
+    parser.add_argument('json', type=str, help='Full path to the json config file.')
+
+    args = parser.parse_args()
+
+    with open(args.json) as j:
+        config = load(j)
+
+    geb_path = config['areas_vector_path']
+
+    geb_vector = read_file(geb_path)
+    geb_vector = geb_vector.loc[geb_vector.BEZ_LAN == 'Niedersachsen']
+    geb_vector = geb_vector.loc[~geb_vector.BEZ_KRS.isna()]
+    geb_vector = geb_vector.dissolve(by='BEZ_LAN')
+
+    x_range = list(range(2342, 2672, 2))
+    y_range = range(5684, 5970, 2)
+
+    unzip_path = Path(config['unzip_lod1_path'])
+    dates = date_range(*config['dates'], freq='1D', inclusive='both')
+
[…]

### 2022-12-18 src/lod1_load/mass_download_lod1.py

@@ -1,0 +1,38 @@
+from argparse import ArgumentParser
+from concurrent.futures import ThreadPoolExecutor
+from functools import partial
+from io import BytesIO
+from json import load
+from pathlib import Path
+
+from geopandas import read_file
+from requests import get
+
+from download_lod1 import download_zips, concat_vector_files
+
+if __name__ == '__main__':
+    parser = ArgumentParser(description='Download all files ')
+    parser.add_argument('json', type=str, help='Full path to the json config file.')
+
+    args = parser.parse_args()
+
+    with open(args.json) as j:
+        config = load(j)
+
+        geb_path = config['areas_vector_path']
+        response = get(config['mass_geojson_url']).content
+    tmp = BytesIO(response)
+    all_files_gdf = read_file(tmp)
+    unzip_path = Path(config['unzip_lod1_path'])
+
+    p_download = partial(download_zips, _save_path=unzip_path)
+
+    existing_files = [x for x in unzip_path.iterdir() if x.suffix == ".shp"]
+    existing_tiles_names = [x.stem.split('_')[1] for x in existing_files]
+
+    to_download_gdf = all_files_gdf.loc[~all_files_gdf['tile_id'].isin(existing_tiles_names)]
+    with ThreadPoolExecutor() as p:
+        is_downloaded = p.map(p_download, to_download_gdf['3DShape'])
+
+    concat_vector_files(unzip_path).to_file(config['save_result_path'])
+

### 2022-12-18 src/wps/birdy_test.py

@@ -1,0 +1,18 @@
+from birdy import WPSClient
+from pathlib import Path
+
+pywps = WPSClient('http://localhost:5000/wps')
+line_path = Path(r'.\..\..\results\least_cost_paths\least_cost_path_test_points_res_10_al_false.gml')
+print(line_path.exists())
+line = open(line_path)
+
+cost_raster = Path(r".\..\..\results\weights\result_res_100_all_touched_True.tif")
+start_features = Path(r"..\..\results\test_points\start_point.gpkg")
+end_features = Path(r"..\..\results\test_points\end_point.gpkg")
+
+pywps.lcp(costs=cost_raster,
+          start=start_features,
+          end=end_features).get(asobj=True)[0]
+
+# pywps.buffer(poly_in=line,
+#              buffer=10.0).get(asobj=True)[0]

### 2022-12-18 src/wps/ows_test.py

@@ -1,0 +1,21 @@
+from owslib.wps import WebProcessingService, GMLMultiPolygonFeatureCollection, printInputOutput, ComplexDataInput
+
+client = WebProcessingService('http://127.0.0.1:5000/wps')
+client.getcapabilities()
+print(client.processes)
+
+process_id = 'MyBuffer'
+
+buffer_process_description = client.describeprocess(process_id)
+for i in buffer_process_description.dataInputs:
+    printInputOutput(i)
+
+# polygon = [(-102.8184, 39.5273), (-102.8184, 37.418), (-101.2363, 37.418), (-101.2363, 39.5273), (-102.8184, 39.5273)]
+in_data = ComplexDataInput(r"E:\Basti\Studium\HSHarz\WissenschaftlichesArbeiten\WebProcessing-StandardisedGISAnalysesforCableRoutePlanning\results\least_cost_paths\least_cost_path_test_points_res_10_al_false.gml")
+# featureCollection = GMLMultiPolygonFeatureCollection([polygon, ])
+
+inputs = [("poly_in", in_data),
+          # ("distance", 10.0),
+          ]
+
+execution = client.execute(process_id, inputs, output="OUTPUT")

### 2022-12-18 src/wps/processes/least_cost_path.py

@@ -1,0 +1,39 @@
+from pywps import Process, ComplexInput, ComplexOutput, Format, FORMATS
+
+from src.least_cost_path.find_least_cost_path import find_least_cost_path
+
+
+class LeastCostPath(Process):
+    def __init__(self):
+        inputs = [ComplexInput('costs', 'Cost Raster', supported_formats=[Format('image/tiff'), ]),
+                  ComplexInput('start', 'Starting Point',
+                               supported_formats=[Format('application/gpkg'), Format('application/json'), ]),
+                  ComplexInput('end', 'Ending Point',
+                               supported_formats=[Format('application/gpkg'), Format('application/json'), ])]
+        outputs = [ComplexOutput('out', 'Referenced Output',
+                                 supported_formats=[
+                                     Format('application/json')
+                                 ])]
+
+        super(LeastCostPath, self).__init__(
+            self._handler,
+            identifier='lcp',
+            title='Process least cost path',
+            abstract='Returns a GeoJSON \
+                with with least cost path from cost raster.',
+            inputs=inputs,
+            outputs=outputs,
+            store_supported=True,
+            status_supported=True
+        )
+
+    def _handler(self, request, response):
+        input_cost_raster = request.inputs['layer'][0].file
+        input_start = request.inputs['startingPoint'][0].file
+        input_end_points = request.inputs['endingPoint'][0].file
+
+        lcp = find_least_cost_path(input_cost_raster, 0, False, input_start, input_end_points)
+
+        response.outputs['out'].output_format = Format(FORMATS['JSON'])
+        response.outputs['out'].data = lcp.to_json(indent=2)
+        return response

### 2022-12-18 paper/paper-CableRoute.tex
Paper: adding the values of the mean minium distances and the path lengths ratio of the zonal stats
line2line_distance.ipynb
 - calculate the zonal stat

@@
 	\subsection{Least Cost Paths}
+
+	\begin{table}[h!]
+	\caption{Least cost paths for the different resolutions of the rasters.}
+	\label{table:2}
+	\centering
+	\begin{tabular}{ r  r  r r }
+		resolution /m & $length_{all touched false} /m$ & $length_{all touched true} /m$ & mean minimum distance /m\\
+		\hline
+		5 & 76136.27					& 78002.00 & 126.04 \\
+		10 & 75430.10 	& 77936.57 & 277.92 \\
+		50 & 76135.02	& 70619,95 & 1140.01 \\
+		100 & 76283.80	& 74120.73				& 1946.41 \\
+	\end{tabular}
+	\end{table}
+	\subsection{Faster Processing of the Cost Path Algorithm}
+
+	The paths between the all\_touched False runs is less distinct. The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m. \todo{add the overall cost}
+
+	\begin{table}[h!]
+		\caption{Ratio Category values of each least cost path.}
+		\label{table:3}
+		\centering
+		\begin{tabular}{ r  r  r  r  r  r  r }
+			resolution /m & all touched & $ r_{Preferential} \% $  & $ r_{No Restriction} \% $ & $ r_{Restricted} \% $ & $ r_{strongly Restricted}\% $ & $ r_{Prohibited} \% $ \\
+			\hline
+			5 & False &  4.29 & 59.01  & 8.88 & 0.73 & 27.08  \\
+			5 & True &  17.38 & 67.25  & 1.17 & 0.97 & 13.23\\
+
+			10 & False &  17.90 & 68.73  & 0.92 & 0.80 & 11.64\\
+			10 & True &  17.37 & 66.42  & 1.47 & 1.38 & 13.36\\
+
+			50 & False &  18.73 & 67.87  & 0.87 & 0.68 & 11.85\\
+			50 & True &  8.58 & 74.95  & 3.95 & 4.08 & 8.45\\
+
+			100 & False &  18.48 & 69.82  & 1.31 & 0.90 & 9.52\\
+			100 & True &  6.16 & 71.74  & 5.75 & 9.69 & 7.04\\
+
+		\end{tabular}
+	\end{table}
+	\todo{Check why values are so different for 5 m all touched False.}
+	\todo{Discuss WHY}
 	\subsection{Faster Processing of the Cost Path Algorithm}
@@
 	\section{Discussion}
-	\todo{discuss aggrgation with max vs sum}
+	\todo{discuss aggregation with max vs sum}
 	\section{Related Works}

### 2022-12-18 src/wps/pywps_test.py
add pywps_test.py  after renaming folder code to src. Try the LeastCostPath Process
 added line2line_distance.ipynb mentioned in last commit

@@ -1,0 +1,12 @@
+import flask
+import pywps
+
+from .processes.buffer import Buffer, MyBuffer
+from .processes.centroids import Centroids
+from .processes.least_cost_path import LeastCostPath
+
+
+service = pywps.Service([Buffer(), MyBuffer(), Centroids(), LeastCostPath()])
+app = flask.Flask(__name__)
+app.route('/wps', methods=['GET', 'POST'])(lambda: service)
+app.run()

### 2023-01-08 paper/paper-CableRoute.tex
updated paper add description for rasterization, including schematic.
added parts for the result tables (for the normal least cost path estimation via qgis plugin), and describes that tables as text.

--- 2022-12-17 paper/paper-CableRoute.tex
@@
 \documentclass[acmtog]{acmart}
+\usepackage{amsmath}
+\usepackage{multirow}
 \usepackage[acronym]{glossaries}
 \usepackage{todonotes}
+\usepackage{tikz}
 
@@
 \acmNumber{4}
-\acmArticle{111}
-\acmMonth{8}
+\acmArticle{28}
+\acmMonth{2}
 
@@
 \newacronym{wps}{wps}{Web Processing Service}
+\newacronym{res}{res}{resolution}
+\newacronym{r}{r}{ratio}
+\newacronym{al}{al}{all touched}
+\newacronym{t}{t}{True}
+\newacronym{f}{f}{False}
+\newacronym{mmd}{mmd}{mean minimum distance}
+\newacronym{agg}{agg}{aggregated}
+\newacronym{rel}{rel}{relative}
+\newacronym{corr}{corr}{corrected}
 
@@
 	\begin{abstract}
-		add as last part
+		add as final part
 	\end{abstract}
@@
 
-	Sometimes, finding the shortest path is not sufficient. Additional parameters play also have to be taken into consideration. As the steepness of a road or the soil,
-	play an important role for the building cost of a road or pipeline \todo{source}.
-For planing the additional routes for a power grid, additional aspects as legal regulations and acceptance by the local population have to be taken in consideration.
+	Sometimes, finding the shortest path is not sufficient.
+	Additional parameters play also have to be taken into consideration.
+	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline
+	\todo{source}.
+	For planing the additional routes for a power grid, additional aspects as legal regulations and acceptance
+	by the local population have to be taken in consideration.
 	Also technical aspects, as the effects on the grid stability might be further points to consider.\cite{schafer_understanding_2022}
@@
 	Field of study are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
-	Areas protected by different European and National conservation laws are provided by the German Environment Agency as \acrfull{wfs}~\cite{noauthor_schutzgebiete_2015}.
-The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography and Geodesy\cite{noauthor_digitales_2021}.
+	Areas protected by different European and National conservation laws are provided by the German Environment Agency
+	as \acrfull{wfs}~\cite{noauthor_schutzgebiete_2015}.
+
+	The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
+	and Geodesy\cite{noauthor_digitales_2021}.
 	The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap.\cite{boeing_osmnx_2017}
-	Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of Lower Saxony\cite{noauthor_opengeodatani_2022}.
+	Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
+	Lower Saxony\cite{noauthor_opengeodatani_2022}.
 	In addition local planning geodata for the land usage are taken from
@@
 	PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
-	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin 'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
+	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin
+	'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
 
-	The different layers from the different entities are optionally filtered, buffered and than converted into rasters.
+	The different layers from the different entities are optionally filtered, buffered and than rasterized.
 	Filtering the layers of the files for special attributes enables to further differentiate further.
 	For examples makes it possible to differentiate between heath and uncultivated land.
-	 Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
+	Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
+	correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
  	Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
- 	The costs of all layers of all layers is aggregated with the maximum function.
+ 	The costs of all layers of is aggregated with the maximum function.
 	Thus, an area covered by multiple layers is uniformly used with the highest costs.
-For every area in the study area.
-that is not covered by any layer, is given the default cost.
-	The costs has been grouped into different levels~(see table \ref{table:1}) starting from preferential areas with very low costs, via no restrictions, which is the default, used when no other layers are covering the local area, to restricted, strongly restricted and prohibited areas with high costs.
+	For every area in the study area, that is not covered by any layer, is given the default cost.
+	The costs has been grouped into different levels~(see table~\ref{table:1}) starting from preferential areas with
+	very low costs, via no restrictions, which is the default, used when no other layers are covering the local area,
+	to restricted, strongly restricted and prohibited areas with high costs.
 	These higher costs resemble the degree how much a local area should be avoided, while routing the path.
-The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels the algorithm is willing to go, for avoiding an area of high costs.
-	Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum, the weight that resembles the costs for these types of areas, has to be especially high.
-	\todo{all touched}
+	The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
+	the algorithm is willing to go, for avoiding an area of high costs.
+	Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum,
+	the weight that resembles the costs for these types of areas, has to be especially high. \\
+	The rasterziation transforms a vector into a raster.
+	The rasterization can be executed in to different ways.
+	In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
+	given resolution and the new affine transformation and the coordinate reference system of the vector.
+	Both rasterization techniques differ  in who the pixels are selected, that describe, that a geo-object overlays the pixel.
+	The pixels can either be described as overlayed with a geo-object, when either the center of that pixel is overlayed
+	by the geo-object, or any part of the pixel is overlayed.
+	Hence, the version with any part overlayed is called all touched True.
+	The version where a overlayed pixel center is require is thus called all touched False.
+	All touched False is considered the default (see figure~\ref{fig:alltouched}).
+	\begin{figure}[!ht]
+		\centering
+
+		\resizebox{8cm}{!}{
+			\begin{tikzpicture}
+				\node[draw] at (0,6.5) {\huge all touched false};
+				\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
+				\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
+
+				\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);
+
+				\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
+				\draw[fill=blue] (11,0)--(11,2)--(10,2)--(10,3)--(11,3)--(11,4)--(12,4)--(12,3)--(13,3)--(13,4)--(14,4)--(14,2)--(13,2)--(13,1)--(12,1)--(12,0)--cycle;
+			\end{tikzpicture}%
+		}%
+		\vspace*{5mm}
+		\resizebox{8cm}{!}{
+			\begin{tikzpicture}
+				\node[draw] at (0,6.5) {\huge all touched true};
+				\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
+				\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
+
+				\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);
+
+				\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
+				\draw[fill=blue] (10,0)--(10,5)--(11,5)--(11,4)--(15,4)--(15,1)--(13,1)--(13,0)--cycle;
+			\end{tikzpicture}%
+		}%
+		\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with weither all-touched False (above), or all touched True (below).}
+		\label{fig:alltouched}
+
+	\end{figure}
 
@@
 	\section{Results}
-	In this chapter we want to show the different cost rasters, that were created from the very same set of layers, but computed for different resolutions.
-From this different rasters the cost paths are calculated and compared.
-In the last step the rasters with lower resolution were used to calculate in a way, that they shall result in similar paths as if the paths were computed from a high resolution raster.
-	\subsection{Cost Rasters}
+	In this chapter we want to show the different cost raster, that were created from the very same set of layers,
+	but computed for different resolutions.
+	From this different raster the cost paths are calculated and compared.
+	In the last step the raster with lower resolution were used to calculate in a way, that they shall result in
+	similar paths as if the paths were computed from a high resolution raster.
+	\subsection{Cost Raster}
 	\subsection{Least Cost Paths}
+	For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
+	and all\_touched True.
+	The distance of both paths is calculated with the mean minimum distance.
+	For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
+	is computed and afterwards the minimum distances are averaged (see equation~\ref{equ:1}).
+	\begin{equation}
+		\label{equ:1}
+		d_{mean} = \frac{\sum_{i=1}^{n} d_{min}(p_i, L_2)}{n} \bigg\vert p_i \in L_1
+	\end{equation}
+
+	Hence, the mean minimum euclidean distance between the two paths can used to compute
+	the similarity.
+	As different table~\ref{table:2} shows the distance between the two paths decreases
+	with increasing resolution.
+	At the same time the differences in the aggregated costs stay constant.
+	\todo{Interpretation} When normalize the aggregated costs by the resolution.
+	Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
+	linear with the resolution.
+	On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
+	the resolution.
+	Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}
+
+	\begin{table*}[t]
+	\caption{Least cost paths as length for the different \acrfull{res} of the rasters, including the \acrfull{mmd} and the \acrfull{agg} costs.
+From the \acrshort{agg} costs the differences of the \acrshort{agg} costs and the \acrfull{corr} \acrshort{agg} by resolution are given.}
+	\label{table:2}
+	\centering
+	\begin{tabular}{ r  r  r  r  r  r  r  r  r}
+		\acrshort{res} /m & $length_{\acrshort{al}=\acrshort{f}} /m$ & $length_{\acrshort{al}=\acrshort{t}} /m$ & \acrshort{mmd} /m & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{t}} $ \\
+		\hline
+		5 & 76136.27	& 78002.00 & 126.04  & 18665.923 & 19616.756 & -850 & 93329.6 &  97584.77 \\
+		10 & 75430.10 	& 77936.57 & 277.92 & 8931.245 & 9731.175 & -799.95 & 89312.45 & 97311.75 \\
+		50 & 76135.02	& 70619,95 & 1140.01 & 1409.023 & 2300.073 & -891.05 & 70451.15 & 115003.65 \\
+		100 & 76283.80	& 74120.73 & 1946.41 & 640.516 & 1572.268 & -931.7 & 64051.6 & 167226.8 \\
+
+	\end{tabular}
+	\end{table*}
+
+	Comparing the all\_touched True rasterization and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterization at the
+	different resolutions the later paths are more similar.
+	The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m.
+
+	The similarity of all all\_touched False runs is higher than, the similar between the all\_touched False and all\_touched True runs with the same resolution, except for the highest resolution.
+
+	Hence, in an overall perspective paths of the all\_touched False runs stay in a
+	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
+	On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
+	But the extent is smaller.
+	The length of the the paths only differs to a maximum of about 10 \%.
+	The least path distance for higher resolutions can be lower, because new paths, between regions that are forbidden
+	in higher resolution can open.
+	On the other side the length of the paths can increase, because with higher resolution more vertices will be used.
+
+	The zonal stat (see table\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
+	percentage of which costs levels are around the path.
+	When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
+	Preferential Level and less the NoRestriction Level.\todo{Interpretation:}
+	The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
+	\todo{Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100 m! No problem because buffering is included in the cost raster.}
+	There is no strong tendency for the all\_touched False least cost paths.
+
+	\todo{used regions}
+
+	\setlength{\tabcolsep}{10pt}
+
+	\begin{table*}[t]
+		\caption{\acrfull{r} Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
+		\label{table:3}
+		\centering
+		\begin{tabular}{ r  r  r r  r r  r r  r r  r r}
+			resolution /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
+			\hline
+			5 & False &  4.7  &  (5.4) & 58.7 & (58.9) & 8.8 & (8.4) & 0.7 & (0.7) & 27.1 & (26.7)  \\
+
+			10 & False &  19.6 & (33.5) & 68.5 & (64.5)  & 1.0 & (0.8) & 0.8 & (0.3) & 10.1 & (0.9)\\
+
+			50 & False &  20.4 & (33.2) & 68.0 & (66.2)  & 0.9 & (0.1) & 0.7 & (0.0) & 10.1 & (0.5)\\
+
+			100 & False &  21.1 & (30.7) & 69.1 & (68.8)  & 1.1 & (0.0) & 0.7 & (0.0) & 7.9 & (0.4) \\
+
+			\hline
+
+			5 & True  &  18.9 & (28.5) & 67.3 & (66.4) & 1.3 & (1.6) & 1.0 & (0.5) & 11.5 & (3.0) \\	
+			10 & True &  18.9 & (33.7) & 66.6 & (63.4)  & 1.6 & (1.4) & 1.4 & (0.6) & 11.5 & (1.0)\\	
+			50 & True &  9.1 & (13.0) & 75.7 & (83.0) & 3.9 & (2.0) & 4.2 & (1.6) & 7.1 & (0.4) \\
+			100 & True &  7.0 & (10.1) & 73.8 & (81.9)  & 5.5 & (3.9) & 8.5 & (3.6) & 5.2 & (0.4) \\
+
+		\end{tabular}
+	\end{table*}
+
+	\todo{Check why values are so different for 5 m all touched False --> simplified line.}
+
 	\subsection{Faster Processing of the Cost Path Algorithm}
+	The original least cost path QGIS plug-in uses stops the least aggregation of the costs after the final end point has been found.
+	This early stopping might result in sub optimal paths around the end points.
+	For further calculation the search for the least cost paths has to be refactored and split the aggregation of the raster for the search of the path.
+	To be able to construct a full aggregated cost raster, aggregated cost path in further steps will be computed completely.\todo{really needed}
+	For the calculation of the least cost paths.
+	The finally least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer
+	The first step to optimize the calculation speed is, only the calculated the least cost paths for this smaller area. 
+	\subsubsection{Apply a minimum sized bounding box around least cost paths}
+	\subsubsection{Apply a convex hull round a the last cost paths}
+	\subsubsection{Remove unreachable points}
 
 	\section{Discussion}
-	\todo{discuss aggrgation with max vs sum}
+	\todo{discuss aggregation with max vs sum}
 	\section{Related Works}

### 2023-01-09 paper/paper-CableRoute.tex
updated paper:
added new ideas for next steps.

@@
 	For every area in the study area, that is not covered by any layer, is given the default cost.
-	The costs has been grouped into different levels~(see table~\ref{table:1}) starting from preferential areas with
+	The costs has been grouped into different levels~(see table~\ref{tab:1}) starting from preferential areas with
 	very low costs, via no restrictions, which is the default, used when no other layers are covering the local area,
@@
 	the weight that resembles the costs for these types of areas, has to be especially high. \\
-	The rasterziation transforms a vector into a raster.
+	The rasterization transforms a vector into a raster.
 	The rasterization can be executed in to different ways.
@@
 	Both rasterization techniques differ  in who the pixels are selected, that describe, that a geo-object overlays the pixel.
-	The pixels can either be described as overlayed with a geo-object, when either the center of that pixel is overlayed
-	by the geo-object, or any part of the pixel is overlayed.
-	Hence, the version with any part overlayed is called all touched True.
-	The version where a overlayed pixel center is require is thus called all touched False.
+	The pixels can either be described as overlaid with a geo-object, when either the center of that pixel is overlaid
+	by the geo-object, or any part of the pixel is overlaid.
+	Hence, the version with any part overlaid is called all touched True.
+	The version where a overlaid pixel center is require is thus called all touched False.
 	All touched False is considered the default (see figure~\ref{fig:alltouched}).
@@
 		\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
-		\label{table:1}
+		\label{tab:1}
 		\centering
@@
 	For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
-	is computed and afterwards the minimum distances are averaged (see equation~\ref{equ:1}).
+	is computed and afterwards the minimum distances are averaged (see equation~\ref{eq:1}).
 	\begin{equation}
-		\label{equ:1}
+		\label{eq:1}
 		d_{mean} = \frac{\sum_{i=1}^{n} d_{min}(p_i, L_2)}{n} \bigg\vert p_i \in L_1
@@
 	The first step to optimize the calculation speed is, only the calculated the least cost paths for this smaller area. 
-	\subsubsection{Apply a minimum sized bounding box around least cost paths}
-	\subsubsection{Apply a convex hull round a the last cost paths}
-	\subsubsection{Remove unreachable points}
+	\subsubsection{Compare least cost paths, for overlay of both rasterizations}
+	\subsubsection{Restrict search to a minimum sized bounding box}
+	\subsubsection{Restrict search to a convex hull}
+	\subsubsection{Restrict search to reachable points}
+
+	\subsubsection{Compare the different Solutions}
+	\subsubsection{Check solution for least cost path between different points}
 
@@
 	\todo{discuss aggregation with max vs sum}
+	\todo{discuss other methods. Off algorithmic speed-ups}
+	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
 	\section{Related Works}

### 2023-01-11 paper/paper-CableRoute.tex
updated paper:
added maps that shows least cost paths for combinations different resolutions and values for parameter all touched.

@@
 \usepackage{multirow}
+\usepackage{subfig}
 \usepackage[acronym]{glossaries}
@@
 	As different table~\ref{table:2} shows the distance between the two paths decreases
-	with increasing resolution.
+	with increasing resolution. In addition, this tendency is depicted in figure~\ref{fig:paths_alltouched} for the calculated cost paths of 5m and 100 m resolutions.
+
 	At the same time the differences in the aggregated costs stay constant.
@@
 	Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}
+
+	\begin{figure}
+		\centering
+
+		\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m.pdf} }}%
+		\qquad
+		\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
+		\caption{Figures of the least cost paths. Contrasting the paths for different resolutions. Using OpenStreetMaps as base map.}
+		\label{fig:paths_resolution}
+	\end{figure}
+
+	\begin{figure}
+	\centering
+
+	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F.pdf} }}%
+	\qquad
+	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T.pdf} }}%
+
+	\caption{Figures of the least cost paths. Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched. Using OpenStreetMaps as base map.}
+	\label{fig:paths_alltouched}
+\end{figure}
 
@@
 	different resolutions the later paths are more similar.
-	The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m.
+	The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m. \todo{Option add 2.5 m and 25 m}
 
@@
 	Hence, in an overall perspective paths of the all\_touched False runs stay in a
-	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
+	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs. This behaviour is depicted in figure~\ref{fig:paths_alltouched}. 
 	On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
@@
 
-	The zonal stat (see table\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
+	The zonal stat (see table~\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
 	percentage of which costs levels are around the path.
@@
 	\begin{table*}[t]
-		\caption{\acrfull{r} Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
+		\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
 		\label{table:3}

### 2023-01-11 paper/paper-CableRoute.tex
updated paper:
minor fixes

@@
 \usepackage{tikz}
+\usepackage{graphicx}
 
@@
 	the similarity.
-	As different table~\ref{table:2} shows the distance between the two paths decreases
-	with increasing resolution. In addition, this tendency is depicted in figure~\ref{fig:paths_alltouched} for the calculated cost paths of 5m and 100 m resolutions.
+	As different table~\ref{tab:2} shows the distance between the two paths decreases
+	with increasing resolution.
+	In addition, this tendency is depicted in figure~\ref{fig:paths_resolution} for the calculated cost paths of 5m and 100 m resolutions.
 
@@
 From the \acrshort{agg} costs the differences of the \acrshort{agg} costs and the \acrfull{corr} \acrshort{agg} by resolution are given.}
-	\label{table:2}
+	\label{tab:2}
 	\centering
@@
 	Hence, in an overall perspective paths of the all\_touched False runs stay in a
-	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs. This behaviour is depicted in figure~\ref{fig:paths_alltouched}. 
+	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
+	This behaviour is depicted in figure~\ref{fig:paths_alltouched}.
 	On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
@@
 	\todo{discuss aggregation with max vs sum}
+	\todo{Scaling with multiple endpoints.}
 	\todo{discuss other methods. Off algorithmic speed-ups}

### 2023-01-14 paper/paper-CableRoute.tex
paper:  - added subchapter cost raster  - updated, added maps  - updated paragraph in Faster Processing

@@
 \usepackage{graphicx}
+\usepackage{makecell}
 
@@
 	%%
-	%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
+	%% The code below is generated by the tool at https://dl.acm.org/ccs.cfm.
 	%% Please copy and paste the code instead of the example below.
@@
 
-	\section{Introduction}
+	\section{Introduction}\label{sec:introduction}
 
@@
 
-	\section{Methods}
+	\section{Methods}\label{sec:methods}
 	We retrieve a set of different spacial data-sets from  public sources as a basis to create the cost raster.
@@
 	The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
-	and Geodesy\cite{noauthor_digitales_2021}.
-	The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap.\cite{boeing_osmnx_2017}
+	and Geodesy~\cite{noauthor_digitales_2021}.
+	The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap~\cite{boeing_osmnx_2017}.
 	Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
@@
 	PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
-	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin
+	The the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin
 	'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
@@
 			\hline
-			Prohibited & 500					& National Parks, Buildings \\
-			strongly Restricted & 10 	& Bird Reserve \\
-			Restricted & 5	& Industrial Areas \\
+			Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
+			strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
+			Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
 			No Restriction 		& 0.5					& Default\\
-			Preferential & 0.1					& Power Grid, Motorway Buffers\\
+			Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
 		\end{tabular}
@@
 
-	\section{Results}
+	\section{Results}\label{sec:results}
 	In this chapter we want to show the different cost raster, that were created from the very same set of layers,
@@
 	similar paths as if the paths were computed from a high resolution raster.
-	\subsection{Cost Raster}
-	\subsection{Least Cost Paths}
+	\subsection{Cost Raster}\label{subsec:cost-raster}
+
+	The cost raster contains all the costs as weights for the geographic region of the study area.
+	The regions outside the study area are given, a no-data value and will not be use for the calculation of the least cost path.
+	It is decided to use the maximum value, for any place in the raster, that is covered by several rules at the same time.
+	When the resolution that is use for the rasterization is smaller than the object size, than the effect of the rasterization with all touched True or False is limited.
+	For all touched True any part of the pixel that is covered by the object, will attribute that whole pixel to the object.
+	Thus, the object appears to be halve a pixel size larger in all directs.
+	As can be seen in figure~\ref{fig:costs_5m} that shows a detailed view of the costs for village of Beverstedt.
+	Due to the maximum aggregation of the costs, the average cost of the raster of all touched true will be over estimated.
+	All touched false will be a better description of the real size of the object.
+	\begin{figure}
+		\centering
+
+		\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alT.png} }}%
+		\qquad
+		\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alF.png} }}%
+		\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 100~m.}
+		\label{fig:costs_5m}
+	\end{figure}
+
+	In contrast when the resolution the is much less, than the size of the object the described behavior changes.
+	On one hand, while the area, of the pixel is increased for all touched True, also the area for which the cost is overestimated increases.
+	On the other hand while the pixel size for all touched  False increases, not only the over estimated area increases for that pixel as describes for all touched true, but in addition the object is less probable situated in the center of the pixel.
+	The consequence of the pixel is, that will decreasing resolution that object is not included in the rasterziation.
+	Hence, hence for all touched false lower resolution  leads ot a loss of information.
+	Because the default is relative small compared the over effect is a underestimation of the costs.
+	The figure~\ref{fig:costs_100m} shows for the resolution of 100 m, that while larger are still included in the map they appear to be much larger.
+	On the contrary smaller objects might not be included.
+	Objects that are small on in one dimension as streets will be included in a stochastic manner.
+	Described by the likely of that a object of that sized overlaps with the pixel center.
+	\begin{figure}
+		\centering
+
+		\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alT.png} }}%
+		\qquad
+		\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alF.png} }}%
+		\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 100~m.}
+		\label{fig:costs_100m}
+	\end{figure}
+	Thus, with having larger areas and more objects, with higher costs, all touched True rasterization might more likely lead to longer roots and more likely in blocking the  direct spatial path.
+
+	\subsection{Least Cost Paths}\label{subsec:least-cost-paths}
 	For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
@@
 		\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
-		\caption{Figures of the least cost paths. Contrasting the paths for different resolutions. Using OpenStreetMaps as base map.}
+		\caption{Figures of the least cost paths.
+Contrasting the paths for different resolutions.
+Paths computed from all touched false raster are indicated by dashed lines.
+Results from all touched True are indicuated by continuous lines.
+Higher resolutions are indicated by the color green, lower resolutions by the color red.
+Using OpenStreetMaps as base map.}
 		\label{fig:paths_resolution}
@@
 
-	\caption{Figures of the least cost paths. Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched. Using OpenStreetMaps as base map.}
+	\caption{Figures of the least cost paths.
+Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched.
+Paths computed from all touched false raster are indicated by dashed lines.
+Results from all touched True are indicuated by continuous lines.
+Higher resolutions are indicated by the color green, lower resolutions by the color red.
+Using OpenStreetMaps as base map.}
 	\label{fig:paths_alltouched}
@@
 		\hline
-		5 & 76136.27	& 78002.00 & 126.04  & 18665.923 & 19616.756 & -850 & 93329.6 &  97584.77 \\
+		5 	& 76136.27	& 78002.00 &  126.04 & 18665.923 & 19616.756 & -850.00 & 93329.60 &  97584.77 \\
 		10 	& 75430.10 	& 77936.57 &  277.92 &  8931.245 &  9731.175 & -799.95 & 89312.45 &  97311.75 \\
 		50 	& 76135.02	& 70619,95 & 1140.01 &  1409.023 &  2300.073 & -891.05 & 70451.15 & 115003.65 \\
-		100 & 76283.80	& 74120.73 & 1946.41 & 640.516 & 1572.268 & -931.7 & 64051.6 & 167226.8 \\
+		100 & 76283.80	& 74120.73 & 1946.41 &   640.516 &  1572.268 & -931.70 & 64051.60 & 167226.80 \\
 
@@
 
-	The zonal stat (see table~\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
+	The zonal stat (see table~\ref{tab:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
 	percentage of which costs levels are around the path.
@@
 		\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
-		\label{table:3}
+		\label{tab:3}
 		\centering
@@
 
-	\subsection{Faster Processing of the Cost Path Algorithm}
-	The original least cost path QGIS plug-in uses stops the least aggregation of the costs after the final end point has been found.
-	This early stopping might result in sub optimal paths around the end points.
-	For further calculation the search for the least cost paths has to be refactored and split the aggregation of the raster for the search of the path.
-	To be able to construct a full aggregated cost raster, aggregated cost path in further steps will be computed completely.\todo{really needed}
-	For the calculation of the least cost paths.
-	The finally least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer
-	The first step to optimize the calculation speed is, only the calculated the least cost paths for this smaller area. 
+	\subsection{Faster Processing of the Cost Path Algorithm}\label{subsec:faster-processing-of-the-cost-path-algorithm}
+
+	The final least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer to the paths resulted by the all touched false rasterization.
+	The first step is to optimize the calculation speed is, only to calculated the least cost paths for this smaller area.
+	Another method, is to improve the prediction of the medium resolution itself and thus reduce the need for a computation in higher resolution.
 	\subsubsection{Compare least cost paths, for overlay of both rasterizations}
-	\subsubsection{Restrict search to a minimum sized bounding box}
+	As all touched true rasterization overestimates and all touched true underestimates the real costs.
+	A weighted mean will describe the real costs more precise.
+	As present work indicates, that the weight should favor the all touched false raster.
+	The best weight should be the percentage of the pixel, which is covered by the object, but this can not be computed in this work.
+	Thus, the best weight has to be searched.
+	An alternative might be to compute the cost raster in higher resolution, but than to reproject the to a smaller resolution with a using a (linear) interpolation of the of the weights.\todo{test this idea}
+	% \subsubsection{Restrict search to a minimum sized bounding box}
 	\subsubsection{Restrict search to a convex hull}
+	\todo{Construct a polygon form the two lines. Buffer the polygon.}
 	\subsubsection{Restrict search to reachable points}
+	For this purpose the aggregated cost as to converted back into a raster.
 
@@
 
-	\section{Discussion}
+	\section{Discussion}\label{sec:discussion}
+	As shown need of computation power increases with the power of two with the resolution.
+	While the error only scales linear with the resolution.
+
+	As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
+	This early stopping might result in sub optimal paths around the end points.
+
+	As a any buffer around buildings are describes as protected areas, power poles are also included.
+	In further works, these should be excluded.
+
 	\todo{discuss aggregation with max vs sum}
@@
 	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
-	\section{Related Works}
-	\section{Conclusion}
+	\section{Related Works}\label{sec:related-works}
+
+	\section{Conclusion}\label{sec:conclusion}
 
@@
 
-\section{Research Methods}
+\section{Research Methods}\label{sec:research-methods}
 
-\subsection{Part One}
+	\subsection{Part One}\label{subsec:part-one}
 
-\subsection{Part Two}
+	\subsection{Part Two}\label{subsec:part-two}
 
-\section{Online Resources}
+	\section{Online Resources}\label{sec:online-resources}
 

### 2023-01-28 src/least_cost_path/dijkstra.py
added missing requirements, typos and types
linear regression of estimation time and total costs

@@ -10,7 +10,7 @@ def dijkstra(start_tuple: tuple[tuple[int, int], Point, int],
              block: list[list[float | None]],
              find_nearest: bool) -> list[tuple[list[tuple[int, int]], list[float | None], list]]:
     """
-    compute the distance between the starting point and any point of he
+    compute the distance between the starting point and any point of the end_list
     :param start_tuple: Starting point (2-tuple of int, in coords of `block`) of the aggregation/length calculation
     :param end_tuples: Ending points(list of 2-tuple of ints) that should be reached
     :param block: The costs/weights as 2S-List of float | None (None == nodata)
@@ -106,7 +106,7 @@ def dijkstra(start_tuple: tuple[tuple[int, int], Point, int],
             continue
         decided.add(current_node)
 
-        # reached destination
+        # reached destination, potential break
         if current_node in end_row_cols:
             path = []
             costs = []
@@ -129,7 +129,7 @@ def dijkstra(start_tuple: tuple[tuple[int, int], Point, int],
             if len(end_row_cols) == 0 or find_nearest:
                 break
 
-        # relax distance
+        # relax distance, aggregate the costs
         for nex in grid.neighbors(current_node):
             new_cost = cost_so_far[current_node] + grid.simple_cost(current_node, nex)
             if nex not in cost_so_far or new_cost < cost_so_far[nex]:

### 2023-01-28 src/least_cost_path/find_least_cost_path.py

@@ -135,8 +135,8 @@ def find_least_cost_path(cost_raster: DataArray, cost_raster_band: int, is_neare
     :param cost_raster: DataArray
     :param cost_raster_band: int
     :param is_nearest: bool
-    :param start_features: GeoSeries
-    :param end_features: GeoSeries
+    :param start_features: GeoDataFrame
+    :param end_features: GeoDataFrame
     :return: GeoDataFrame
     """
     raster_2d = cost_raster[cost_raster_band]

### 2023-01-28 src/lod1_load/download_lod1.py

@@ -37,7 +37,7 @@ def concat_vector_files(in_dir: Path) -> GeoDataFrame:
     for f in shp_files:
         try:
             lod1_vectors.append(read_file(f))
-        except UnsupportedGeometryTypeError as e:
+        except UnsupportedGeometryTypeError:
             error_counter = error_counter + 1
     if error_counter > 0:
         warn(f'{error_counter} could not be concatenated.')

### 2023-01-28 src/wps/flask/pywps_flask.py

@@ -15,7 +15,7 @@ def outputfile(filename: str):
     Send the files with the given json file in folder `/tmp`.
     `/tmp` has to be configured in `.cfg` file for PyWPS as outputpath.
     :param filename: str, name of the json file
-    :return: falsk.Resonse or abort (None) transmitting the input json file
+    :return: flask.Response or abort (None) transmitting the input json file
     """
     print(filename)
     target_file = Path('/tmp') / filename

### 2023-01-28 src/wps/processes/least_cost_path.py

@@ -10,7 +10,7 @@ class LeastCostPath(Process):
     Process Class to compute the least cost path from input.
     __init__ describes the data.
     _handle computes the least cost path.
-    Class instiacted by server.
+    Class instanced by server.
     """
     def __init__(self):
         inputs = [ComplexInput('costs', 'Cost Raster', supported_formats=[Format('image/tiff'), ]),

### 2023-02-04 src/least_cost_path/dijkstra.py
update .gitignore with ide and latex file extensions
dijkstra.py
- refactor
expose the GRID class to the module
split the dijkstra function in a dijkstra generator (for all end positions) and a backtrack_path function (a call for each end position)

find_least_cost_path.py
 -  add matrix2raster function that converts the aggregated cost map dict yield by the dijkstra generator into a raster
 - find_least_cost_path function needed to be adapted to the refactor

 interpolate_costs.ipynb
 - jupyter notebook compute the cost raster
  1) raster resized with bilinear interpolation
  2) mixed all tou

@@ -5,86 +5,87 @@ import collections
 from shapely.geometry import Point
 
 
+class Grid:
+    def __init__(self, matrix: list[list[float]]):
+        self.map = matrix
+        self.h = len(matrix)
+        self.w = len(matrix[0])
+
+    def _in_bounds(self, _id: tuple[int, int]) -> bool:
+        """
+        Is the tuple inside the `block` coordinates
+        :param _id: tuple[int, int] coords of the point in `block` coords
+        :return: bool
+        """
+        x, y = _id
+        return 0 <= x < self.h and 0 <= y < self.w
+
+    def _passable(self, _id: tuple[int, int]) -> bool:
+        """
+        Can this point be reached (is a valid point, with valid weight?)
+        :param _id: tuple[int, int] coords of the point in `block` coords
+        :return: bool
+        """
+        x, y = _id
+        return self.map[x][y] is not None
+
+    def is_valid(self, _id: tuple[int, int]) -> bool:
+        """
+        Can this inside the block and can be reached
+        :param _id:tuple[int, int] coords of the point in `block` coords
+        :return: bool
+        """
+        return self._in_bounds(_id) and self._passable(_id)
+
+    def neighbors(self, _id: tuple[int, int]) -> list[tuple[int, int]]:
+        """
+        Compute all valid neighbors of the current point
+        :param _id: tuple[int, int] coords of the point in `block` coords
+        :return: list[tuple[int, int]]
+        """
+        x, y = _id
+        results = [(x + 1, y), (x, y - 1), (x - 1, y), (x, y + 1),
+                   (x + 1, y - 1), (x + 1, y + 1), (x - 1, y - 1), (x - 1, y + 1)]
+        results = list(filter(self.is_valid, results))
+        return results
+
+    def simple_cost(self, current: tuple[int, int], _next: tuple[int, int]) -> float:
+        """
+        Compute the euclidean distance between current and _next point
+        :param current: tuple[int, int] coords of the point in `block` coords
+        :param _next: tuple[int, int] coords of the point in `block` coords
+        :return: float
+        """
+        cx, cy = current
+        nx, ny = _next
+        current_value = self.map[cx][cy]
+        offset_value = self.map[nx][ny]
+        if cx == nx or cy == ny:
+            return (current_value + offset_value) / 2
+        else:
+            return sqrt(2) * (current_value + offset_value) / 2
+
+
 def dijkstra(start_tuple: tuple[tuple[int, int], Point, int],
[…]

### 2023-02-04 src/least_cost_path/find_least_cost_path.py

@@ -1,12 +1,15 @@
 from math import floor
 
+import numpy as np
 from geopandas import read_file, GeoSeries, GeoDataFrame
 from pandas import concat
+from pathlib import Path
 from rioxarray import open_rasterio
 from shapely.geometry import Point, LineString, MultiPoint, box
 from xarray import DataArray
 
-from src.least_cost_path.dijkstra import dijkstra
+from src.least_cost_path.dijkstra import dijkstra, backtrack_path
+from src.rules2weights import write_compressed
 from argparse import ArgumentParser
 
 
@@ -93,6 +96,25 @@ def raster2matrix(block: DataArray) -> tuple[list[list[None | float]], bool]:
     return _matrix, _contains_negative
 
 
+def matrix2raster(matrix: dict, alike_raster: DataArray, nodata: None | float = None) -> DataArray:
+    """
+    Convert the 2D-List back into the original raster format
+    :param matrix: list[list[None | float]] the 2D-List to convert
+    :param alike_raster: the raster with the correct crs, shape, nodata
+    :param nodata: if given: value to use as nodata
+    :return: DataArray
+    """
+    return_raster = alike_raster.copy()
+    return_raster = return_raster.rio.write_nodata(nodata)
+    return_raster.data = np.full_like(return_raster.data, fill_value=return_raster.rio.nodata)
+
+    for i in range(alike_raster.rio.height):
+        for j in range(alike_raster.rio.width):
+            if (i, j) in matrix:
+                return_raster[i, j] = matrix[(i, j)]
+    return return_raster
+
+
 def create_points_from_path(_cost_raster: DataArray, min_cost_path: list[tuple[int, int]], start_point: Point,
                             end_point: Point) -> list[Point]:
     """
@@ -129,14 +151,15 @@ def create_path_feature_from_points(_path_points: list[Point], attr_vals: tuple[
 
 
 def find_least_cost_path(cost_raster: DataArray, cost_raster_band: int, is_nearest: bool, start_features: GeoDataFrame,
-                         end_features: GeoDataFrame) -> GeoDataFrame:
+                         end_features: GeoDataFrame, aggregated_save: None | str = None) -> GeoDataFrame:
     """
     Compute the Least Cost Path with Dijkstra Algorithm
     :param cost_raster: DataArray
     :param cost_raster_band: int
     :param is_nearest: bool
-    :param start_features: GeoDataFrame
-    :param end_features: GeoDataFrame
[…]

### 2023-02-04 src/rules2weights.py

@@ -19,7 +19,7 @@ from tqdm import tqdm
 from xarray import DataArray
 
 
-def write_compress(r: DataArray, save: Path) -> None:
+def write_compressed(r: DataArray, save: Path) -> None:
     """
     Saving the raster r at save
     :param r: raster as DataArray
@@ -52,7 +52,7 @@ def __filter(total_gdf: GeoDataFrame, column: str, _rule: Node, ) -> GeoDataFram
     :param total_gdf: original GeoDataFrame
     :param column: column name (str) from which the attributes will be used
     :param _rule: rule (as Node) that applies for the values in the column of the GeoDataFrame
-    :return: GeoDataFrame
+    :return: GeoDataFrame with the filter applied
     """
     child_results: list[GeoDataFrame] = []
     for item in _rule.items:
@@ -241,7 +241,7 @@ def process_base_rule(df: DataFrame, _res: float | tuple[float, float], _crs: in
     raster = raster['Weights']
 
     save_path.parent.mkdir(parents=True, exist_ok=True)
-    write_compress(raster, save_path)
+    write_compressed(raster, save_path)
     return raster
 
 
@@ -259,7 +259,7 @@ def process_rule(nmd_tpl, example_da: DataArray, _crs: int,
     :return: list[str| Path]
     """
     # resolution from example_Data
-    res = int(abs(example_da.rio.transform()[0]))
+    res = float(abs(example_da.rio.transform()[0]))
 
     save_path = _save_dir / f"res_{res}/all_touched_{_all_touched}/rule_{nmd_tpl.LineNum}_{nmd_tpl.Description}.tif"
     if not save_path.exists():  # reuse existing results
@@ -269,7 +269,7 @@ def process_rule(nmd_tpl, example_da: DataArray, _crs: int,
         raster = raster['Weights']
         raster.data[example_da == example_da.rio.nodata] = raster.rio.nodata
         save_path.parent.mkdir(parents=True, exist_ok=True)
-        write_compress(raster, save_path)
+        write_compressed(raster, save_path)
     return [save_path, ]
 
 
@@ -278,7 +278,7 @@ if __name__ == '__main__':
     parser.add_argument('config_file_path', type=str, help='Full path to the .xlsx config file.')
     parser.add_argument('vectors_main_folder', type=str, help='Main Path to the vector files.')
     parser.add_argument('save_dir', type=str, help='Path where this projects raster files can the saved')
-    parser.add_argument('-r', '--resolution', type=int, default=1000, help='Resolution to use.')
[…]

### 2023-02-05 paper/methods.tex
refactor paper-CableRoute.tex to exclude the larger sections results and methods.
These were moved into there own files

--- 2023-01-11 paper/paper-CableRoute.tex
@@
 
-\documentclass[acmtog]{acmart}
-\usepackage{amsmath}
-\usepackage{multirow}
-\usepackage{subfig}
-\usepackage[acronym]{glossaries}
-\usepackage{todonotes}
-\usepackage{tikz}
-
-\AtBeginDocument{%
-	\providecommand\BibTeX{{%
-			Bib\TeX}}}
-
-\setcopyright{acmcopyright}
-\copyrightyear{2022}
-\acmYear{2022}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%%
-%% These commands are for a JOURNAL article.
-\acmJournal{JACM}
-\acmVolume{37}
-\acmNumber{4}
-\acmArticle{28}
-\acmMonth{2}
-
-\newacronym{wfs}{wfs}{Web Feature Service}
-\newacronym{wps}{wps}{Web Processing Service}
-\newacronym{res}{res}{resolution}
-\newacronym{r}{r}{ratio}
-\newacronym{al}{al}{all touched}
-\newacronym{t}{t}{True}
-\newacronym{f}{f}{False}
-\newacronym{mmd}{mmd}{mean minimum distance}
-\newacronym{agg}{agg}{aggregated}
-\newacronym{rel}{rel}{relative}
-\newacronym{corr}{corr}{corrected}
-
-\begin{document}
-	\title{Web Processing - Standardised GIS Analyses for Cable Route Planning}
-
-	\author{Sebastian Heiden}
-	\email{u38439@hs-harz.de}
-	\affiliation{%
-		\institution{Harz University of Applied Sciences}
-		\streetaddress{Friedrichstrasse 57-59}
-		\city{Wernigerode}
-		\state{Saxony-Anhalt}
-		\country{Germany}
-		\postcode{38855}
-	}
-
-	\renewcommand{\shortauthors}{Heiden}
-
-	\begin{abstract}
-		add as final part
-	\end{abstract}
-
-	%%
-	%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-	%% Please copy and paste the code instead of the example below.
-	%%
-	\begin{CCSXML}
-		<ccs2012>
-		<concept>
-		<concept_id>10010520.10010553.10010562</concept_id>
-		<concept_desc>Computer systems organization~Embedded systems</concept_desc>
-		<concept_significance>500</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10010520.10010575.10010755</concept_id>
-		<concept_desc>Computer systems organization~Redundancy</concept_desc>
-		<concept_significance>300</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10010520.10010553.10010554</concept_id>
-		<concept_desc>Computer systems organization~Robotics</concept_desc>
-		<concept_significance>100</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10003033.10003083.10003095</concept_id>
-		<concept_desc>Networks~Network reliability</concept_desc>
-		<concept_significance>100</concept_significance>
-		</concept>
-		</ccs2012>
-	\end{CCSXML}
-
-	\ccsdesc[500]{Computer systems organization~Embedded systems}
-	\ccsdesc[300]{Computer systems organization~Redundancy}
-	\ccsdesc{Computer systems organization~Robotics}
-	\ccsdesc[100]{Networks~Network reliability}
-
-	%%
-	%% Keywords. The author(s) should pick words that accurately describe
-	%% the work being presented. Separate the keywords with commas.
-	\keywords{datasets, neural networks, gaze detection, text tagging}
-
-	\received{20 February 2007}
-	\received[revised]{12 March 2009}
-	\received[accepted]{5 June 2009}
-
-	%%
-	%% This command processes the author and affiliation and title
-	%% information and builds the first part of the formatted document.
-	\maketitle
-
-	\section{Introduction}
-
-	Sometimes, finding the shortest path is not sufficient.
-	Additional parameters play also have to be taken into consideration.
-	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline
-	\todo{source}.
-	For planing the additional routes for a power grid, additional aspects as legal regulations and acceptance
-	by the local population have to be taken in consideration.
-	Also technical aspects, as the effects on the grid stability might be further points to consider.\cite{schafer_understanding_2022}
-	\todo{what is the least cost path}
-
-	\section{Methods}
 We retrieve a set of different spacial data-sets from  public sources as a basis to create the cost raster.
@@
 
-	The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
-	and Geodesy\cite{noauthor_digitales_2021}.
-	The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap.\cite{boeing_osmnx_2017}
+The nationwide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
+and Geodesy~\cite{noauthor_digitales_2021}.
+The nationwide power grid (tags: 'power': line) has been retrieved via OpenStreetMap~\cite{boeing_osmnx_2017}.
 Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
 Lower Saxony\cite{noauthor_opengeodatani_2022}.
-	In addition local planning geodata for the land usage are taken from
+In addition, local planning geodata for the land usage are taken
 from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)\cite{noauthor_metropolplaner_2022}
@@
 PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
-	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin
-	'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
+The initial implementation of the least cost path algorithm is based on the implementation for the QGIS-Plugin
+'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.
 
-	The different layers from the different entities are optionally filtered, buffered and than rasterized.
-	Filtering the layers of the files for special attributes enables to further differentiate further.
+The different layers from the different entities are optionally filtered, buffered and then rasterized.
+Filtering the layers of the files for special attributes enables to further differentiate.
 For examples makes it possible to differentiate between heath and uncultivated land.
@@
 For every area in the study area, that is not covered by any layer, is given the default cost.
-	The costs has been grouped into different levels~(see table~\ref{tab:1}) starting from preferential areas with
+The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from preferential areas with
 very low costs, via no restrictions, which is the default, used when no other layers are covering the local area,
@@
 given resolution and the new affine transformation and the coordinate reference system of the vector.
-	Both rasterization techniques differ  in who the pixels are selected, that describe, that a geo-object overlays the pixel.
+Both rasterization techniques differ in how the pixels are selected, that describe, that a geo-object overlays the pixel.
 The pixels can either be described as overlaid with a geo-object, when either the center of that pixel is overlaid
@@
 		\hline
-			Prohibited & 500					& National Parks, Buildings \\
-			strongly Restricted & 10 	& Bird Reserve \\
-			Restricted & 5	& Industrial Areas \\
+		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
+		strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
+		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
 		No Restriction 		& 0.5					& Default\\
-			Preferential & 0.1					& Power Grid, Motorway Buffers\\
+		Preferential 		& 0.1					& \makecell[lt]{Power Grid,\\ Motorway and Railway Buffers}\\
 	\end{tabular}
@@
 The completed list of layers and the processing applied to them, can be found in Supplement S1. \todo{create the Supplement from the processing rules}
-
-	\section{Results}
-	In this chapter we want to show the different cost raster, that were created from the very same set of layers,
-	but computed for different resolutions.
-	From this different raster the cost paths are calculated and compared.
-	In the last step the raster with lower resolution were used to calculate in a way, that they shall result in
-	similar paths as if the paths were computed from a high resolution raster.
-	\subsection{Cost Raster}
-	\subsection{Least Cost Paths}
-	For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
-	and all\_touched True.
-	The distance of both paths is calculated with the mean minimum distance.
-	For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
-	is computed and afterwards the minimum distances are averaged (see equation~\ref{eq:1}).
-	\begin{equation}
-		\label{eq:1}
-		d_{mean} = \frac{\sum_{i=1}^{n} d_{min}(p_i, L_2)}{n} \bigg\vert p_i \in L_1
-	\end{equation}
-
-	Hence, the mean minimum euclidean distance between the two paths can used to compute
-	the similarity.
-	As different table~\ref{table:2} shows the distance between the two paths decreases
-	with increasing resolution. In addition, this tendency is depicted in figure~\ref{fig:paths_alltouched} for the calculated cost paths of 5m and 100 m resolutions.
-
-	At the same time the differences in the aggregated costs stay constant.
-	\todo{Interpretation} When normalize the aggregated costs by the resolution.
-	Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
-	linear with the resolution.
-	On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
-	the resolution.
-	Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}
-
-	\begin{figure}
-		\centering
-
-		\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m.pdf} }}%
-		\qquad
-		\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
-		\caption{Figures of the least cost paths. Contrasting the paths for different resolutions. Using OpenStreetMaps as base map.}
-		\label{fig:paths_resolution}
-	\end{figure}
-
-	\begin{figure}
-	\centering
-
-	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F.pdf} }}%
-	\qquad
-	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T.pdf} }}%
-
-	\caption{Figures of the least cost paths. Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched. Using OpenStreetMaps as base map.}
-	\label{fig:paths_alltouched}
-\end{figure}
-
-	\begin{table*}[t]
-	\caption{Least cost paths as length for the different \acrfull{res} of the rasters, including the \acrfull{mmd} and the \acrfull{agg} costs.
-From the \acrshort{agg} costs the differences of the \acrshort{agg} costs and the \acrfull{corr} \acrshort{agg} by resolution are given.}
-	\label{table:2}
-	\centering
-	\begin{tabular}{ r  r  r  r  r  r  r  r  r}
-		\acrshort{res} /m & $length_{\acrshort{al}=\acrshort{f}} /m$ & $length_{\acrshort{al}=\acrshort{t}} /m$ & \acrshort{mmd} /m & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{t}} $ \\
-		\hline
-		5 & 76136.27	& 78002.00 & 126.04  & 18665.923 & 19616.756 & -850 & 93329.6 &  97584.77 \\
-		10 & 75430.10 	& 77936.57 & 277.92 & 8931.245 & 9731.175 & -799.95 & 89312.45 & 97311.75 \\
-		50 & 76135.02	& 70619,95 & 1140.01 & 1409.023 & 2300.073 & -891.05 & 70451.15 & 115003.65 \\
-		100 & 76283.80	& 74120.73 & 1946.41 & 640.516 & 1572.268 & -931.7 & 64051.6 & 167226.8 \\
-
-	\end{tabular}
-	\end{table*}
-
-	Comparing the all\_touched True rasterization and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterization at the
-	different resolutions the later paths are more similar.
-	The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m. \todo{Option add 2.5 m and 25 m}
-
-	The similarity of all all\_touched False runs is higher than, the similar between the all\_touched False and all\_touched True runs with the same resolution, except for the highest resolution.
-
-	Hence, in an overall perspective paths of the all\_touched False runs stay in a
-	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs. This behaviour is depicted in figure~\ref{fig:paths_alltouched}. 
-	On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
-	But the extent is smaller.
-	The length of the the paths only differs to a maximum of about 10 \%.
-	The least path distance for higher resolutions can be lower, because new paths, between regions that are forbidden
-	in higher resolution can open.
-	On the other side the length of the paths can increase, because with higher resolution more vertices will be used.
-
-	The zonal stat (see table~\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
-	percentage of which costs levels are around the path.
-	When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
-	Preferential Level and less the NoRestriction Level.\todo{Interpretation:}
-	The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
-	\todo{Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100 m! No problem because buffering is included in the cost raster.}
-	There is no strong tendency for the all\_touched False least cost paths.
-
-	\todo{used regions}
-
-	\setlength{\tabcolsep}{10pt}
-
-	\begin{table*}[t]
-		\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
-		\label{table:3}
-		\centering
-		\begin{tabular}{ r  r  r r  r r  r r  r r  r r}
-			resolution /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
-			\hline
-			5 & False &  4.7  &  (5.4) & 58.7 & (58.9) & 8.8 & (8.4) & 0.7 & (0.7) & 27.1 & (26.7)  \\
-
-			10 & False &  19.6 & (33.5) & 68.5 & (64.5)  & 1.0 & (0.8) & 0.8 & (0.3) & 10.1 & (0.9)\\
-
-			50 & False &  20.4 & (33.2) & 68.0 & (66.2)  & 0.9 & (0.1) & 0.7 & (0.0) & 10.1 & (0.5)\\
-
-			100 & False &  21.1 & (30.7) & 69.1 & (68.8)  & 1.1 & (0.0) & 0.7 & (0.0) & 7.9 & (0.4) \\
-
-			\hline
-
-			5 & True  &  18.9 & (28.5) & 67.3 & (66.4) & 1.3 & (1.6) & 1.0 & (0.5) & 11.5 & (3.0) \\	
-			10 & True &  18.9 & (33.7) & 66.6 & (63.4)  & 1.6 & (1.4) & 1.4 & (0.6) & 11.5 & (1.0)\\	
-			50 & True &  9.1 & (13.0) & 75.7 & (83.0) & 3.9 & (2.0) & 4.2 & (1.6) & 7.1 & (0.4) \\
-			100 & True &  7.0 & (10.1) & 73.8 & (81.9)  & 5.5 & (3.9) & 8.5 & (3.6) & 5.2 & (0.4) \\
-
-		\end{tabular}
-	\end{table*}
-
-	\todo{Check why values are so different for 5 m all touched False --> simplified line.}
-
-	\subsection{Faster Processing of the Cost Path Algorithm}
-	The original least cost path QGIS plug-in uses stops the least aggregation of the costs after the final end point has been found.
-	This early stopping might result in sub optimal paths around the end points.
-	For further calculation the search for the least cost paths has to be refactored and split the aggregation of the raster for the search of the path.
-	To be able to construct a full aggregated cost raster, aggregated cost path in further steps will be computed completely.\todo{really needed}
-	For the calculation of the least cost paths.
-	The finally least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer
-	The first step to optimize the calculation speed is, only the calculated the least cost paths for this smaller area. 
-	\subsubsection{Compare least cost paths, for overlay of both rasterizations}
-	\subsubsection{Restrict search to a minimum sized bounding box}
-	\subsubsection{Restrict search to a convex hull}
-	\subsubsection{Restrict search to reachable points}
-
-	\subsubsection{Compare the different Solutions}
-	\subsubsection{Check solution for least cost path between different points}
-
-	\section{Discussion}
-	\todo{discuss aggregation with max vs sum}
-	\todo{discuss other methods. Off algorithmic speed-ups}
-	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
-	\section{Related Works}
-	\section{Conclusion}
-
-\begin{acks}
-	\ldots
-\end{acks}
-
-%%
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{Web_Processing.bib}
-
-%%
-%% If your work has an appendix, this is the place to put it.
-\appendix
-
-\section{Research Methods}
-
-\subsection{Part One}
-
-\subsection{Part Two}
-
-\section{Online Resources}
-
-\end{document}
-\endinput
-%%
-%% End of file `sample-acmsmall.tex'.

### 2023-02-05 paper/paper-CableRoute.tex
added missing requirements, typos and types
linear regression of estimation time and total costs
refactor paper-CableRoute.tex to exclude the larger sections results and methods.
These were moved into there own files

--- 2023-01-11 paper/paper-CableRoute.tex
@@
 \newacronym{f}{f}{False}
-\newacronym{mmd}{mmd}{mean minimum distance}
-\newacronym{agg}{agg}{aggregated}
+\newacronym{agg.}{agg.}{aggregated}
 \newacronym{rel}{rel}{relative}
@@
 
-	\section{Methods}
-	We retrieve a set of different spacial data-sets from  public sources as a basis to create the cost raster.
-	Field of study are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
-	Areas protected by different European and National conservation laws are provided by the German Environment Agency
-	as \acrfull{wfs}~\cite{noauthor_schutzgebiete_2015}.
+	\section{Methods}\label{sec:methods}
+	\input{methods}
 
-	The nation wide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
-	and Geodesy\cite{noauthor_digitales_2021}.
-	The nation wide power grid (tags: 'power': line) has been retrieved via OpenStreetMap.\cite{boeing_osmnx_2017}
-	Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
-	Lower Saxony\cite{noauthor_opengeodatani_2022}.
-	In addition local planning geodata for the land usage are taken from
-	from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)\cite{noauthor_metropolplaner_2022}
+	\section{Results}\label{sec:results}
+	\input{results}
 
-	PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
-	The for the initial implementation of the least cost path algorithm the implementation for the QGIS-Plugin
-	'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} has been taken into account.
+	\section{Discussion}\label{sec:discussion}
+	As shown need of computation power increases with the power of two with the resolution.
+	While the error only scales linear with the resolution.
 
-	The different layers from the different entities are optionally filtered, buffered and than rasterized.
-	Filtering the layers of the files for special attributes enables to further differentiate further.
-	For examples makes it possible to differentiate between heath and uncultivated land.
-	Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
-	correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
- 	Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
- 	The costs of all layers of is aggregated with the maximum function.
-	Thus, an area covered by multiple layers is uniformly used with the highest costs.
-	For every area in the study area, that is not covered by any layer, is given the default cost.
-	The costs has been grouped into different levels~(see table~\ref{tab:1}) starting from preferential areas with
-	very low costs, via no restrictions, which is the default, used when no other layers are covering the local area,
-	to restricted, strongly restricted and prohibited areas with high costs.
-	These higher costs resemble the degree how much a local area should be avoided, while routing the path.
-	The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
-	the algorithm is willing to go, for avoiding an area of high costs.
-	Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum,
-	the weight that resembles the costs for these types of areas, has to be especially high. \\
-	The rasterization transforms a vector into a raster.
-	The rasterization can be executed in to different ways.
-	In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
-	given resolution and the new affine transformation and the coordinate reference system of the vector.
-	Both rasterization techniques differ  in who the pixels are selected, that describe, that a geo-object overlays the pixel.
-	The pixels can either be described as overlaid with a geo-object, when either the center of that pixel is overlaid
-	by the geo-object, or any part of the pixel is overlaid.
-	Hence, the version with any part overlaid is called all touched True.
-	The version where a overlaid pixel center is require is thus called all touched False.
-	All touched False is considered the default (see figure~\ref{fig:alltouched}).
-	\begin{figure}[!ht]
-		\centering
+	As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
+	This early stopping might result in sub optimal paths around the end points.
 
-		\resizebox{8cm}{!}{
-			\begin{tikzpicture}
-				\node[draw] at (0,6.5) {\huge all touched false};
-				\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
-				\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
+	As a any buffer around buildings are describes as protected areas, power poles are also included.
+	In further works, these should be excluded.
 
-				\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);
-
-				\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
-				\draw[fill=blue] (11,0)--(11,2)--(10,2)--(10,3)--(11,3)--(11,4)--(12,4)--(12,3)--(13,3)--(13,4)--(14,4)--(14,2)--(13,2)--(13,1)--(12,1)--(12,0)--cycle;
-			\end{tikzpicture}%
-		}%
-		\vspace*{5mm}
-		\resizebox{8cm}{!}{
-			\begin{tikzpicture}
-				\node[draw] at (0,6.5) {\huge all touched true};
-				\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
-				\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
-
-				\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);
-
-				\draw[step=1cm,gray,very thin] (9.1,-1.9) grid (16.9,5.9);
-				\draw[fill=blue] (10,0)--(10,5)--(11,5)--(11,4)--(15,4)--(15,1)--(13,1)--(13,0)--cycle;
-			\end{tikzpicture}%
-		}%
-		\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with weither all-touched False (above), or all touched True (below).}
-		\label{fig:alltouched}
-
-	\end{figure}
-
-	\begin{table}[h!]
-		\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}
-		\label{tab:1}
-		\centering
-		\begin{tabular}{ l  r  l }
-			Cost Level & Cost &  Example\\
-			\hline
-			Prohibited & 500					& National Parks, Buildings \\
-			strongly Restricted & 10 	& Bird Reserve \\
-			Restricted & 5	& Industrial Areas \\
-			No Restriction & 0.5					& Default\\
-			Preferential & 0.1					& Power Grid, Motorway Buffers\\
-		\end{tabular}
-	\end{table}
-	The completed list of layers and the processing applied to them, can be found in Supplement S1. \todo{create the Supplement from the processing rules}
-
-	\section{Results}
-	In this chapter we want to show the different cost raster, that were created from the very same set of layers,
-	but computed for different resolutions.
-	From this different raster the cost paths are calculated and compared.
-	In the last step the raster with lower resolution were used to calculate in a way, that they shall result in
-	similar paths as if the paths were computed from a high resolution raster.
-	\subsection{Cost Raster}
-	\subsection{Least Cost Paths}
-	For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
-	and all\_touched True.
-	The distance of both paths is calculated with the mean minimum distance.
-	For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
-	is computed and afterwards the minimum distances are averaged (see equation~\ref{eq:1}).
-	\begin{equation}
-		\label{eq:1}
-		d_{mean} = \frac{\sum_{i=1}^{n} d_{min}(p_i, L_2)}{n} \bigg\vert p_i \in L_1
-	\end{equation}
-
-	Hence, the mean minimum euclidean distance between the two paths can used to compute
-	the similarity.
-	As different table~\ref{tab:2} shows the distance between the two paths decreases
-	with increasing resolution.
-	In addition, this tendency is depicted in figure~\ref{fig:paths_resolution} for the calculated cost paths of 5m and 100 m resolutions.
-
-	At the same time the differences in the aggregated costs stay constant.
-	\todo{Interpretation} When normalize the aggregated costs by the resolution.
-	Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
-	linear with the resolution.
-	On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
-	the resolution.
-	Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}
-
-	\begin{figure}
-		\centering
-
-		\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m.pdf} }}%
-		\qquad
-		\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
-		\caption{Figures of the least cost paths. Contrasting the paths for different resolutions. Using OpenStreetMaps as base map.}
-		\label{fig:paths_resolution}
-	\end{figure}
-
-	\begin{figure}
-	\centering
-
-	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F.pdf} }}%
-	\qquad
-	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T.pdf} }}%
-
-	\caption{Figures of the least cost paths. Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched. Using OpenStreetMaps as base map.}
-	\label{fig:paths_alltouched}
-\end{figure}
-
-	\begin{table*}[t]
-	\caption{Least cost paths as length for the different \acrfull{res} of the rasters, including the \acrfull{mmd} and the \acrfull{agg} costs.
-From the \acrshort{agg} costs the differences of the \acrshort{agg} costs and the \acrfull{corr} \acrshort{agg} by resolution are given.}
-	\label{tab:2}
-	\centering
-	\begin{tabular}{ r  r  r  r  r  r  r  r  r}
-		\acrshort{res} /m & $length_{\acrshort{al}=\acrshort{f}} /m$ & $length_{\acrshort{al}=\acrshort{t}} /m$ & \acrshort{mmd} /m & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg} $costs_{\acrshort{al}=\acrshort{t}} $ \\
-		\hline
-		5 & 76136.27	& 78002.00 & 126.04  & 18665.923 & 19616.756 & -850 & 93329.6 &  97584.77 \\
-		10 & 75430.10 	& 77936.57 & 277.92 & 8931.245 & 9731.175 & -799.95 & 89312.45 & 97311.75 \\
-		50 & 76135.02	& 70619,95 & 1140.01 & 1409.023 & 2300.073 & -891.05 & 70451.15 & 115003.65 \\
-		100 & 76283.80	& 74120.73 & 1946.41 & 640.516 & 1572.268 & -931.7 & 64051.6 & 167226.8 \\
-
-	\end{tabular}
-	\end{table*}
-
-	Comparing the all\_touched True rasterization and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterization at the
-	different resolutions the later paths are more similar.
-	The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m. \todo{Option add 2.5 m and 25 m}
-
-	The similarity of all all\_touched False runs is higher than, the similar between the all\_touched False and all\_touched True runs with the same resolution, except for the highest resolution.
-
-	Hence, in an overall perspective paths of the all\_touched False runs stay in a
-	similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
-	This behaviour is depicted in figure~\ref{fig:paths_alltouched}.
-	On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
-	But the extent is smaller.
-	The length of the the paths only differs to a maximum of about 10 \%.
-	The least path distance for higher resolutions can be lower, because new paths, between regions that are forbidden
-	in higher resolution can open.
-	On the other side the length of the paths can increase, because with higher resolution more vertices will be used.
-
-	The zonal stat (see table~\ref{table:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
-	percentage of which costs levels are around the path.
-	When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
-	Preferential Level and less the NoRestriction Level.\todo{Interpretation:}
-	The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
-	\todo{Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100 m! No problem because buffering is included in the cost raster.}
-	There is no strong tendency for the all\_touched False least cost paths.
-
-	\todo{used regions}
-
-	\setlength{\tabcolsep}{10pt}
-
-	\begin{table*}[t]
-		\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
-		\label{table:3}
-		\centering
-		\begin{tabular}{ r  r  r r  r r  r r  r r  r r}
-			resolution /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
-			\hline
-			5 & False &  4.7  &  (5.4) & 58.7 & (58.9) & 8.8 & (8.4) & 0.7 & (0.7) & 27.1 & (26.7)  \\
-
-			10 & False &  19.6 & (33.5) & 68.5 & (64.5)  & 1.0 & (0.8) & 0.8 & (0.3) & 10.1 & (0.9)\\
-
-			50 & False &  20.4 & (33.2) & 68.0 & (66.2)  & 0.9 & (0.1) & 0.7 & (0.0) & 10.1 & (0.5)\\
-
-			100 & False &  21.1 & (30.7) & 69.1 & (68.8)  & 1.1 & (0.0) & 0.7 & (0.0) & 7.9 & (0.4) \\
-
-			\hline
-
-			5 & True  &  18.9 & (28.5) & 67.3 & (66.4) & 1.3 & (1.6) & 1.0 & (0.5) & 11.5 & (3.0) \\	
-			10 & True &  18.9 & (33.7) & 66.6 & (63.4)  & 1.6 & (1.4) & 1.4 & (0.6) & 11.5 & (1.0)\\	
-			50 & True &  9.1 & (13.0) & 75.7 & (83.0) & 3.9 & (2.0) & 4.2 & (1.6) & 7.1 & (0.4) \\
-			100 & True &  7.0 & (10.1) & 73.8 & (81.9)  & 5.5 & (3.9) & 8.5 & (3.6) & 5.2 & (0.4) \\
-
-		\end{tabular}
-	\end{table*}
-
-	\todo{Check why values are so different for 5 m all touched False --> simplified line.}
-
-	\subsection{Faster Processing of the Cost Path Algorithm}
-	The original least cost path QGIS plug-in uses stops the least aggregation of the costs after the final end point has been found.
-	This early stopping might result in sub optimal paths around the end points.
-	For further calculation the search for the least cost paths has to be refactored and split the aggregation of the raster for the search of the path.
-	To be able to construct a full aggregated cost raster, aggregated cost path in further steps will be computed completely.\todo{really needed}
-	For the calculation of the least cost paths.
-	The finally least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer
-	The first step to optimize the calculation speed is, only the calculated the least cost paths for this smaller area. 
-	\subsubsection{Compare least cost paths, for overlay of both rasterizations}
-	\subsubsection{Restrict search to a minimum sized bounding box}
-	\subsubsection{Restrict search to a convex hull}
-	\subsubsection{Restrict search to reachable points}
-
-	\subsubsection{Compare the different Solutions}
-	\subsubsection{Check solution for least cost path between different points}
-
-	\section{Discussion}
 	\todo{discuss aggregation with max vs sum}
@@
 	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
-	\section{Related Works}
-	\section{Conclusion}
+	\section{Related Works}\label{sec:related-works}
+
+	\section{Conclusion}\label{sec:conclusion}
 
 \begin{acks}
-	\ldots
+	$\ldots$
 \end{acks}

### 2023-02-05 paper/results.tex
refactor paper-CableRoute.tex to exclude the larger sections results and methods.
These were moved into there own files

% !TeX spellcheck = en_US

In this chapter we want to show the different cost raster, that were created from the very same set of layers,
but computed for different resolutions.
From this different raster the cost paths are calculated and compared.
In the last step the raster with lower resolution were used to calculate in a way, that they shall result in
similar paths as if the paths were computed from a high resolution raster.
\subsection{Cost Raster}\label{subsec:cost-raster}

The cost raster contains all the costs as weights for the geographic region of the study area.
The regions outside the study area are given, a no-data value and will not be use for the calculation of the least cost path.
It is decided to use the maximum value, for any place in the raster, that is covered by several rules at the same time.
When the resolution that is use for the rasterization is smaller than the object size, than the effect of the rasterization with all touched True or False is limited.
For all touched True any part of the pixel that is covered by the object, will attribute that whole pixel to the object.
Thus, the object appears to be halve a pixel size larger in all directs.
As can be seen in figure~\ref{fig:costs_5m} that shows a detailed view of the costs for village of Beverstedt.
Due to the maximum aggregation of the costs, the average cost of the raster of all touched true will be over estimated.
All touched false will be a better description of the real size of the object.
\begin{figure}
	\centering

	\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alT.png} }}%
	\qquad
	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alF.png} }}%
	\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 5~m.}
	\label{fig:costs_5m}
\end{figure}

In contrast when the resolution the is much less, than the size of the object the described behavior changes.
On one hand, while the area, of the pixel is increased for all touched True, also the area for which the cost is overestimated increases.
On the other hand while the pixel size for all touched  False increases, not only the over estimated area increases for that pixel as describes for all touched true, but in addition the object is less probable situated in the center of the pixel.
The consequence of the pixel is, that will decreasing resolution that object is not included in the rasterziation.
Hence, hence for all touched false lower resolution  leads ot a loss of information.
Because the default is relative small compared the over effect is a underestimation of the costs.
The figure~\ref{fig:costs_100m} shows for the resolution of 100 m, that while larger are still included in the map they appear to be much larger.
On the contrary smaller objects might not be included.
Objects that are small on in one dimension as streets will be included in a stochastic manner.
Described by the likely of that a object of that sized overlaps with the pixel center.
\begin{figure}
	\centering

	\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alT.png} }}%
	\qquad
	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alF.png} }}%
	\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 100~m.}
	\label{fig:costs_100m}
\end{figure}
Thus, with having larger areas and more objects, with higher costs, all touched True rasterization might more likely lead to longer roots and more likely in blocking the  direct spatial path.

\subsection{Least Cost Paths}\label{subsec:least-cost-paths}
For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
and all\_touched True.
The distance of both paths is calculated with the mean minimum distance.
For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
is computed and afterwards the minimum distances are averaged (see equation~\ref{eq:1}).
\begin{equation}
	\label{eq:1}
	d_{mean} = \frac{1}{|L_1|} \sum_{i=1}^{n} d_{min}(p_i, L_2) \bigg\vert p_i \in L_1
\end{equation}

\begin{equation}
\label{eq:2}
d_{max} = max(\sum_{i=1}^{n} d_{min}(p_i, L_2)) \bigg\vert p_i \in L_1
\end{equation}

Hence, the mean minimum euclidean distance between the two paths can used to compute
the similarity.
As different table~\ref{tab:2} shows the distance between the two paths decreases
with increasing resolution.
In addition, this tendency is depicted in figure~\ref{fig:paths_resolution} for the calculated cost paths of 5m and 100 m resolutions.

At the same time the differences in the aggregated costs stay constant.
\todo{Interpretation} When normalize the aggregated costs by the resolution.
Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
linear with the resolution.
On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
the resolution.
Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}

\begin{figure}
	\centering

	\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m.pdf} }}%
	\qquad
	\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
	\caption{Figures of the least cost paths. Contrasting the paths for different resolutions. Paths computed from all touched false raster are indicated by dashed lines. Results from all touched True are indicuated by continuous lines. Higher resolutions are indicated by the color green, lower resolutions by the color red. Using OpenStreetMaps as base map.}
	\label{fig:paths_resolution}
\end{figure}

\begin{figure}
	\centering

	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F.pdf} }}%
	\qquad
	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T.pdf} }}%

	\caption{Figures of the least cost paths. Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched. Paths computed from all touched false raster are indicated by dashed lines. Results from all touched True are indicuated by continuous lines. Higher resolutions are indicated by the color green, lower resolutions by the color red. Using OpenStreetMaps as base map.}
	\label{fig:paths_alltouched}
\end{figure}

\begin{table*}[t]
	\caption{Least cost paths as length for the different \acrfull{res} of the raster, including the mean minimum distance and the maximum minimum distance and the \acrfull{agg.} costs. From the \acrshort{agg.} costs the differences of the \acrshort{agg.} costs and the \acrfull{corr} \acrshort{agg.} by resolution are given.} 
	\label{tab:2}
	\centering
	\begin{tabular}{ r  r  r  r  r  r  r  r  r  r}
		\acrshort{res} /m & $length_{\acrshort{al}=\acrshort{f}} /m$ & $length_{\acrshort{al}=\acrshort{t}} /m$ & $d_{mean}$ /m & $d_{max}/m$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{t}} $ \\
		\hline
		5 	& 76136.27	& 78002.00 &  126.04 & 1065.00 & 18665.923 & 19616.756 & -850.00 & 93329.60 &  97584.77 \\
		10 	& 75430.10 	& 77936.57 &  277.92 & 1590.00 &  8931.245 &  9731.175 & -799.95 & 89312.45 &  97311.75 \\
		25 	& 75422.85 	& 78422.85 &  313.75 & 1621.15 &  3354.869 &  3872.656 & -517.78 & 83871.73 &  96816.40 \\
		50 	& 76135.02	& 70619,95 & 1140.01 & 4950.00 &  1409.023 &  2300.073 & -891.05 & 70451.15 & 115003.65 \\
		100 & 76283.80	& 74120.73 & 1946.41 & 6016.64 &   640.516 &  1572.268 & -931.70 & 64051.60 & 167226.80 \\

	\end{tabular}
\end{table*}

Comparing the all\_touched True rasterization and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterization at the
different resolutions the later paths are more similar.
The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m.

The similarity of all all\_touched False runs is higher than, the similar between the all\_touched False and all\_touched True runs with the same resolution, except for the highest resolution.

Hence, in an overall perspective paths of the all\_touched False runs stay in a
similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
This behaviour is depicted in figure~\ref{fig:paths_alltouched}.
On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
But the extent is smaller.
The length of the the paths only differs to a maximum of about 10 \%.
The least path distance for higher resolutions can be lower, because new paths, between regions that are forbidden
in higher resolution can open.
On the other side the length of the paths can increase, because with higher resolution more vertices will be used.

The zonal stat (see table~\ref{tab:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
percentage of which costs levels are around the path.
When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
Preferential Level and less the NoRestriction Level.\todo{Interpretation:}
The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
\todo{Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100 m! No problem because buffering is included in the cost raster.}
There is no strong tendency for the all\_touched False least cost paths.

\subsection{Execution time}\label{subsec:execution-time}
In theory, the execution time should increase with the square of the resolution, because higher resolutions result in a higher number of pixels and thus data points the aggregated costs needs to be calculated for. 
A full logarithmic fit for several repetitions of the execution shows, that the execution time scales with power of $2.1997  \pm 0.007$ of the inverse resolution. 
With the caveat of a low number of samples this can equally be success fully be fitted to a second degree polynomial of the inverse resolution with a $r^2$ for the test set of 0.99 and a the squared inverse of the resolution with a $r^2$ for the test set of 0.99.
Hence, the order of magnitude

The total execution time consists of two parts. 
The aggregation of the costs and the back tracking of the least cost to find the path.
\todo{Find out whether back tacking scales linear or square.}

\todo{used regions}

\setlength{\tabcolsep}{10pt}

\begin{table*}[t]
	\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
	\label{tab:3}
	\centering
	\begin{tabular}{ r  r  r r  r r  r r  r r  r r}
		resolution /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
		\hline
		5 & False &  4.7  &  (5.4) & 58.7 & (58.9) & 8.8 & (8.4) & 0.7 & (0.7) & 27.1 & (26.7)  \\
		10 & False &  19.6 & (33.5) & 68.5 & (64.5)  & 1.0 & (0.8) & 0.8 & (0.3) & 10.1 & (0.9)\\
		25 & False &  19.2 & (34.2) & 68.9 & (64.9)  & 1.0 & (0.2) & 0.7 & (0.1) & 9.7 & (0.6)\\
		50 & False &  20.4 & (33.2) & 68.0 & (66.2)  & 0.9 & (0.1) & 0.7 & (0.0) & 10.1 & (0.5)\\
		100 & False &  21.1 & (30.7) & 69.1 & (68.8)  & 1.1 & (0.0) & 0.7 & (0.0) & 7.9 & (0.4) \\

		\hline

		5 & True  &  18.9 & (28.5) & 67.3 & (66.4) & 1.3 & (1.6) & 1.0 & (0.5) & 11.5 & (3.0) \\	
		10 & True &  18.9 & (33.7) & 66.6 & (63.4)  & 1.6 & (1.4) & 1.4 & (0.6) & 11.5 & (1.0)\\	
		25 & True &  18.7 & (31.9) & 65.5 & (65.5)  & 2.0 & (1.3) & 2.5 & (0.7) & 11.4 & (0.6)\\
		50 & True &  9.1 & (13.0) & 75.7 & (83.0) & 3.9 & (2.0) & 4.2 & (1.6) & 7.1 & (0.4) \\
		100 & True &  7.0 & (10.1) & 73.8 & (81.9)  & 5.5 & (3.9) & 8.5 & (3.6) & 5.2 & (0.4) \\	
	\end{tabular}
\end{table*}

\subsection{Faster Processing of the Cost Path Algorithm}\label{subsec:faster-processing-of-the-cost-path-algorithm}

The final least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer to the paths resulted by the all touched false rasterization.
The first step is to optimize the calculation speed is, only to calculated the least cost paths for this smaller area.
Another method, is to improve the prediction of the medium resolution itself and thus reduce the need for a computation in higher resolution.
\subsubsection{Compare least cost paths, for overlay of both rasterizations}
As all touched true rasterization overestimates and all touched true underestimates the real costs.
A weighted mean will describe the real costs more precise.
As present work indicates, that the weight should favor the all touched false raster.
The best weight should be the percentage of the pixel, which is covered by the object, but this can not be computed in this work.
Thus, the best weight has to be searched.
An alternative might be to compute the cost raster in higher resolution, but than to reproject the to a smaller resolution with a using a (linear) interpolation of the of the weights.\todo{test this idea}
\subsubsection{Compare least cost paths, for down sampled cost paths}
As an alternative to the super position of the all touched true and false raster for the same resolution, the all touched false raster was down sampled to 10~m, 25~m, 50~m and 100~m with bi-linear interpolation.
With this method smaller structures still can be seen in the cost raster, although the resolution is reduced.

% \subsubsection{Restrict search to a minimum sized bounding box}
\subsubsection{Restrict search to a convex hull}
\todo{Construct a polygon form the two lines. Buffer the polygon.}
\todo{max minimum path distance}
\todo{crossing points of two paths}
\subsubsection{Restrict search to reachable points}
For this purpose the aggregated cost as to converted back into a raster.

\subsubsection{Compare the different Solutions}
\subsubsection{Check solution for least cost path between different points}

### 2023-02-09 paper/methods.tex
Added Docker/Docker-Compose files. Not working with the jupyter notebooks.
In general fixed spelling.

added to latex:
 - description of starting end ending points,
 - first bits on the aggregation algorithm
 - figs of cost raster -> aggregated costs -> least cost path
 - a first part on the restricted search space by a buffer

@@
 The completed list of layers and the processing applied to them, can be found in Supplement S1. \todo{create the Supplement from the processing rules}
+
+The whole stack of the cost raster, via the aggregated costs to the resulting least cost path can is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).
+Because of the early stopping strategy used in the aggregation sub-paths with high costs have not been fully computed.Instead the execution stops early after reaching the end point.
+Because the end point ends at a power transformer, which is a building type, the paths ends at in a Prohibited Area.
+Therefore areas even farther away from the starting point have been explored first.
+
+\begin{figure}
+	\centering
+
+	\subfloat[\centering Cost Raster.]{{\includegraphics[width=.25\linewidth]{./images/CostRasterExample.png} }}%
+	\qquad
+	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.25\linewidth]{./images/AggregatedCosts.png} }}%
+	\qquad
+	\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.25\linewidth]{./images/LeastCostPathExample.png} }}%
+
+	\caption{Figures of Cost raster and the resulting aggregated Costs and the least cost path for a resolution of 50~m.}
+	\label{fig:costs2path}
+\end{figure}

### 2023-02-09 paper/paper-CableRoute.tex
Added Docker/Docker-Compose files. Not working with the jupyter notebooks.
In general fixed spelling.

added to latex:
 - description of starting end ending points,
 - first bits on the aggregation algorithm
 - figs of cost raster -> aggregated costs -> least cost path
 - a first part on the restricted search space by a buffer

@@
 
+	The aggregation is implemented in a priority queue. 
+	Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.
+	The aggregation is implemented with early stopping. 
+	After the last end point has been has got an aggregated cost the aggregation stops and the back tracking starts.
+
 	\section{Methods}\label{sec:methods}
@@
 
-	\todo{discuss aggregation with max vs sum}
-	\todo{Scaling with multiple endpoints.}
-	\todo{discuss other methods. Off algorithmic speed-ups}
-	\todo{discuss other technical methods of speed-ups, as precompilation, JIT, vector units, multi-core}
+	Difference aggregation of the rule of a cost raster with max vs sum
+
+	Scaling with multiple endpoints.
+	Not algorithmic speed-ups as precompilation, JIT, vector units, multi-core
 	\section{Related Works}\label{sec:related-works}

### 2023-02-09 paper/results.tex
Added Docker/Docker-Compose files. Not working with the jupyter notebooks.
In general fixed spelling.

added to latex:
 - description of starting end ending points,
 - first bits on the aggregation algorithm
 - figs of cost raster -> aggregated costs -> least cost path
 - a first part on the restricted search space by a buffer

@@
 and all\_touched True.
+
+The hypothetical path starts at a transformer about 6~km north the container terminal Bremerhaven and then end at a transformer in the south east of the Osterholz county.
+
 The distance of both paths is calculated with the mean minimum distance.
@@
 with increasing resolution.
-In addition, this tendency is depicted in figure~\ref{fig:paths_resolution} for the calculated cost paths of 5m and 100 m resolutions.
+In addition, this tendency is depicted in figure~\ref{fig:paths_resolution} for the calculated cost paths of 5~m and 100~m resolutions.
+
+In contrast the largest distance between the paths (see equation~\ref{eq:2}) will be used, to estimate the minimum distance to the areas that still should be examined to grantee a, that at least cost path found in this limited space is still likely to be optimal.
 
@@
 The aggregation of the costs and the back tracking of the least cost to find the path.
-\todo{Find out whether back tacking scales linear or square.}
+\todo{back tacking scales with number of points.}
 
@@
 Thus, the best weight has to be searched.
-An alternative might be to compute the cost raster in higher resolution, but than to reproject the to a smaller resolution with a using a (linear) interpolation of the of the weights.\todo{test this idea}
+An alternative might be to compute the cost raster in higher resolution, but than to reproject the to a smaller resolution with a using a (linear) interpolation of the of the weights.
+
+While the aggregated thus can be speed up.
+The time needed for the back tracking stays unchanged.
+
 \subsubsection{Compare least cost paths, for down sampled cost paths}
@@
 % \subsubsection{Restrict search to a minimum sized bounding box}
-\subsubsection{Restrict search to a convex hull}
-\todo{Construct a polygon form the two lines. Buffer the polygon.}
-\todo{max minimum path distance}
-\todo{crossing points of two paths}
+\subsubsection{Restrict search to a buffered around the least cost paths}
+Construct a polygon form the two lines.
+Buffer the polygon with twice the max minimum path distance.
+
+This enabled the possibility  to run a 2.5~m resolution cost raster and clip it to the extent of the polygon. 
+This clipped 2.5~m raster for all touched true did only slightly change the path. 
+While the all touched False raster, leads to a totally new segment at the end of the path. 
+Due to the small resolution, that a small path with the extent of about 2.5 m width became passable.
+This small path is a power line next to road between which build a cone in a protected landscape area.
+The street and the protected landscape area are both restricted areas, while the power line is preferential.
+This showing, that the way the cost raster is created in the first place can play a vital role, in the end result. So that a nuance, can cause a detour.
+When this behavior occurs, the polygon can not grantee to include the least cost path.
+Therefore this polygon should be overlayed with a polygon around the shortest path.
+
 \subsubsection{Restrict search to reachable points}

### 2023-02-09 src/least_cost_path/find_least_cost_path.py

@@ -172,7 +172,8 @@ def find_least_cost_path(cost_raster: DataArray, cost_raster_band: int, is_neare
 
     _path_features = []
     for itr, (_map, _indexes,  _end_node) in enumerate(dijkstra(start_tuples[0], end_tuples, matrix, is_nearest)):
-        (path, costs), terminal_tuple = *backtrack_path(_map, _indexes, _end_node, start_tuples[0]), end_tuples[itr]
+        (path, costs) = backtrack_path(_map, _indexes, _end_node, start_tuples[0])
+        terminal_tuple = end_tuples[itr]
         if aggregated_save:
             aggregated_raster = matrix2raster(_map, raster_2d, -9999)
             write_compressed(aggregated_raster, Path(aggregated_save))

### 2023-02-19 paper/discussion.tex
added Supplement S1
paper:
results.tex:
added part for the sub-sections:
 1) for overlay of both rasterizations
 2) for down sampled cost paths

refactor: added discussion.tex
added citations, parts for the Introduction

update format of least_cost_path_evaluation.ipynb

As shown, the need for computation time increases with the power of two with the resolution. 
Similarly, the usage of main memory increases. 
Which in tern limits the processable data points and resolution.
On the error only scales linear with the resolution. 
Hence, there is a diminishing return of smaller errors, compared to compute time and resources used.

Therefore, this worked tries a) to reduce the needed compute power and b) decrease the deviation for a given resolution.

For b), two methods have been used to increase the similarity of the paths computed from medium resolution raster, to the path of the highest resolution raster.
These methods can be used as surrogates for the more complex calculation of the least cost path with the higher resolution.
In the first method a bi-linear down sampling of the higher resolution raster has been applied and in the second method, the all touched false and true raster where averaged in different ratios, to computed the optimum mixed raster, for the used costs.
While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the down sampling method is more
simple and does not need to be optimized for the given costs.
Hence, it is easier and widely applicable.

For a), to reduce the computational complexity, the calculation of the medium resolution raster has been applied, to reduce the search space of the aggregation.
While this method reduces the calculation time and usage of main memory, the backtracking part stays unchanged.

As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
This early stopping might result in sub-optimal paths around the end points for some edge cases, where the connection via another neighbor might be more optimal.

The set of rules that are used create the cost raster, includes a rule for creating a buffer around buildings  which is set to the level \textbf{prohibited} areas.
While the used data set for buildings, also includes power poles, the area of the level \textbf{preferential} power grid, includes isles of high costs. 
In further works, these should be excluded.

In this work, the effect of computation costs and deviation of the results, is examined for a very limited set of points.
Also, only the costs of finding the least cost path from one starting point, to a single end point has been considered.
When, multiple endpoints are used, the computation cost for the aggregated cost raster has only to computed once.
On the other hand the effect early stopping is not as dramatic, because the algorithm would only stop early for the last end point with the highest cost.
When calculating several paths from one raster, the benefits of chapter~\ref{subsec:faster-processing-of-the-cost-path-algorithm} is reduced.
Especially, the pre-calculation on lower resolution raster and buffering of the resulting paths as a restricted search area decreases in it effectivity.
As the number of paths rises, fewer of points in higher resolution cost raster will be deselected for the final estimation of the least cost path.

In this work, the single cost raster layers aggregated with the maximum function. 
Another possible aggregation function is the sum.
Each aggregation function can be justified, by a different interpretation of the costs and its scale.
The reasoning of selection the maximum function is, limiting the maximum costs.
When the \textbf{prohibited} level is used as the highest level.
Than higher levels are prohibited within the concept.
This an area even more prohibited, when it is included in two different rules of the level?
On the other side, This aggregation is unable to differentiate between these nuances, as  the sum function is.
This way, one can differential between, different sub-levels.
One consequence is, that a small mediation between high and low single cost raster, in the final cost raster is possible.

The fact, that all touched False raster lead to more similar results compared to high resolution raster, is due to the fact, that the default level is relatively low.
In cause the default level would rise, the effect probably would flip for low resolution raster.
For high resolution raster, this effect would still hold true, because the fact, that the raster pixel center is used for sampling, reflects the original geometry better.

Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100~m!
No problem because buffering is included in the cost raster.
Not algorithmic speed-ups as pre-compilation, JIT, vector units, multi-core

### 2023-02-19 paper/methods.tex
added Supplement S1
paper:
results.tex:
added part for the sub-sections:
 1) for overlay of both rasterizations
 2) for down sampled cost paths

refactor: added discussion.tex
added citations, parts for the Introduction

update format of least_cost_path_evaluation.ipynb

--- 2023-02-05 paper/methods.tex
@@
 \end{table}
-The completed list of layers and the processing applied to them, can be found in Supplement S1. \todo{create the Supplement from the processing rules}
+The completed list of layers and the processing applied to them, can be found in \textbf{Supplement S1}.
+
+The whole stack of the cost raster, via the aggregated costs to the resulting least cost path can is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).
+Because of the early stopping strategy used in the aggregation sub-paths with high costs have not been fully computed.
+Instead, the execution stops early after reaching the end point.
+Because the end point ends at a power transformer, which is a building type, the paths ends at in a Prohibited Area.
+Therefore, areas even farther away from the starting point have been explored first.
+
+\begin{figure}
+	\centering
+
+	\subfloat[\centering Cost Raster.]{{\includegraphics[width=.45\linewidth]{./images/CostRasterExample.png} }}%
+	\qquad
+	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.45\linewidth]{./images/AggregatedCosts.png} }}%
+	\qquad
+	\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPathExample.png} }}%
+
+	\caption{Figures of Cost raster and the resulting aggregated Costs and the least cost path for a resolution of 50~m.}
+	\label{fig:costs2path}
+\end{figure}

### 2023-02-19 paper/paper-CableRoute.tex
added Supplement S1
paper:
results.tex:
added part for the sub-sections:
 1) for overlay of both rasterizations
 2) for down sampled cost paths

refactor: added discussion.tex
added citations, parts for the Introduction

update format of least_cost_path_evaluation.ipynb

@@
 	Sometimes, finding the shortest path is not sufficient.
-	Additional parameters play also have to be taken into consideration.
+	Additional parameters also have to be taken into consideration.
 	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline
-	\todo{source}.
+	\cite{suleiman_optimal_2015}.
 	For planing the additional routes for a power grid, additional aspects as legal regulations and acceptance
@@
 	Also technical aspects, as the effects on the grid stability might be further points to consider.\cite{schafer_understanding_2022}
-	\todo{what is the least cost path}
 
+	Due to the increasing need for renewable energy wind off shore wind turbines are source which currently supplies 5.5~\% annual percentage in 2020 of the German energy mix.~\cite{noauthor_nettostromerzeugung_2021}
+	For wind energy in general highest efficiency can only be archived with high wind speeds. Highest wind speeds are reached along the cost lines in the north of Germany.
+	In contrast, most consumers, as industry are situated in south Germany.
+	Source and consumer have to be linked via new power lines.
+
+	Other methods as the multi-criteria decision analysis take further aspects and stakeholder as experts, public and decision maker and into consideration.\cite{bertsch_participatory_2016}
+
+	For planning a possible route for a power line the least cost path algorithm is applied.
+	The least cost path algorithm is a Dijkstra algorithm applied on a raster map.
+	The vertices of the graph are the centers on neighboring pixels in al possible eight directions.
+	The weights of the edges are the local costs to travel from one pixel to the neighboring pixel.
+	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
+	The least cost path algorithm consists of at least two sequential steps.
+1) The first steps is to aggregate the costs of traveling from the starting point to a given set of end points.
+This steps generates the aggregated cost raster of the total cost that are needed to travel to a point, from the cost raster.
+	2) In the next step the back tracking, the route of the actual least cost path is calculated.
+For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.
+
+	\todo{Were to put the following part?}
 	The aggregation is implemented in a priority queue. 
@@
 	\section{Methods}\label{sec:methods}
+
 	\input{methods}
@@
 	\section{Results}\label{sec:results}
+
 	\input{results}
@@
 	\section{Discussion}\label{sec:discussion}
-	As shown need of computation power increases with the power of two with the resolution.
-	While the error only scales linear with the resolution.
+	\input{discussion}
 
-	As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
-	This early stopping might result in sub optimal paths around the end points.
-
-	As a any buffer around buildings are describes as protected areas, power poles are also included.
-	In further works, these should be excluded.
-
-	Difference aggregation of the rule of a cost raster with max vs sum
-
-	Scaling with multiple endpoints.
-	Not algorithmic speed-ups as precompilation, JIT, vector units, multi-core
 	\section{Related Works}\label{sec:related-works}
@@
 	\section{Conclusion}\label{sec:conclusion}
+	When using the researched strategies, the on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
+	\todo{Befragung: local population, relative weights per district?}
 

### 2023-02-19 paper/results.tex
added Supplement S1
paper:
results.tex:
added part for the sub-sections:
 1) for overlay of both rasterizations
 2) for down sampled cost paths

refactor: added discussion.tex
added citations, parts for the Introduction

update format of least_cost_path_evaluation.ipynb

@@
 similar paths as if the paths were computed from a high resolution raster.
+
 \subsection{Cost Raster}\label{subsec:cost-raster}
@@
 \end{equation}
+This equation is used, to measure the extent of similarity between the paths.
 
@@
 the resolution.
-Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.\todo{show map}
+Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.
 
@@
 	\begin{tabular}{ r  r  r  r  r  r  r  r  r  r}
-		\acrshort{res} /m & $length_{\acrshort{al}=\acrshort{f}} /m$ & $length_{\acrshort{al}=\acrshort{t}} /m$ & $d_{mean}$ /m & $d_{max}/m$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{t}} $ \\
+		\acrshort{res} /m & $l_{\acrshort{al}=\acrshort{f}} /m$ & $l_{\acrshort{al}=\acrshort{t}} /m$ & $d_{mean}$ /m & $d_{max}/m$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{t}} $ \\
 		\hline
@@
 The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
-\todo{Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100 m! No problem because buffering is included in the cost raster.}
+
 There is no strong tendency for the all\_touched False least cost paths.
@@
 \subsection{Execution time}\label{subsec:execution-time}
+
 In theory, the execution time should increase with the square of the resolution, because higher resolutions result in a higher number of pixels and thus data points the aggregated costs needs to be calculated for.
@@
 The aggregation of the costs and the back tracking of the least cost to find the path.
-\todo{back tacking scales with number of points.}
-
-\todo{used regions}
+% \todo{back tacking scales with number of points.}
 
@@
 	\begin{tabular}{ r  r  r r  r r  r r  r r  r r}
-		resolution /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
+		res /m & all touched & \multicolumn{2}{c}{ $ r_{Preferential} \% $}  & \multicolumn{2}{c}{ $ r_{No Restriction} \% $ }  & \multicolumn{2}{c}{ $ r_{Restricted} \% $}  & \multicolumn{2}{ c }{ $ r_{strongly Restricted}\% $ } & \multicolumn{2}{c}{ $ r_{Prohibited} \% $ } \\
 		\hline
@@
 \subsubsection{Compare least cost paths, for overlay of both rasterizations}
+
 As all touched true rasterization overestimates and all touched true underestimates the real costs.
@@
 
+The optimal ratio of overlaying all touched false and all touched true cost raster for 10~m resolution is estimated via the resulting least cost path. 
+The mean distance of least cost paths for different ratio has been estimated to the path from the all touched false raster of the next higher (5~m) resolution. 
+Table~\ref{tab:4} shows, that distance decreases with the higher ratio from 1:1, 2:1, 4:1 and than increases with higher ratio 8:1, 16:1 and so on.
+In addition, the distances to the original paths from the all touched false and all touched true raster change with the applied ratio of the overlayed raster.
+So that, a lower ratio increases the similarity to the path of the all touched true raster, the similarity to the paths from the all touched false raster increases with the ratio.
+
+\begin{table}[h!]
+	\caption{Length of the path computed from the overlaying of all touched false and true raster and the mean distance of the paths to the paths calculated from the all touched false 5~m resolution  and all touched false and true raster of 10~m resolution.}
+	\label{tab:4}
+	\centering
+	\begin{tabular}{ r  r  r  r}
+		r & $d_{5~al=f}$ /m &  $d_{10~al=f}$ /m & $d_{10~al=t}$ /m \\
+		\hline
+
+		  1:1  &    119.6 &  285.50 &  47.21\\
+		  2:1  &    97.11 &  263.51 &  74.19\\
+		  4:1  &    40.13 &  206.38 & 100.16\\
+		  8:1  &    41.73 &  169.02 & 137.34\\
+		 16:1  &    56.69 &  153.32 & 152.71\\
+		 32:1  &    56.69 &  145.56 & 162.09\\
+		 64:1  &   163.48 &   10.61 & 272.36\\
+	  % 128:1  &   163.82 &   10.29 & 276.71\\
+
+	\end{tabular}
+\end{table}
+
 \subsubsection{Compare least cost paths, for down sampled cost paths}
+
 As an alternative to the super position of the all touched true and false raster for the same resolution, the all touched false raster was down sampled to 10~m, 25~m, 50~m and 100~m with bi-linear interpolation.
 With this method smaller structures still can be seen in the cost raster, although the resolution is reduced.
+The distances of the paths that are computed from the bi-linear down sampled  5~m resolution raster (all touched false) is from table~\ref{tab:5} shows, that a low distance to path of the 5~m resolution raster was only obtained when down sampling to a resolution not to low.
+Only when down sampling to a resolution of 10~m, is the computed path considerable close to the path from the higher resolution raster.
+Is holds true, but to a much reduced extend to the path computed from the raster, that was down sampled to a resolution of 25~m.
+The opposite is true for the lower resolution raster which is more similar to paths computed from the all touched true cost raster.
+For every path from a down sampled raster is more similar to a path computed from an all touched true raster, than an all touched false raster, although the all touched false raster of the 5 m resolution was used for down sampling.
+
+\begin{table}[h!]
+	\caption{Length of the path computed from the bi-linear down sampled raster and the mean distance of the paths from the down sampled raster to the paths calculated from the all touched true and all touched false raster of the same resolution as the down sampled raster.}
+	\label{tab:5}
+	\centering
+	\begin{tabular}{ r  r  r  r  r}
+		res /m & l /m  & $d_{5~m}$ /m &  $d_{al=f}$ /m & $d_{al=t}$ /m \\
+		\hline
+
+		10  & 75980.62 &   59.34 &  219.35 & 143.63\\
+		25  & 70205.27 &  385.79 &  558.12 & 432.77\\
+		50  & 69217.86 &  730.82 &  693.42 & 255.70\\
+		100 & 66667.87 &  1681.3 & 1605.63 & 400.55\\
+
+	\end{tabular}
+\end{table}
 
@@
 \subsubsection{Restrict search to a buffered around the least cost paths}
+
 Construct a polygon form the two lines.
@@
 The street and the protected landscape area are both restricted areas, while the power line is preferential.
-This showing, that the way the cost raster is created in the first place can play a vital role, in the end result. So that a nuance, can cause a detour.
+This showing, that the way the cost raster is created in the first place can play a vital role, in the end result.
+So that a nuance, can cause a detour.
 When this behavior occurs, the polygon can not grantee to include the least cost path.
@@
 
-\subsubsection{Restrict search to reachable points}
-For this purpose the aggregated cost as to converted back into a raster.
+%\subsubsection{Restrict search to reachable points}
+%For this purpose the aggregated cost as to converted back into a raster.
 
-\subsubsection{Compare the different Solutions}
 \subsubsection{Check solution for least cost path between different points}

### 2023-03-04 paper/discussion.tex
added Supplement S1
Related works
Examine solutions

@@
+% !TeX spellcheck = en_US
 As shown, the need for computation time increases with the power of two with the resolution. 
@@
 
-Therefore, this worked tries a) to reduce the needed compute power and b) decrease the deviation for a given resolution.
+Therefore, this worked tries a) to reduce the needed to compute power and b) decrease the deviation for a given resolution.
 
@@
 These methods can be used as surrogates for the more complex calculation of the least cost path with the higher resolution.
-In the first method a bi-linear down sampling of the higher resolution raster has been applied and in the second method, the all touched false and true raster where averaged in different ratios, to computed the optimum mixed raster, for the used costs.
+In the first method a bi-linear down sampling of the higher resolution raster has been applied and in the second method, the all touched false and true raster where averaged in different ratios, to compute the optimum mixed raster, for the used costs.
 While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the down sampling method is more
@@
 As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
-This early stopping might result in sub-optimal paths around the end points for some edge cases, where the connection via another neighbor might be more optimal.
+This early stopping might result in suboptimal paths around the end points for some edge cases, where the connection via another neighbor might be more optimal.
 
-The set of rules that are used create the cost raster, includes a rule for creating a buffer around buildings  which is set to the level \textbf{prohibited} areas.
-While the used data set for buildings, also includes power poles, the area of the level \textbf{preferential} power grid, includes isles of high costs. 
+The set of rules that are used create the cost raster, includes a rule for creating a buffer around buildings  which is set to the level \textit{prohibited} areas.
+The resolution of the medium level raster needs to be high enough to crudely show every detail, Hence at least in the magnitude of the minimum object size plus twice its buffer, in this example in the range of at least 25~m or 10~m.
+Although the overall result can change, due super position of objects and there buffer.
+Nevertheless, does higher resolution resemble, the geography more precise, but in addition show paths that are too small, to be taken.
+On the other hand, as the run from the starting point to the end point at a resolution of 2.5~m shows, new ways, that could not not be reached before, do hint inconsistencies of the rules that construct the cost raster.
+
+Object and there buffer might leave a  path that, is smaller than 10~m resolution.
+So a second rule to find a good resolution, that is should be in the magnitude of the size of the structure  to be build.
+While the used data set for buildings, also includes power poles, the area of the level \textit{preferential} power grid, includes isles of high costs. 
 In further works, these should be excluded.
@@
 When calculating several paths from one raster, the benefits of chapter~\ref{subsec:faster-processing-of-the-cost-path-algorithm} is reduced.
-Especially, the pre-calculation on lower resolution raster and buffering of the resulting paths as a restricted search area decreases in it effectivity.
+Especially, the pre-calculation on lower resolution raster and buffering of the resulting paths as a restricted search area decreases in its effectivity.
 As the number of paths rises, fewer of points in higher resolution cost raster will be deselected for the final estimation of the least cost path.
+The least cost algorithm does only select the most cost-effective path.
+Therefore, paths of similar, but slightly worse cost are stay unknown.
+An end-user might be interested in selecting a path from a set of possible paths of similar aggregated costs and apply an adapted set evaluation criteria.
+This can be achieved a by adapting the backtracking, to get an array of paths or by applying perturbation on the costs.
 
@@
 The reasoning of selection the maximum function is, limiting the maximum costs.
-When the \textbf{prohibited} level is used as the highest level.
-Than higher levels are prohibited within the concept.
-This an area even more prohibited, when it is included in two different rules of the level?
+When the \textit{prohibited} level is used as the highest level.
+Then higher levels are \textit{prohibited} within the concept.
+This an area even more \textit{prohibited}, when it is included in two different rules of the level?
 On the other side, This aggregation is unable to differentiate between these nuances, as  the sum function is.
-This way, one can differential between, different sub-levels.
+This way, one can differential between, different sublevels.
 One consequence is, that a small mediation between high and low single cost raster, in the final cost raster is possible.

### 2023-03-04 paper/methods.tex
added Supplement S1
Related works
Examine solutions

@@
 Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
-Lower Saxony\cite{noauthor_opengeodatani_2022}.
+Lower Saxony~\cite{noauthor_opengeodatani_2022}.
 In addition, local planning geodata for the land usage are taken
-from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)\cite{noauthor_metropolplaner_2022}
+from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)~\cite{noauthor_metropolplaner_2022}.
 
-PyWPS\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
+PyWPS~\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
 The initial implementation of the least cost path algorithm is based on the implementation for the QGIS-Plugin
-'Least Cost Path'\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.
+'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.
 
@@
 For every area in the study area, that is not covered by any layer, is given the default cost.
-The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from preferential areas with
-very low costs, via no restrictions, which is the default, used when no other layers are covering the local area,
-to restricted, strongly restricted and prohibited areas with high costs.
+The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from \textit{preferential} areas with
+very low costs, via \textit{no restriction}s, which is the default, used when no other layers are covering the local area,
+to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
 These higher costs resemble the degree how much a local area should be avoided, while routing the path.
@@
 the algorithm is willing to go, for avoiding an area of high costs.
-Thus, as prohibited areas describe a legal obligation, not to use these areas or only to the utmost minimum,
+Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
 the weight that resembles the costs for these types of areas, has to be especially high. \\
@@
 Instead, the execution stops early after reaching the end point.
-Because the end point ends at a power transformer, which is a building type, the paths ends at in a Prohibited Area.
+Because the end point ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} Area.
 Therefore, areas even farther away from the starting point have been explored first.

### 2023-03-04 paper/paper-CableRoute.tex
added Supplement S1
Related works
Examine solutions

@@
 \newacronym{corr}{corr}{corrected}
+\newacronym{MCDA}{MCDA}{multi-criteria decision analysis}
 
@@
 	\begin{abstract}
-		add as final part
+		The least cost paths for a power line route from a power station near the German shore line is computed to a power station further inland.
+		Down sampling and clipping studied for a speed-up of the computation.
+		The path from clipped cost raster often matches the original path.
+		The speed-up of the clipping can not be known before hand.
 	\end{abstract}
@@
 		<concept>
-		<concept_id>10010520.10010553.10010562</concept_id>
-		<concept_desc>Computer systems organization~Embedded systems</concept_desc>
+		<concept_id>10010147.10010341</concept_id>
+		<concept_desc>Computing methodologies~Modeling and simulation</concept_desc>
 		<concept_significance>500</concept_significance>
@@
 		<concept>
-		<concept_id>10010520.10010575.10010755</concept_id>
-		<concept_desc>Computer systems organization~Redundancy</concept_desc>
-		<concept_significance>300</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10010520.10010553.10010554</concept_id>
-		<concept_desc>Computer systems organization~Robotics</concept_desc>
-		<concept_significance>100</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10003033.10003083.10003095</concept_id>
-		<concept_desc>Networks~Network reliability</concept_desc>
+		<concept_id>10002944.10011123.10011133</concept_id>
+		<concept_desc>General and reference~Estimation</concept_desc>
 		<concept_significance>100</concept_significance>
@@
 
-	\ccsdesc[500]{Computer systems organization~Embedded systems}
-	\ccsdesc[300]{Computer systems organization~Redundancy}
-	\ccsdesc{Computer systems organization~Robotics}
-	\ccsdesc[100]{Networks~Network reliability}
+	\ccsdesc[500]{Computing methodologies~Modeling and simulation}
+	\ccsdesc[100]{General and reference~Estimation}
 
@@
 	%% the work being presented. Separate the keywords with commas.
-	\keywords{datasets, neural networks, gaze detection, text tagging}
+	\keywords{GIS, least cost path}
 
@@
 	by the local population have to be taken in consideration.
-	Also technical aspects, as the effects on the grid stability might be further points to consider.\cite{schafer_understanding_2022}
+	Also technical aspects, as the effects on the grid stability might be further points to consider~\cite{schafer_understanding_2022}.
 
-	Due to the increasing need for renewable energy wind off shore wind turbines are source which currently supplies 5.5~\% annual percentage in 2020 of the German energy mix.~\cite{noauthor_nettostromerzeugung_2021}
-	For wind energy in general highest efficiency can only be archived with high wind speeds. Highest wind speeds are reached along the cost lines in the north of Germany.
+	Due to the increasing need for renewable energy wind off shore wind turbines are source which currently supplies 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}
+.	For wind energy in general highest efficiency can only be archived with high wind speeds.
+	Highest wind speeds are reached along the cost lines in the north of Germany.
 	In contrast, most consumers, as industry are situated in south Germany.
@@
 
-	Other methods as the multi-criteria decision analysis take further aspects and stakeholder as experts, public and decision maker and into consideration.\cite{bertsch_participatory_2016}
+	Other methods as the multi-criteria decision analysis take further aspects and stakeholder as experts, public and decision maker and into consideration~\cite{bertsch_participatory_2016}.
 
@@
 	\section{Methods}\label{sec:methods}
-
 	\input{methods}
@@
 	\section{Results}\label{sec:results}
-
 	\input{results}
@@
 	\section{Related Works}\label{sec:related-works}
+	In contrast to GIS analysis, further effort can be done with a \acrfull{MCDA}\todo{Pohekar and Ramachandran 2004, Wang 2009, Riberio 2011}.
+	The uncertainties connected with \acrshort{MCDA} are data uncertainties, preferential uncertainties and model uncertainties.
+	These uncertainties are examined by a with a sensitivity analysis \todo{Eßer-Frey} or simulation~\cite{butler_simulation_1997}.
+
+	The different models of the power market models span from not modeling the grid at all to models that includes Kirchhoffs laws.
+	However, the actual planning of power grids is computational complex and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
+	Some models include not only the grid of a nation alone~\cite{dietrich_will_2010}, but models do not include both model the expansion planing for a longer time period and the effect of power transition~\cite{bertsch_participatory_2016}.
+	But economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an important role~\cite{hauff_gesellschaftliche_2011}.
 
@@
 	When using the researched strategies, the on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
-	\todo{Befragung: local population, relative weights per district?}
+	\todo{survey: local population, relative weights per district?}
 
@@
 %% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+%%\bibliographystyle{ACM-Reference-Format}
+\bibliographystyle{unsrt}
 \bibliography{Web_Processing.bib}

### 2023-03-04 paper/results.tex
added Supplement S1
Related works
Examine solutions

@@
 When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
-Preferential Level and less the NoRestriction Level.\todo{Interpretation:}
+\textit{Preferential Level} and less the\textit{NoRestriction} Level.\todo{Interpretation:}
 The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
@@
 
-\subsubsection{Check solution for least cost path between different points}
+\subsubsection{Examine the proposed solutions}
+To broaden the view and verify the current result, four different routes should be found with the presented strategies applied.
+Two paths form the starting point to two new points in the south east of the investigated area and two paths from the north, north east of the investigates area to the ending point.
+
+For three of the four routes the least cost path computation from the clipped raster, was able to compute the expected result.
+For the forth path, the
+least cost path from the 5~m resolution raster was out side of the buffer around the least cost path from the 50~m resolution raster.
+The speed up from the clipping of the higher resolution raster depends on the amount of pixels that, have been clipped. 
+Bi-linear down sampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
+The resolution corrected aggregated costs of the least cost path from the down sampled raster are higher, than those, from the higher 5~m resolution raster and the medium 10~m resolution raster.
+In addition, is the distance, from this path to the path from the high resolution raster much higher, than the distance of path from the original 10~m resolution raster to the path from the 5~m resolution raster (see table~\ref{tab:6}).
+\todo{interpretation Why, down sampling so different???}
+
+\begin{table}[h!]
+	\caption{Length of the path, the resolution corrected costs and the mean distance to the path created from the 5~m resolution all touched false raster for the four control routes, for the reference path of constructed from the 5~m  and 10~m raster and the down 5~m to 10~m down sampled raster and the 5~m clipped raster.}
+	\label{tab:6}
+	\centering
+	\begin{tabular}{ r  r  r  r  r}
+		Route & Method & $length /m$ & $costs_{al=f}$ & $d_{mean}$ /m \\
+		\hline
+		P1-E & 5~m 			& 107889.58 & 208547.79 &        \\
+		 	 & Clipped 		& 107889.58 & 208547.79 &   0.0  \\
+		 	 & Down			&  96754.20 & 212910.95 & 628.05 \\
+		 	 & 10~m 		& 107232.94 & 203010.15 & 103.49 \\
+		\hline
+		P2-E & 5~m 			& 103706.41 & 155567.86 &        \\
+		 	 & Clipped 		& 103706.41 & 155567.86 &   0.00 \\
+		 	 & Down 	    &  92403.29 & 158238.58	& 639.88 \\
+		 	 & 10~m 		& 104249.94 & 149899.72 & 177.68 \\
+		\hline
+		S-P3 & 5~m 			& 102187.09 & 34503.81 	&         \\
+			 & Clipped 		&  90377.39 & 37926.14 	& 4465.37 \\
+			 & Down 	    &  94125.55 & 37574.93 	&  742.41 \\
+			 & 10~m 		& 102461.58 & 32445.96 	&   81.18 \\
+		\hline
+		S-P4 & 5~m 			& 96449.22 	& 33865.53 	&  		\\
+		 	 & Clipped 		& 96449.22 	& 33865.53 	& 0.00 	\\
+			 & Down 		& 87861.12 	& 36462.72 	& 796.40\\
+			 & 10~m 		& 96739.47 	& 31899.31	& 83.46 \\
+
+	\end{tabular}
+\end{table}

### 2023-03-07 paper/discussion.tex
spelling
Begining of the introduction and the related works section updated

@@
 The set of rules that are used create the cost raster, includes a rule for creating a buffer around buildings  which is set to the level \textit{prohibited} areas.
-The resolution of the medium level raster needs to be high enough to crudely show every detail, Hence at least in the magnitude of the minimum object size plus twice its buffer, in this example in the range of at least 25~m or 10~m.
+The resolution of the medium level raster needs to be high enough to crudely show every detail, Hence at least in the magnitude of the minimum object size plus twice its buffer, in this example in the range of at least 25~m or 10~m to include the streets in all touched false raster.
+Other details as rivers and houses, ae already includes in lower resolution raster, because of the buffering.
 Although the overall result can change, due super position of objects and there buffer.
 Nevertheless, does higher resolution resemble, the geography more precise, but in addition show paths that are too small, to be taken.
-On the other hand, as the run from the starting point to the end point at a resolution of 2.5~m shows, new ways, that could not not be reached before, do hint inconsistencies of the rules that construct the cost raster.
+On the other hand, as the run from the starting point to the end point at a resolution of 2.5~m shows, new ways, that could not be reached before, do hint inconsistencies of the rules that construct the cost raster.
 
 Object and there buffer might leave a  path that, is smaller than 10~m resolution.
-So a second rule to find a good resolution, that is should be in the magnitude of the size of the structure  to be build.
+So a second rule to find a good resolution, that it should be in the magnitude of the size of the structure  to be build.
 While the used data set for buildings, also includes power poles, the area of the level \textit{preferential} power grid, includes isles of high costs. 
@@
 No problem because buffering is included in the cost raster.
-Not algorithmic speed-ups as pre-compilation, JIT, vector units, multi-core
+Not algorithmic speed-ups as pre-compilation, JIT, vector units, multi-core, weight range

### 2023-03-07 paper/methods.tex
spelling
Begining of the introduction and the related works section updated

@@
 	}%
-	\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with weither all-touched False (above), or all touched True (below).}
+	\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with either all-touched False (above), or all touched True (below).}
 	\label{fig:alltouched}
@@
 
-	\subfloat[\centering Cost Raster.]{{\includegraphics[width=.45\linewidth]{./images/CostRasterExample.png} }}%
-	\qquad
-	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.45\linewidth]{./images/AggregatedCosts.png} }}%
-	\qquad
-	\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPathExample.png} }}%
+	\subfloat[\centering Cost Raster.]{{\includegraphics[width=.30\linewidth]{./images/CostRasterExample_cut.png} }}%
+	\enskip
+	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.30\linewidth]{./images/AggregatedCosts_cut.png} }}%
+	\enskip
+	\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.30\linewidth]{./images/LeastCostPathExample_cut.png} }}%
 
@@
 \end{figure}
+
+\todo{Were to put the following part?}
+The aggregation is implemented in a priority queue. 
+Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.
+The aggregation is implemented with early stopping. 
+After the last end point has been has got an aggregated cost the aggregation stops and the back tracking starts.

### 2023-03-07 paper/paper-CableRoute.tex
spelling
Begining of the introduction and the related works section updated

@@
 
-	Sometimes, finding the shortest path is not sufficient.
+	The German power system add a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
+	At the same time old power plants e.g. nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
+	These older power plants were mainly situated in southern and central Germany. 
+	At the same time , the power sink industrial and private demand do not shift northwards.
+	Therefore, the renewable energies have to be transported from north to south which increases the congestion in the power grid. 
+	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.
+
+	But for building power lines, finding the shortest path is not sufficient.
 	Additional parameters also have to be taken into consideration.
-	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline
-	\cite{suleiman_optimal_2015}.
-	For planing the additional routes for a power grid, additional aspects as legal regulations and acceptance
+	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
+	For planing the additional routes for a power grid, further aspects as legal regulations and acceptance
 	by the local population have to be taken in consideration.
@@
 	Due to the increasing need for renewable energy wind off shore wind turbines are source which currently supplies 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}
-.	For wind energy in general highest efficiency can only be archived with high wind speeds.
+.	
 	Highest wind speeds are reached along the cost lines in the north of Germany.
-	In contrast, most consumers, as industry are situated in south Germany.
-	Source and consumer have to be linked via new power lines.
 
@@
 	The vertices of the graph are the centers on neighboring pixels in al possible eight directions.
-	The weights of the edges are the local costs to travel from one pixel to the neighboring pixel.
+	This enables to find routes on graphs that are not predefined like road notworks.
+	The weights between the edges are the local costs to travel from one pixel center to the neighboring pixel center.
 	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
@@
 	For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.
-
-	\todo{Were to put the following part?}
-	The aggregation is implemented in a priority queue. 
-	Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.
-	The aggregation is implemented with early stopping. 
-	After the last end point has been has got an aggregated cost the aggregation stops and the back tracking starts.
 
@@
 	\section{Related Works}\label{sec:related-works}
-	In contrast to GIS analysis, further effort can be done with a \acrfull{MCDA}\todo{Pohekar and Ramachandran 2004, Wang 2009, Riberio 2011}.
-	The uncertainties connected with \acrshort{MCDA} are data uncertainties, preferential uncertainties and model uncertainties.
-	These uncertainties are examined by a with a sensitivity analysis \todo{Eßer-Frey} or simulation~\cite{butler_simulation_1997}.
 
-	The different models of the power market models span from not modeling the grid at all to models that includes Kirchhoffs laws.
-	However, the actual planning of power grids is computational complex and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
-	Some models include not only the grid of a nation alone~\cite{dietrich_will_2010}, but models do not include both model the expansion planing for a longer time period and the effect of power transition~\cite{bertsch_participatory_2016}.
-	But economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an important role~\cite{hauff_gesellschaftliche_2011}.
+	Older models are e.g.
+are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
+
+	In contrast to a GIS analysis, these further factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
+
+	Stakeholder as decision maker can be included and combined with an expert system~\cite{bertsch_participatory_2016}.
+
+	The uncertainties connected with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are examined by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
+	Both inter and intra criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.
+
+	The different models of the power system, that are used in the expert system span from not modeling the grid at all to models that includes Kirchhoffs laws~\cite{bertsch_participatory_2016}.
+	The calculation of a power system is complex, as changing one edge will change the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
+	Some models include not only the grid of a nation alone, but consider neighboring states or the overall European power system~\cite{dietrich_will_2010}.
 
@@
 
-\begin{acks}
-	$\ldots$
-\end{acks}
+%%\begin{acks}
+%%	$\ldots$
+%%\end{acks}
 
@@
 %%\bibliographystyle{ACM-Reference-Format}
-\bibliographystyle{unsrt}
+\bibliographystyle{unsrtdin}
 \bibliography{Web_Processing.bib}
@@
 %% If your work has an appendix, this is the place to put it.
-\appendix
+%%\appendix
 
-\section{Research Methods}\label{sec:research-methods}
+%%\section{Research Methods}\label{sec:research-methods}
 
-	\subsection{Part One}\label{subsec:part-one}
+%%	\subsection{Part One}\label{subsec:part-one}
 
-	\subsection{Part Two}\label{subsec:part-two}
+%%	\subsection{Part Two}\label{subsec:part-two}
 
-	\section{Online Resources}\label{sec:online-resources}
+%%	\section{Online Resources}\label{sec:online-resources}
 

### 2023-03-08 paper/discussion.tex
Updated Discussion & Conclussion
readability test for the first two sections

@@
 
-Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100~m!
-No problem because buffering is included in the cost raster.
-Not algorithmic speed-ups as pre-compilation, JIT, vector units, multi-core, weight range
+The fact that the differences in the costs between the paths from the all touched false to the all touched true rasters stay almost constant over the different resolution is probably due to the fact that, the costs per resolution stays in narrow range of costs.
+This effect of the similar aggregated costs per resolution could also be seen in the testing paths, even when the paths changed strongly with the change of resolution or algorithm applied.
+Which might be a hint for an evenly spatial distribution of the costs.
+
+The all touched true cost raster show every detail, but the sampling with all touched true increases the size of the features.
+The fact that the aggregated costs per resolution for all touched true rasters overestimates the costs when computing the path from a low resolution raster, might be due to the fact, that the values of the costs are un-evenly distributed, but that the high costs are much more frequent differ and the costs are exponential scaled exponentially.

### 2023-03-08 paper/methods.tex
Updated Discussion & Conclussion
readability test for the first two sections

@@
 
-PyWPS~\cite{noauthor_welcome_2016} is used to offer the least cost path algorithm as a \acrfull{wps}.
+PyWPS~\cite{noauthor_welcome_2016} is used to provide the least cost path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
+As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sending the cost raster and receiving the resulting least cost path.
 The initial implementation of the least cost path algorithm is based on the implementation for the QGIS-Plugin
@@
 Filtering the layers of the files for special attributes enables to further differentiate.
-For examples makes it possible to differentiate between heath and uncultivated land.
+For examples makes it possible to differentiate between heath and uncultivated land in the land coverage.
 Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
@@
 Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
-The costs of all layers of is aggregated with the maximum function.
+The costs of all layers is aggregated with the maximum function.
 Thus, an area covered by multiple layers is uniformly used with the highest costs.
-For every area in the study area, that is not covered by any layer, is given the default cost.
+Any place in the study area, that is not covered by any layer and thus does not have a weight yet, is given the default cost.
+
 The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from \textit{preferential} areas with
@@
 to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
-These higher costs resemble the degree how much a local area should be avoided, while routing the path.
+These higher costs resemble the degree how much a place of this cost should be avoided, while routing the path.
 The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
-the algorithm is willing to go, for avoiding an area of high costs.
+the algorithm is willing to go, for avoiding an area of this high costs.
 Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
-the weight that resembles the costs for these types of areas, has to be especially high. \\
+the weight that resembles the costs for these types of areas, has to be especially high.\\
+
+All these layers are provided as vectors.
+While the least cost algorithm use raster data.
 The rasterization transforms a vector into a raster.
-The rasterization can be executed in to different ways.
+The rasterization can be executed in two different ways.
 In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
 given resolution and the new affine transformation and the coordinate reference system of the vector.
-Both rasterization techniques differ in how the pixels are selected, that describe, that a geo-object overlays the pixel.
-The pixels can either be described as overlaid with a geo-object, when either the center of that pixel is overlaid
+Both rasterization techniques differ in the selection of the pixel, that describe the original polygon.
+The pixels can be selected , if either the center of that pixel is overlaid
 by the geo-object, or any part of the pixel is overlaid.
-Hence, the version with any part overlaid is called all touched True.
-The version where a overlaid pixel center is require is thus called all touched False.
+The version with any part overlaid is called all touched True.
+The version where a overlaid pixel center is require is called all touched False.
 All touched False is considered the default (see figure~\ref{fig:alltouched}).
@@
 \end{table}
-The completed list of layers and the processing applied to them, can be found in \textbf{Supplement S1}.
 
-The whole stack of the cost raster, via the aggregated costs to the resulting least cost path can is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).
-Because of the early stopping strategy used in the aggregation sub-paths with high costs have not been fully computed.
-Instead, the execution stops early after reaching the end point.
+The completed list of layers and their processing, can be found in \textbf{Supplement S1}.
+
+All three steps of the generation of the least cost path: generation of the cost raster, aggregation and back tracking is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).
+
+The chosen implementation applies early stopping.
+Therefore, the costs for points that are not needed to try to connect to the end point are not aggregated.
+After an aggregated cost for every end point has been found, the aggregation stops and the back tracking starts.
 Because the end point ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} Area.
@@
 
-\todo{Were to put the following part?}
-The aggregation is implemented in a priority queue. 
-Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.
-The aggregation is implemented with early stopping. 
-After the last end point has been has got an aggregated cost the aggregation stops and the back tracking starts.
+%% The aggregation is implemented in a priority queue. 
+%% Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.

### 2023-03-08 paper/paper-CableRoute.tex
Updated Discussion & Conclussion
readability test for the first two sections

@@
 	\begin{abstract}
-		The least cost paths for a power line route from a power station near the German shore line is computed to a power station further inland.
-		Down sampling and clipping studied for a speed-up of the computation.
+		The least cost paths for a power line route from a power station is computed to a power station further inland.
+		Down sampling and clipping as studied as techniques to speed-up the computation.
 		The path from clipped cost raster often matches the original path.
@@
 
-	The German power system add a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
-	At the same time old power plants e.g. nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
+	The German power system adds a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
+	At the same time old power plants e.g.\ nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
 	These older power plants were mainly situated in southern and central Germany. 
-	At the same time , the power sink industrial and private demand do not shift northwards.
+	The power sink, which is industrial and private demand does not shift northwards.
 	Therefore, the renewable energies have to be transported from north to south which increases the congestion in the power grid. 
@@
 	by the local population have to be taken in consideration.
-	Also technical aspects, as the effects on the grid stability might be further points to consider~\cite{schafer_understanding_2022}.
+	Also technical aspects, as the effects on the grid stability are further points take into consideration~\cite{schafer_understanding_2022}.
 
-	Due to the increasing need for renewable energy wind off shore wind turbines are source which currently supplies 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}
-.	
-	Highest wind speeds are reached along the cost lines in the north of Germany.
+	Due to the increasing need for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.
 
-	Other methods as the multi-criteria decision analysis take further aspects and stakeholder as experts, public and decision maker and into consideration~\cite{bertsch_participatory_2016}.
+	Other methods as the \acrfull{MCDA}  consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).
 
 	For planning a possible route for a power line the least cost path algorithm is applied.
-	The least cost path algorithm is a Dijkstra algorithm applied on a raster map.
-	The vertices of the graph are the centers on neighboring pixels in al possible eight directions.
+	The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
+	The vertices of the graph are the pixel centers, that are connected to the eight neighboring pixel centers via the edges.
 	This enables to find routes on graphs that are not predefined like road notworks.
-	The weights between the edges are the local costs to travel from one pixel center to the neighboring pixel center.
+	The weights of the edges are the local costs to travel from one pixel center to the neighboring pixel center.
 	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
 	The least cost path algorithm consists of at least two sequential steps.
-	1) The first steps is to aggregate the costs of traveling from the starting point to a given set of end points.
-	This steps generates the aggregated cost raster of the total cost that are needed to travel to a point, from the cost raster.
+	1) The first step is to aggregate the costs of traveling from the starting point to a given set of end points.
+	This step generates the aggregated cost raster of traveling from the starting point to any point of the cost raster.
 	2) In the next step the back tracking, the route of the actual least cost path is calculated.
 	For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.
+
+	Some implementations switch the role of start and ending points, so that either many starting points and one end point, or one end point and many endpoints can be used.
+	In some implementation there is a extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
+	path to the starting point as integer values.
 
@@
 
-	Older models are e.g.
-are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
+	Older models are e.g.\ are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
 
@@
 	\section{Conclusion}\label{sec:conclusion}
-	When using the researched strategies, the on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
-	\todo{survey: local population, relative weights per district?}
+	The cost of computation of the least cost path scales with the square of the resolution.
+	The difference between the aggregated costs of the paths per resolution computed from the all touched false raster and the all touched true raster only shows a linear decrease.
+	Hence, the gain in precision per compute time decreases.
+	The presented methods tried to circumvent this bottleneck.
+	When using the researched strategies, on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
+	The paths might change very strongly even with small changes in total costs.
+	Hence to develop an alternative backtracking algorithm that, that generates a set of possible paths within a corridor of costs might be a good strategy, to offer alternative routes and show variations between these routes at the same time.
+	An alternative method for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
+	This in turn increases the computation time again.
+
+	In an actual search for least cost path a survey might be used as method to estimate the costs.
+	When the sample size of the survey is large enough, the
+	weights might take local differences relative acceptance into consideration.
+
+	Only changes in the algorithm have been applied in the search for a speed-up.
+	There a still other methods as just-in-time compilation.
 

### 2023-03-08 paper/results.tex
Updated Discussion & Conclussion
readability test for the first two sections

@@
 
-At the same time the differences in the aggregated costs stay constant.
-\todo{Interpretation} When normalize the aggregated costs by the resolution.
-Than On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
+At the same time the differences in the aggregated costs stay almost constant.
+When normalize the aggregated costs by the resolution.
+On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
 linear with the resolution.
@@
 When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
-\textit{Preferential Level} and less the\textit{NoRestriction} Level.\todo{Interpretation:}
+\textit{Preferential Level} and less the\textit{NoRestriction} Level.
 The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
@@
 In addition, is the distance, from this path to the path from the high resolution raster much higher, than the distance of path from the original 10~m resolution raster to the path from the 5~m resolution raster (see table~\ref{tab:6}).
-\todo{interpretation Why, down sampling so different???}
+\todo{interpretation: Why result from down sampling so different? Don't know.}
 

### 2023-03-08 paper/discussion.tex
grammar + change to en_GB

@@
-% !TeX spellcheck = en_US
+% !TeX spellcheck = en_GB
 As shown, the need for computation time increases with the power of two with the resolution. 
@@
 As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
-This early stopping might result in suboptimal paths around the end points for some edge cases, where the connection via another neighbor might be more optimal.
+This early stopping might result in suboptimal paths around the end points for some edge cases, where the connection via another neighbour might be more optimal.
 
@@
 On the other side, This aggregation is unable to differentiate between these nuances, as  the sum function is.
-This way, one can differential between, different sublevels.
+This way, one can differential between, different sub-levels.
 One consequence is, that a small mediation between high and low single cost raster, in the final cost raster is possible.
@@
 In cause the default level would rise, the effect probably would flip for low resolution raster.
-For high resolution raster, this effect would still hold true, because the fact, that the raster pixel center is used for sampling, reflects the original geometry better.
+For high resolution raster, this effect would still hold true, because the fact, that the raster pixel centre is used for sampling, reflects the original geometry better.
 

### 2023-03-08 paper/methods.tex
grammar + change to en_GB

@@
-% !TeX spellcheck = en_US
+% !TeX spellcheck = en_GB
 
-We retrieve a set of different spacial data-sets from  public sources as a basis to create the cost raster.
-Field of study are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
+We retrieve a set of different spacial data-sets from  public sources as a basis for creating the cost raster.
+The study area are the counties of Cuxhaven and Osterholz in the state of Lower Saxony, Germany.
 Areas protected by different European and National conservation laws are provided by the German Environment Agency
@@
 and Geodesy~\cite{noauthor_digitales_2021}.
-The nationwide power grid (tags: 'power': line) has been retrieved via OpenStreetMap~\cite{boeing_osmnx_2017}.
-Local data as houses at Level of Detail 1 are offered by the State Office for Geoinformation and Land Surveying of
+The national power grid (tags: 'power': line) has been retrieved via OpenStreetMap~\cite{boeing_osmnx_2017}.
+Local data as houses at Level of Detail 1 are provided by the State Office for Geoinformation and Land Surveying of
 Lower Saxony~\cite{noauthor_opengeodatani_2022}.
-In addition, local planning geodata for the land usage are taken
-from 'Metropolplaner' (Planing data Lower Saxony \& Bremen)~\cite{noauthor_metropolplaner_2022}.
+In addition, local planning geodata for the land use are taken
+from 'Metropolplaner' (Planning data Lower Saxony \& Bremen)~\cite{noauthor_metropolplaner_2022}.
 
 PyWPS~\cite{noauthor_welcome_2016} is used to provide the least cost path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
-As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sending the cost raster and receiving the resulting least cost path.
-The initial implementation of the least cost path algorithm is based on the implementation for the QGIS-Plugin
+As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster and receives the resulting least cost path.
+The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
 'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.
 
-The different layers from the different entities are optionally filtered, buffered and then rasterized.
-Filtering the layers of the files for special attributes enables to further differentiate.
-For examples makes it possible to differentiate between heath and uncultivated land in the land coverage.
-Adding a buffer can be used either used to convert a line objects as power grids and streets into a polygon with the
+The different layers of the different entities are optionally filtered, buffered and then rasterised.
+Filtering the layers of the files for special attributes enables further differentiation.
+For examples, it possible to differentiate between heath and uncultivated land in the land cover.
+Adding a buffer can be used either used to convert a line objects such as power grids and streets into a polygon with the
 correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
-Each of theses rasterized layers are given a weight, or cost that expresses the cost of using land covered by this layer.
+Each of theses rasterised layers are given a weight, or cost which expresses the cost of using land covered by this layer.
 The costs of all layers is aggregated with the maximum function.
-Thus, an area covered by multiple layers is uniformly used with the highest costs.
-Any place in the study area, that is not covered by any layer and thus does not have a weight yet, is given the default cost.
+Thus, an area covered by several layers is uniformly used with the highest cost.
+Any place in the study area, that is not covered by any layer and thus does not yet have a weight, is given the default cost.
 
-The costs have been grouped into different levels~(see table~\ref{tab:1}) starting from \textit{preferential} areas with
+The costs have been grouped into different levels~(see table~\ref{tab:1}), starting from \textit{preferential} areas with
 very low costs, via \textit{no restriction}s, which is the default, used when no other layers are covering the local area,
 to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
-These higher costs resemble the degree how much a place of this cost should be avoided, while routing the path.
-The ratio of the higher costs to the lower costs directly translates into the additional diversion in pixels
-the algorithm is willing to go, for avoiding an area of this high costs.
+These higher costs resemble the degree to which a place with this cost should be avoided, while routing the path.
+The ratio of the higher costs to the lower costs directly translates into the additional detour in pixels
+the algorithm is willing to go.
 Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
@@
 All these layers are provided as vectors.
-While the least cost algorithm use raster data.
-The rasterization transforms a vector into a raster.
-The rasterization can be executed in two different ways.
-In both ways, the rasterization can be imaged, as the old vector is layered over the new raster grid with the new
+The Least Cost Path algorithm uses raster data.
+Rasterisation transforms a vector into a raster.
+Rasterisation can be done in two different ways.
+In both ways, the rasterisation can be imaged, as the old vector is superimposed on the new raster grid with the new
 given resolution and the new affine transformation and the coordinate reference system of the vector.
-Both rasterization techniques differ in the selection of the pixel, that describe the original polygon.
-The pixels can be selected , if either the center of that pixel is overlaid
+Both rasterisation techniques differ in the selection of the pixel, that describe the original polygon.
+The pixels can be selected , if either the centre of that pixel is overlapped
 by the geo-object, or any part of the pixel is overlaid.
-The version with any part overlaid is called all touched True.
-The version where a overlaid pixel center is require is called all touched False.
+The version with any part overlapped is called all touched True.
+The version where a overlapped pixel centre is require is called all touched False.
 All touched False is considered the default (see figure~\ref{fig:alltouched}).
@@
 
-The completed list of layers and their processing, can be found in \textbf{Supplement S1}.
+The complete list of layers and their processing, can be found in \textbf{Supplement S1}.
 
-All three steps of the generation of the least cost path: generation of the cost raster, aggregation and back tracking is shown with an example for a starting cost raster of 50~m resolution and a rasterization with all touched False (see \ref{fig:costs2path}).
+All three steps of the generation of the least cost path: generation of the cost raster, aggregation and backtracking is shown with an example for a starting cost raster of 50~m resolution and a rasterisation with all touched False (see \ref{fig:costs2path}).
 
@@
 Therefore, the costs for points that are not needed to try to connect to the end point are not aggregated.
-After an aggregated cost for every end point has been found, the aggregation stops and the back tracking starts.
-Because the end point ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} Area.
-Therefore, areas even farther away from the starting point have been explored first.
+After finding an aggregated cost for every end point, the aggregation stops and the backtracking starts.
+Because the path ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} area.
+Therefore, areas even further away from the starting point have been explored first.
 
@@
 
-%% The aggregation is implemented in a priority queue. 
-%% Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbors, than a point with higher costs.
+%% Aggregation is implemented in a priority queue. 
+%% Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbours, than a point with higher cost.

### 2023-03-08 paper/paper-CableRoute.tex
grammar + change to en_GB

@@
-% !TeX spellcheck = en_US
+% !TeX spellcheck = en_GB
 
@@
 	\begin{abstract}
-		The least cost paths for a power line route from a power station is computed to a power station further inland.
-		Down sampling and clipping as studied as techniques to speed-up the computation.
-		The path from clipped cost raster often matches the original path.
-		The speed-up of the clipping can not be known before hand.
+		The least cost paths for a power line from a power station at the cost to a power station further inland are calculated.
+		Down-sampling and clipping are investigated as techniques to speed up the calculation.
+		The path from the clipped cost raster is often the same as the original path.
+		The speed-up of the clipping can not be known in advance.
 	\end{abstract}
@@
 
-	The German power system adds a lot of renewable energy sources in the north, because wind energy reaches highest efficiency with high wind speeds.
-	At the same time old power plants e.g.\ nuclear, hard coal and lignite are going out off service~\cite{eser-frey_analyzing_2012}.
-	These older power plants were mainly situated in southern and central Germany. 
-	The power sink, which is industrial and private demand does not shift northwards.
-	Therefore, the renewable energies have to be transported from north to south which increases the congestion in the power grid. 
+	The German power system is adding a lot of renewable energy sources in the north, because wind energy reaches its highest efficiency with high wind speeds.
+	At the same time old power plants e.g.\ nuclear, hard coal and lignite are being phased out~\cite{eser-frey_analyzing_2012}.
+	These older power plants were mainly located in southern and central Germany. 
+	The energy sink,  industrial and private demand, is not shifting north.
+	Therefore, the renewable energy has to be transported from north to south which increases the congestion in the power grid. 
 	The amount of offshore wind power, that the German energy system can use, can be greatly increased by adding new power lines~\cite{leuthold_nodal_2005}.
 
-	But for building power lines, finding the shortest path is not sufficient.
-	Additional parameters also have to be taken into consideration.
+	However, it is not enough to find the shortest route when building new power lines.
+	Other parameters have to be taken into consideration.
 	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
-	For planing the additional routes for a power grid, further aspects as legal regulations and acceptance
+	When planing the additional routes for a power grid, further aspects such as legal regulations and acceptance
 	by the local population have to be taken in consideration.
@@
 
-	Due to the increasing need for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.
+	Due to the increasing demand for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.
 
-	Other methods as the \acrfull{MCDA}  consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).
+	Other methods such as the \acrfull{MCDA}  consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).
 
-	For planning a possible route for a power line the least cost path algorithm is applied.
+	The least cost path algorithm is used to plan a possible route for a power line.
 	The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
-	The vertices of the graph are the pixel centers, that are connected to the eight neighboring pixel centers via the edges.
-	This enables to find routes on graphs that are not predefined like road notworks.
-	The weights of the edges are the local costs to travel from one pixel center to the neighboring pixel center.
-	The costs might be physical costs, as the local slope, but also can be composed on other factors as the acceptance rate for the land usage of a given class.
-	The least cost path algorithm consists of at least two sequential steps.
-	1) The first step is to aggregate the costs of traveling from the starting point to a given set of end points.
-	This step generates the aggregated cost raster of traveling from the starting point to any point of the cost raster.
-	2) In the next step the back tracking, the route of the actual least cost path is calculated.
-	For each ending point the path via the lowest cost neighbor is taken until arriving at the starting point.
+	The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
+	This makes it possible to find routes on graphs that are not predefined, such as road networks.
+	The weights of the edges are the local costs of travel from one pixel center to the neighbouring pixel center.
+	The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate for the land usage of a given class.
+	The Least Cost Path algorithm consists of at least two successive steps.
+	1) The first step is to aggregate the costs of travelling from the starting point to a given set of end points.
+	This step generates the aggregated cost raster of travelling from the starting point to any point of the cost raster.
+	2) In the next step the backtracking, the route of the actual Least Cost Path is calculated.
+	For each end point the path via the lowest cost neighbour is taken until the starting point is reached.
 
-	Some implementations switch the role of start and ending points, so that either many starting points and one end point, or one end point and many endpoints can be used.
-	In some implementation there is a extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
+	Some implementations switch the role of start and ending points, so that either many start points and one single end point, or one single end point and many end points can be used.
+	In some implementation there is an extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
 	path to the starting point as integer values.
@@
 
-	Older models are e.g.\ are optimizing the well fare, economic modeling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
+	Older models are e.g.\ are optimizing the well fare, economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
 
@@
 
-	The different models of the power system, that are used in the expert system span from not modeling the grid at all to models that includes Kirchhoffs laws~\cite{bertsch_participatory_2016}.
+	The different models of the power system, that are used in the expert system span from not modelling the grid at all to models that includes Kirchhoffs laws~\cite{bertsch_participatory_2016}.
 	The calculation of a power system is complex, as changing one edge will change the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
-	Some models include not only the grid of a nation alone, but consider neighboring states or the overall European power system~\cite{dietrich_will_2010}.
+	Some models include not only the grid of a nation alone, but consider neighbouring states or the overall European power system~\cite{dietrich_will_2010}.
 

### 2023-03-08 paper/results.tex
grammar + change to en_GB

@@
 On one hand, while the area, of the pixel is increased for all touched True, also the area for which the cost is overestimated increases.
-On the other hand while the pixel size for all touched  False increases, not only the over estimated area increases for that pixel as describes for all touched true, but in addition the object is less probable situated in the center of the pixel.
-The consequence of the pixel is, that will decreasing resolution that object is not included in the rasterziation.
+On the other hand while the pixel size for all touched  False increases, not only the over estimated area increases for that pixel as describes for all touched true, but in addition the object is less probable situated in the centre of the pixel.
+The consequence of the pixel is, that will decreasing resolution that object is not included in the rasterisation.
 Hence, hence for all touched false lower resolution  leads ot a loss of information.
@@
 Objects that are small on in one dimension as streets will be included in a stochastic manner.
-Described by the likely of that a object of that sized overlaps with the pixel center.
+Described by the likely of that a object of that sized overlaps with the pixel centre.
 \begin{figure}
@@
 Another method, is to improve the prediction of the medium resolution itself and thus reduce the need for a computation in higher resolution.
-\subsubsection{Compare least cost paths, for overlay of both rasterizations}
+\subsubsection{Compare least cost paths, for overlay of both rasterisations}
 
@@
 Table~\ref{tab:4} shows, that distance decreases with the higher ratio from 1:1, 2:1, 4:1 and than increases with higher ratio 8:1, 16:1 and so on.
-In addition, the distances to the original paths from the all touched false and all touched true raster change with the applied ratio of the overlayed raster.
+In addition, the distances to the original paths from the all touched false and all touched true raster change with the applied ratio of the superimposed raster.
 So that, a lower ratio increases the similarity to the path of the all touched true raster, the similarity to the paths from the all touched false raster increases with the ratio.
@@
 So that a nuance, can cause a detour.
-When this behavior occurs, the polygon can not grantee to include the least cost path.
-Therefore this polygon should be overlayed with a polygon around the shortest path.
+When this behaviour occurs, the polygon can not grantee to include the least cost path.
+Therefore this polygon should be overlapped with a polygon around the shortest path.
 

### 2023-03-08 paper/results.tex
grammar + change to en_GB

@@
 It is decided to use the maximum value, for any place in the raster, that is covered by several rules at the same time.
-When the resolution that is use for the rasterization is smaller than the object size, than the effect of the rasterization with all touched True or False is limited.
+When the resolution that is use for the rasterisation is smaller than the object size, than the effect of the rasterisation with all touched True or False is limited.
 For all touched True any part of the pixel that is covered by the object, will attribute that whole pixel to the object.
@@
 \end{figure}
-Thus, with having larger areas and more objects, with higher costs, all touched True rasterization might more likely lead to longer roots and more likely in blocking the  direct spatial path.
+Thus, with having larger areas and more objects, with higher costs, all touched True rasterisation might more likely lead to longer roots and more likely in blocking the  direct spatial path.
 
 \subsection{Least Cost Paths}\label{subsec:least-cost-paths}
-For each resolution the costs paths were estimated for the rasterization with setting all\_touched False
+For each resolution the costs paths were estimated for the rasterisation with setting all\_touched False
 and all\_touched True.
@@
 
-Comparing the all\_touched True rasterization and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterization at the
+Comparing the all\_touched True rasterisation and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterisation at the
 different resolutions the later paths are more similar.
@@
 percentage of which costs levels are around the path.
-When using all\_touched True rasterization with higher resolution the tendency is to use a higher percentage of the
+When using all\_touched True rasterisation with higher resolution the tendency is to use a higher percentage of the
 \textit{Preferential Level} and less the\textit{NoRestriction} Level.
@@
 
-As all touched true rasterization overestimates and all touched true underestimates the real costs.
+As all touched true rasterisation overestimates and all touched true underestimates the real costs.
 A weighted mean will describe the real costs more precise.

### 2023-03-11 paper/discussion.tex
style, grammar format, figures

--- 2023-03-07 paper/discussion.tex
@@
-% !TeX spellcheck = en_US
-As shown, the need for computation time increases with the power of two with the resolution. 
-Similarly, the usage of main memory increases. 
-Which in tern limits the processable data points and resolution.
-On the error only scales linear with the resolution. 
-Hence, there is a diminishing return of smaller errors, compared to compute time and resources used.
+% !TeX spellcheck = en_GB
+In theory, the need for computing time increases with the resolution as power of two. 
+Similarly, the use of main memory increases. 
+This again limits the number data points, that can be processed and resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls have been used for the aggregation of higher resolution.
 
-Therefore, this worked tries a) to reduce the needed to compute power and b) decrease the deviation for a given resolution.
+On the other hand, the similarity to higher resolutions only scales linear with the resolution. 
+Thus, there is a diminishing return of smaller errors, compared to compute time and resources used.
 
-For b), two methods have been used to increase the similarity of the paths computed from medium resolution raster, to the path of the highest resolution raster.
-These methods can be used as surrogates for the more complex calculation of the least cost path with the higher resolution.
-In the first method a bi-linear down sampling of the higher resolution raster has been applied and in the second method, the all touched false and true raster where averaged in different ratios, to compute the optimum mixed raster, for the used costs.
-While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the down sampling method is more
-simple and does not need to be optimized for the given costs.
-Hence, it is easier and widely applicable.
+Therefore, this work attempts a) to reduce the computation power needed and b) reduce the deviation for a given resolution compared to a higher resolution raster.
 
-For a), to reduce the computational complexity, the calculation of the medium resolution raster has been applied, to reduce the search space of the aggregation.
-While this method reduces the calculation time and usage of main memory, the backtracking part stays unchanged.
+For a), to reduce the computational complexity, clipping of the high resolution has been applied, to reduce the search space of the aggregation.
+While this method reduces the computation time and memory usage, the backtracking part stays unchanged.
 
-As the original least cost path QGIS plug-in the used algorithm stops the least aggregation of the costs after the final end point has been found.
-This early stopping might result in suboptimal paths around the end points for some edge cases, where the connection via another neighbor might be more optimal.
+For b), two methods have been used to increase the similarity of the paths computed from the medium resolution raster, to the path of the highest resolution raster.
+These methods are used as surrogates for the more complex calculation of the Least Cost Path with the higher resolution.
+In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all touched false and true raster where averaged in different ratios, to compute the optimum weighted cost raster.
+While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
+This disadvantage could be reduced by normalising the costs.
 
-The set of rules that are used create the cost raster, includes a rule for creating a buffer around buildings  which is set to the level \textit{prohibited} areas.
-The resolution of the medium level raster needs to be high enough to crudely show every detail, Hence at least in the magnitude of the minimum object size plus twice its buffer, in this example in the range of at least 25~m or 10~m to include the streets in all touched false raster.
-Other details as rivers and houses, ae already includes in lower resolution raster, because of the buffering.
-Although the overall result can change, due super position of objects and there buffer.
-Nevertheless, does higher resolution resemble, the geography more precise, but in addition show paths that are too small, to be taken.
-On the other hand, as the run from the starting point to the end point at a resolution of 2.5~m shows, new ways, that could not be reached before, do hint inconsistencies of the rules that construct the cost raster.
+This early stopping may result in suboptimal paths around the end points for some edge cases, where the connection via another neighbour might be more optimal.
 
-Object and there buffer might leave a  path that, is smaller than 10~m resolution.
-So a second rule to find a good resolution, that it should be in the magnitude of the size of the structure  to be build.
-While the used data set for buildings, also includes power poles, the area of the level \textit{preferential} power grid, includes isles of high costs. 
-In further works, these should be excluded.
+The set of rules that are used create the cost raster, includes a rule to create buffer around buildings which is set to the level \textit{prohibited} areas.
+In all touched false rasters.
+The resolution of the medium level raster needs to be high enough to show every detail.
+At least in the magnitude of the minimum object size plus twice its buffer.
+This is true for the 10~m resolution raster and less true for the 25~m resolution raster that misses some details for roads for all touched false raster.
+Other details such as rivers and houses, are already included in the lower resolution raster, due to larger buffers.
+The Least Cost Path algorithm searches for an optimal path as a line.
+As lines do not have a width, the route found might contain bottlenecks, that have a smaller width than the object that should be placed there.
+Therefore, the used resolution should not be smaller than, the width of the object that should be placed.
+This can be avoided by downsampling, but by weighting the medium resolution all touched true or false rasters.
 
-In this work, the effect of computation costs and deviation of the results, is examined for a very limited set of points.
-Also, only the costs of finding the least cost path from one starting point, to a single end point has been considered.
-When, multiple endpoints are used, the computation cost for the aggregated cost raster has only to computed once.
-On the other hand the effect early stopping is not as dramatic, because the algorithm would only stop early for the last end point with the highest cost.
-When calculating several paths from one raster, the benefits of chapter~\ref{subsec:faster-processing-of-the-cost-path-algorithm} is reduced.
-Especially, the pre-calculation on lower resolution raster and buffering of the resulting paths as a restricted search area decreases in its effectivity.
-As the number of paths rises, fewer of points in higher resolution cost raster will be deselected for the final estimation of the least cost path.
-The least cost algorithm does only select the most cost-effective path.
-Therefore, paths of similar, but slightly worse cost are stay unknown.
-An end-user might be interested in selecting a path from a set of possible paths of similar aggregated costs and apply an adapted set evaluation criteria.
-This can be achieved a by adapting the backtracking, to get an array of paths or by applying perturbation on the costs.
+This paper examines the effect of computational costs and deviation of the results, is examined for a very limited set of points.
+Also, only the cost of finding the Least Cost Path from a single start point, to a single end point has been considered.
+If multiple endpoints are used, the computational cost for the aggregated cost raster has only to be paid once.
 
-In this work, the single cost raster layers aggregated with the maximum function. 
-Another possible aggregation function is the sum.
-Each aggregation function can be justified, by a different interpretation of the costs and its scale.
-The reasoning of selection the maximum function is, limiting the maximum costs.
-When the \textit{prohibited} level is used as the highest level.
-Then higher levels are \textit{prohibited} within the concept.
-This an area even more \textit{prohibited}, when it is included in two different rules of the level?
-On the other side, This aggregation is unable to differentiate between these nuances, as  the sum function is.
-This way, one can differential between, different sublevels.
-One consequence is, that a small mediation between high and low single cost raster, in the final cost raster is possible.
+If multiple paths are calculated from a single raster, the speed-up benefit is reduced.
+Especially, pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective.
+As the number of paths increases, fewer pixels are clipped.
+The Least Cost Path algorithm does only select the single most cost-effective path.
+Therefore, paths of similar, but slightly higher costs remain unknown.
+In additional, slight variations on the costs rasters can lead to very different paths, although the costs will not change much.
+An end-user may be interested in selecting a path from a set of similar aggregated costs and applying their own evaluation criteria.
+This can be achieved a by adjusting the backtracking and return polygons, or by applying perturbation on the costs.
 
-The fact, that all touched False raster lead to more similar results compared to high resolution raster, is due to the fact, that the default level is relatively low.
-In cause the default level would rise, the effect probably would flip for low resolution raster.
-For high resolution raster, this effect would still hold true, because the fact, that the raster pixel center is used for sampling, reflects the original geometry better.
+In this work, the intermediate cost raster layers are aggregated using the maximum function. 
+Another possible aggregation function is the sum or average.
+Each aggregation function can be justified, by a different interpretation of the cost and its scale.
 
-Interpretation: Interpretation the Cost Paths are much nearer smaller, than 100~m!
-No problem because buffering is included in the cost raster.
-Not algorithmic speed-ups as pre-compilation, JIT, vector units, multi-core, weight range
+When the \textit{prohibited} level is used as the highest level, then summing the two highest levels would result in a new highest level. 
+Also, the maximum function does not interfere with the nodata value. 
+Although this can be done by a nansum- / nanmean-function, if the nodata value is set to \textit{not a number} during the aggregation.
+The disadvantage of aggregation with the maximum function is, that this aggregation is unable to distinguish between nuances of different overlapping intermediate costs.
+On sum or average aggregated rasters, one can distinguish between, different sublevels.
+
+The fact that all touched false rasters produce more similar results than high resolution raster, is probably due to the fact, that the default level is relatively low.
+As the default level increases, the effect would probably be reduced for low resolution rasters.
+For high resolution raster, the effect would still be present, because the fact, that the raster pixel centre is used for sampling, reflects the original geometry better.
+
+This effect of the similar aggregated costs per resolution could also be seen in the test paths, even when the paths varied greatly with the change in resolution or algorithm.
+This could be an indication of an even spatial distribution of the costs.
+
+The all touched true cost raster shows every detail, but the sampling with all touched true increases the size of the features.
+The fact that the aggregated costs per resolution for all touched true rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the values of the costs are not uniformly distributed, but that the high costs are much more frequent, because the costs are exponential scaled.

### 2023-03-11 paper/methods.tex
grammar + change to en_GB
style, grammar format, figures

@@
 
-The nationwide land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
+The national land coverage (ATKIS) with a scale of 1:250000 are provided by the Federal Agency for Cartography
 and Geodesy~\cite{noauthor_digitales_2021}.
@@
 The version with any part overlapped is called all touched True.
-The version where a overlapped pixel centre is require is called all touched False.
+The version where an overlapped pixel centre is require is called all touched False.
 All touched False is considered the default (see figure~\ref{fig:alltouched}).
@@
 		\begin{tikzpicture}
-			\node[draw] at (0,6.5) {\huge all touched false};
+			\node[draw] at (0,6.5) {\Huge all\_touched=false};
 			\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
-			\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
+			\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.3,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
 
-			\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);
+			\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\Huge rasterisation} (8.5, 2);
 
@@
 		\begin{tikzpicture}
-			\node[draw] at (0,6.5) {\huge all touched true};
+			\node[draw] at (0,6.5) {\Huge all\_touched=true};
 			\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
-			\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.2,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
+			\draw[fill=blue, thick] (-0.2, 0.2)--(-0.6, 3.1)--(-0.3,4.1)--(1.2,3.4)--(2.8,3.6)--(3.2, 2.6)--(3.4, 2.1)--(1.8,1.3)--(1.2, 0.1)--cycle;
 
-			\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\huge rasterization} (8.5, 2);
+			\draw[ultra thick,->, red] (6.5, 2) --node[above=1mm] {\Huge rasterisation} (8.5, 2);
 

### 2023-03-11 paper/paper-CableRoute.tex
style, grammar format, figures

@@
 \newacronym{wps}{wps}{Web Processing Service}
-\newacronym{res}{res}{resolution}
-\newacronym{r}{r}{ratio}
-\newacronym{al}{al}{all touched}
-\newacronym{t}{t}{True}
-\newacronym{f}{f}{False}
-\newacronym{agg.}{agg.}{aggregated}
-\newacronym{rel}{rel}{relative}
-\newacronym{corr}{corr}{corrected}
 \newacronym{MCDA}{MCDA}{multi-criteria decision analysis}
@@
 
-	\received{20 February 2007}
-	\received[revised]{12 March 2009}
-	\received[accepted]{5 June 2009}
+	\received{14 March 2023}
+	%%\received[revised]{12 March 2009}
+	%%\received[accepted]{5 June 2009}
 
@@
 
-	Other methods such as the \acrfull{MCDA}  consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).
+	Other methods such as \acrfull{MCDA} consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).
 
@@
 
-	Older models are e.g.\ are optimizing the well fare, economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance more and more play an important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
+	The different models of the power system, that are used in the expert system range from not modelling the grid at all to models that include Kirchhoffs laws~\cite{bertsch_participatory_2016}.
+	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
+	Some models consider not only the grid of one nation, but consider neighbouring states or the whole European power system~\cite{dietrich_will_2010}.	
+	Older national power grid models are e.g.\ are optimising the well fare, but economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an increasingly important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
 
-	In contrast to a GIS analysis, these further factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
+	In contrast to a GIS analysis, these other factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
+	Stakeholders as decision makers can be included and combined with an expert system~\cite{bertsch_participatory_2016}.
 
-	Stakeholder as decision maker can be included and combined with an expert system~\cite{bertsch_participatory_2016}.
-
-	The uncertainties connected with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are examined by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
-	Both inter and intra criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.
-
-	The different models of the power system, that are used in the expert system span from not modelling the grid at all to models that includes Kirchhoffs laws~\cite{bertsch_participatory_2016}.
-	The calculation of a power system is complex, as changing one edge will change the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
-	Some models include not only the grid of a nation alone, but consider neighbouring states or the overall European power system~\cite{dietrich_will_2010}.
+	The uncertainties associated with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are investigated by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
+	Both inter- and intra-criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.
 
 	\section{Conclusion}\label{sec:conclusion}
-	The cost of computation of the least cost path scales with the square of the resolution.
+	The cost of computing of the Least Cost Path scales with the square of the resolution.
 	The difference between the aggregated costs of the paths per resolution computed from the all touched false raster and the all touched true raster only shows a linear decrease.
-	Hence, the gain in precision per compute time decreases.
-	The presented methods tried to circumvent this bottleneck.
-	When using the researched strategies, on medium resolution raster, the computation for high resolution results might successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
-	The paths might change very strongly even with small changes in total costs.
-	Hence to develop an alternative backtracking algorithm that, that generates a set of possible paths within a corridor of costs might be a good strategy, to offer alternative routes and show variations between these routes at the same time.
-	An alternative method for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
-	This in turn increases the computation time again.
+	Therefore, the gain in accuracy per compute time decreases.
+	The presented methods have attempted to circumvent this bottleneck.
+	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
+	The paths can change significantly even for a small change in total cost.
+	Therefore, to develop an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy, to offer alternative paths and at the same time show variations between these paths.
+	An alternative for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
+	This again increases the computation time.
 
-	In an actual search for least cost path a survey might be used as method to estimate the costs.
-	When the sample size of the survey is large enough, the
-	weights might take local differences relative acceptance into consideration.
+	In an actual search for Least Cost Path search, a survey could be used as a method to estimate the costs.
+	If the sample size of the survey is large enough, the
+	weights could take  into account local differences in relative acceptance.
 
-	Only changes in the algorithm have been applied in the search for a speed-up.
-	There a still other methods as just-in-time compilation.
+	Only changes in the algorithm have been applied in the search for acceleration.
+	Therefore, other methods such as just-in-time compilation have not been tested.
 

### 2023-03-11 paper/results.tex
style, grammar format, figures

@@
 
-In this chapter we want to show the different cost raster, that were created from the very same set of layers,
-but computed for different resolutions.
-From this different raster the cost paths are calculated and compared.
-In the last step the raster with lower resolution were used to calculate in a way, that they shall result in
-similar paths as if the paths were computed from a high resolution raster.
+In this chapter we want to show the different cost rasters, that were created from the same set of layers at different resolutions.
+The Least Cost Path is estimated from this set of rasters.
+In the last step the Least Cost Path is computed from the lower resolution rasters and compared with the Least Cost Paths computed from a high resolution raster.
 
@@
 
-The cost raster contains all the costs as weights for the geographic region of the study area.
-The regions outside the study area are given, a no-data value and will not be use for the calculation of the least cost path.
-It is decided to use the maximum value, for any place in the raster, that is covered by several rules at the same time.
-When the resolution that is use for the rasterisation is smaller than the object size, than the effect of the rasterisation with all touched True or False is limited.
+The cost raster contains all the costs for the geographical region of the study area.
+%%The surrounding of the study area, uses a no-data value and will not be use for the calculation of the Least Cost Path.
+The different cost layers in the study area are aggregated by the maximum function.
+If the resolution is higher than the object size, then the effect of the rasterisation with all touched True or False is limited.
 For all touched True any part of the pixel that is covered by the object, will attribute that whole pixel to the object.
-Thus, the object appears to be halve a pixel size larger in all directs.
-As can be seen in figure~\ref{fig:costs_5m} that shows a detailed view of the costs for village of Beverstedt.
-Due to the maximum aggregation of the costs, the average cost of the raster of all touched true will be over estimated.
-All touched false will be a better description of the real size of the object.
+This makes the object appear a halve a pixel size larger in all directs.
+As can be seen in figure~\ref{fig:costs_5m}, which shows a detailed view of the costs for the village of Beverstedt.
+ %% Due to the maximum aggregation of the costs, the average cost of the raster of all touched true will be over estimated.
+The rasterisation with all touched false will be a better description of the real size of the object, for high resolution.
 \begin{figure}
@@
 
-	\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alT.png} }}%
+	\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alT_v2_small.png} }}%
 	\qquad
-	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alF.png} }}%
+	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alF_v2_small.png} }}%
 	\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 5~m.}
@@
 
-In contrast when the resolution the is much less, than the size of the object the described behavior changes.
-On one hand, while the area, of the pixel is increased for all touched True, also the area for which the cost is overestimated increases.
-On the other hand while the pixel size for all touched  False increases, not only the over estimated area increases for that pixel as describes for all touched true, but in addition the object is less probable situated in the centre of the pixel.
-The consequence of the pixel is, that will decreasing resolution that object is not included in the rasterisation.
-Hence, hence for all touched false lower resolution  leads ot a loss of information.
-Because the default is relative small compared the over effect is a underestimation of the costs.
-The figure~\ref{fig:costs_100m} shows for the resolution of 100 m, that while larger are still included in the map they appear to be much larger.
-On the contrary smaller objects might not be included.
-Objects that are small on in one dimension as streets will be included in a stochastic manner.
-Described by the likely of that a object of that sized overlaps with the pixel centre.
+In contrast, if the resolution is smaller, than the object-size this behaviour changes.
+In general, while the area, of the pixel increases for all touched True, the area for which the cost is overestimated also increases.
+As the pixel size for all touched  False increases, not only does the overestimated area increases, but in addition the object is less probable situated in the centre of the pixel. 
+Therefore, object is sampled at random.
+Hence, all touched false leads to a loss of information for smaller objects.
+\todo{Put this part into Discussion?} Since the default cost is much smaller than the average cost, this method underestimates the cost.
+The figure~\ref{fig:costs_100m} shows, that for the resolution of 100 m, that larger objects are still included in the map, but appear to be larger.
+But smaller objects, such as roads, are not fully or randomly included.
 \begin{figure}
@@
 
-	\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alT.png} }}%
+	\subfloat[\centering All touched: True.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alT_v2_small.png} }}%
 	\qquad
-	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alF.png} }}%
+	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alF_v2_small.png} }}%
 	\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 100~m.}
@@
 \end{figure}
-Thus, with having larger areas and more objects, with higher costs, all touched True rasterisation might more likely lead to longer roots and more likely in blocking the  direct spatial path.
+\todo{Put this part into Discussion?}Thus, with having larger areas and more objects, with higher costs, all touched True rasterisation might more likely lead to longer roots and more likely to block the  direct spatial path.
 
 \subsection{Least Cost Paths}\label{subsec:least-cost-paths}
-For each resolution the costs paths were estimated for the rasterisation with setting all\_touched False
-and all\_touched True.
+For each resolution the Least Cost Paths were estimated from the all\_touched False and all\_touched True rasters.
 
-The hypothetical path starts at a transformer about 6~km north the container terminal Bremerhaven and then end at a transformer in the south east of the Osterholz county.
+For the study, a start point were chosen at a transformer about 6~km north of the container terminal Bremerhaven and an end point at a transformer in the southeast of the Osterholz county.
 
-The distance of both paths is calculated with the mean minimum distance.
-For every vertex $P_i$ in the path $L_1$ the minimum between the vertex and the path $L_2$
-is computed and afterwards the minimum distances are averaged (see equation~\ref{eq:1}).
+The distance of Least Cost Paths of the all touched false to the paths of the all touched true raster is calculated by the mean minimum distance.
+For each vertex $P_i$ in the path $L_1$ the minimum distance between the vertex and the path $L_2$
+is calculated and then the minimum distances are averaged (see equation~\ref{eq:1}).
 \begin{equation}
 	\label{eq:1}
-	d_{mean} = \frac{1}{|L_1|} \sum_{i=1}^{n} d_{min}(p_i, L_2) \bigg\vert p_i \in L_1
+	d_{mean} = \frac{1}{|L_1|} \sum_{i=1}^{n} d_{min}(p_i, L_2) \Bigr\vert p_i \in L_1
 \end{equation}
-This equation is used, to measure the extent of similarity between the paths.
+This equation is used, to measure the degree of similarity between the paths.
 
-\begin{equation}
-\label{eq:2}
-d_{max} = max(\sum_{i=1}^{n} d_{min}(p_i, L_2)) \bigg\vert p_i \in L_1
-\end{equation}
-
-Hence, the mean minimum euclidean distance between the two paths can used to compute
-the similarity.
-As different table~\ref{tab:2} shows the distance between the two paths decreases
+Table~\ref{tab:2} shows the distance between the two paths decreases
 with increasing resolution.
@@
 
-In contrast the largest distance between the paths (see equation~\ref{eq:2}) will be used, to estimate the minimum distance to the areas that still should be examined to grantee a, that at least cost path found in this limited space is still likely to be optimal.
-
-At the same time the differences in the aggregated costs stay almost constant.
-When normalize the aggregated costs by the resolution.
-On one hand it can be seen, that the all\_touched False under estimate the costs and that this tendency scales
-linear with the resolution.
-On the other hand, the all\_touched True least cost over estimates the aggregated costs on a linear scale of
+At the same time, the differences in the aggregated costs per resolution remain almost constant.
+On one hand it can be seen, that the all\_touched False underestimates the costs and that this tendency scales
+linearly with the resolution.
+On the other hand, the all\_touched True least cost overestimates the aggregated costs on a linear scale of
 the resolution.
-Therefore, the difference for the normalized least cost paths is reduced on scale by the resolution.
 
@@
 
-	\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m.pdf} }}%
+	\subfloat[\centering resolution of 5 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_5m_v2_small.png} }}%
 	\qquad
-	\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m.pdf} }}%
+	\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m_v2_small.png} }}%
 	\caption{Figures of the least cost paths.
@@
 
-	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F.pdf} }}%
+	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F_v2_small.png} }}%
 	\qquad
-	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T.pdf} }}%
+	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T_v2_small.png} }}%
 
@@
 \begin{table*}[t]
-	\caption{Least cost paths as length for the different \acrfull{res} of the raster, including the mean minimum distance and the maximum minimum distance and the \acrfull{agg.} costs.
-From the \acrshort{agg.} costs the differences of the \acrshort{agg.} costs and the \acrfull{corr} \acrshort{agg.} by resolution are given.}
+	\caption{Least cost paths as length for the different resolution of the raster, including the mean minimum distance and the maximum minimum distance and the agg.
+costs.
+From the agg.
+costs the differences of the agg.
+costs and the agg.
+costs per resolution are given.}
 	\label{tab:2}
@@
 	\begin{tabular}{ r  r  r  r  r  r  r  r  r  r}
-		\acrshort{res} /m & $l_{\acrshort{al}=\acrshort{f}} /m$ & $l_{\acrshort{al}=\acrshort{t}} /m$ & $d_{mean}$ /m & $d_{max}/m$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{f}}$ & \acrshort{agg.}  $ cost_{\acrshort{al}=\acrshort{t}}$ &  $\Delta $ costs & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{f}}$ & \acrshort{corr} \acrshort{agg.} $costs_{\acrshort{al}=\acrshort{t}} $ \\
+		res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m & $d_{max}/m$ & agg.  $ cost_{al=f}$ & agg. $ cost_{al=t}$ &  $\Delta $ costs & agg. $costs_{al=f} /m$ & agg. $costs_{al=t} /m$ \\
 		\hline
-		5 	& 76136.27	& 78002.00 &  126.04 & 1065.00 & 18665.923 & 19616.756 & -850.00 & 93329.60 &  97584.77 \\
-		10 	& 75430.10 	& 77936.57 &  277.92 & 1590.00 &  8931.245 &  9731.175 & -799.95 & 89312.45 &  97311.75 \\
-		25 	& 75422.85 	& 78422.85 &  313.75 & 1621.15 &  3354.869 &  3872.656 & -517.78 & 83871.73 &  96816.40 \\
-		50 	& 76135.02	& 70619,95 & 1140.01 & 4950.00 &  1409.023 &  2300.073 & -891.05 & 70451.15 & 115003.65 \\
-		100 & 76283.80	& 74120.73 & 1946.41 & 6016.64 &   640.516 &  1572.268 & -931.70 & 64051.60 & 167226.80 \\
+		5 	& 76136.3	& 78002.0 &  126.0 & 1065.0 & 18665.9 & 19616.8 & -850.00 & 93329.6 &  97584.8 \\
+		10 	& 75430.1 	& 77936.6 &  277.9 & 1590.0 &  8931.2 &  9731.2 & -799.95 & 89312.5 &  97311.8 \\
+		25 	& 75422.9 	& 78422.9 &  313.8 & 1621.2 &  3354.9 &  3872.7 & -517.78 & 83871.7 &  96816.4 \\
+		50 	& 76135.0	& 70620.0 & 1140.0 & 4950.0 &  1409.0 &  2300.1 & -891.05 & 70451.2 & 115003.7 \\
+		100 & 76283.8	& 74120.7 & 1946.4 & 6016.6 &   640.5 &  1572.3 & -931.70 & 64051.6 & 167226.8 \\
 
@@
 
-Comparing the all\_touched True rasterisation and all\_touched False for the same resolution in contrast with the paths of all\_touched False rasterisation at the
-different resolutions the later paths are more similar.
-The mean average distance between the 100 m resolution run and the 5 m resolution run is 257.97 m.
+When estimating distance between the Least Costs Paths from all\_touched True rasterisation and all\_touched False at the same resolution 
+the mean minimum distance between the 100~m resolution paths is 1946.41~m and between the 5~m resolution paths is 126.04~m.
+The similarity for the all touched False paths is higher, than for the all touched true raster.
+The distance for the 100~m path to the 5~m resolution path is 243.42~m for all touched false and 2109.44~m for all touched true.
 
-The similarity of all all\_touched False runs is higher than, the similar between the all\_touched False and all\_touched True runs with the same resolution, except for the highest resolution.
+When comparing the similarity between the all touched false paths themselves and the similarity between the all touched false rasters to the all touched true rasters of the same resolution: The similarity in between the all all\_touched False paths is higher than, than for most paths of same resolution.
+Namely except for the highest resolution.
 
-Hence, in an overall perspective paths of the all\_touched False runs stay in a
-similar region, while paths of the all\_touched True coverage strongly to the paths all\_touched False runs.
-This behaviour is depicted in figure~\ref{fig:paths_alltouched}.
-On a more detailed level, it can be seen, that also the Paths of all\_touched 	converges the all\_touched True path.
-But the extent is smaller.
-The length of the the paths only differs to a maximum of about 10 \%.
-The least path distance for higher resolutions can be lower, because new paths, between regions that are forbidden
-in higher resolution can open.
-On the other side the length of the paths can increase, because with higher resolution more vertices will be used.
+This behaviour is shown in figure~\ref{fig:paths_alltouched}.
+On a more detailed level, it can be seen, that the paths of all\_touched False also converge directly to the all touched True paths, but the extent is smaller.
+The length of the paths only differs by a maximum of about 10 \%.
+On the other hand, side the length of the paths can increase, because with more vertices are used with higher resolution.
 
-The zonal stat (see table~\ref{tab:3}) for a buffer of 100 m (5 m) around the path has been used, to estimate the
-percentage of which costs levels are around the path.
-When using all\_touched True rasterisation with higher resolution the tendency is to use a higher percentage of the
-\textit{Preferential Level} and less the\textit{NoRestriction} Level.
-The ratio of the 100 m buffered least cost path, strongly shifted  to Levels lower costs.
+The zonal stat (see table~\ref{tab:3}) for a buffer of 100~m (5~m) around the path has been used, to estimate the
+percentage of each costs levels around the path.
+When using all\_touched True rasterisation at higher resolution, the tendency is to use a higher percentage of the
+\textit{Preferential Level} and less of the\textit{NoRestriction} Level.
+The ratio of the 100~m buffered Least Cost Path, strongly shifted  to Levels lower costs.
 
@@
 
-In theory, the execution time should increase with the square of the resolution, because higher resolutions result in a higher number of pixels and thus data points the aggregated costs needs to be calculated for.
+In theory, the execution time increases with the square of the resolution, because higher resolutions result in a higher number of pixels and thus data points the aggregated costs needs to be calculated for.
 A full logarithmic fit for several repetitions of the execution shows, that the execution time scales with power of $2.1997  \pm 0.007$ of the inverse resolution.
-With the caveat of a low number of samples this can equally be success fully be fitted to a second degree polynomial of the inverse resolution with a $r^2$ for the test set of 0.99 and a the squared inverse of the resolution with a $r^2$ for the test set of 0.99.
-Hence, the order of magnitude
 
@@
 The aggregation of the costs and the back tracking of the least cost to find the path.
-% \todo{back tacking scales with number of points.}
 
@@
 \begin{table*}[t]
-	\caption{\acrfull{r} of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
+	\caption{resolution (r) of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
 	\label{tab:3}
@@
 
-The final least costs paths should between the least cost paths of the lower resolution, with a tendency to be nearer to the paths resulted by the all touched false rasterization.
-The first step is to optimize the calculation speed is, only to calculated the least cost paths for this smaller area.
+The first step is to optimise the computational speed, by a reduced area.
 Another method, is to improve the prediction of the medium resolution itself and thus reduce the need for a computation in higher resolution.
+
 \subsubsection{Compare least cost paths, for overlay of both rasterisations}
 
-As all touched true rasterisation overestimates and all touched true underestimates the real costs.
-A weighted mean will describe the real costs more precise.
-As present work indicates, that the weight should favor the all touched false raster.
-The best weight should be the percentage of the pixel, which is covered by the object, but this can not be computed in this work.
-Thus, the best weight has to be searched.
-An alternative might be to compute the cost raster in higher resolution, but than to reproject the to a smaller resolution with a using a (linear) interpolation of the of the weights.
+For the example paths shown, all touched true rasterisation overestimates the true costs and all touched true underestimates them.
 
-While the aggregated thus can be speed up.
+A weighted average of the costs could therefore be a more accurate measure and make estimated medium resolution Least Cost Path more similar, to high resolution paths.
+As all example show, the weighting should favour the all touched false raster.
+\todo{move to discussion/conclussion:}The best weight should be the percentage of the pixels, which is covered by the object, but this can not be calculated in this work.
+An alternative would be to compute the cost raster at a high resolution and reproject them to a medium resolution by a (linear) interpolation of the weights.
+
+\todo{move to planning:}This will speed-up the aggregation.
 The time needed for the back tracking stays unchanged.
 
-The optimal ratio of overlaying all touched false and all touched true cost raster for 10~m resolution is estimated via the resulting least cost path. 
-The mean distance of least cost paths for different ratio has been estimated to the path from the all touched false raster of the next higher (5~m) resolution. 
-Table~\ref{tab:4} shows, that distance decreases with the higher ratio from 1:1, 2:1, 4:1 and than increases with higher ratio 8:1, 16:1 and so on.
-In addition, the distances to the original paths from the all touched false and all touched true raster change with the applied ratio of the superimposed raster.
-So that, a lower ratio increases the similarity to the path of the all touched true raster, the similarity to the paths from the all touched false raster increases with the ratio.
+The optimal ratio of overlaying all touched false and all touched true cost raster for 10~m resolution is estimated via similarity of the resulting Least Cost Path to the path of the original high Resolution raster.
+The mean distance of Least Cost Paths for different ratio is estimated to the path from the all touched false raster of the higher (5~m) resolution. 
+Table~\ref{tab:4} shows, that the distance decreases, with increasing ratio (1:1, 2:1, 4:1) and after this optimum is reached, increases with increasing ratio (8:1, 16:1 and so on).
+Comparing  the similarity of the paths from the rasters of the different ratios to normal paths with 10~m resolution, paths with a higher ratio of all touched true is nearer to the all touched true paths.
+Paths with a ratio in favour of all touched false are much closer to the all touched false paths.
 
@@
 
-		  1:1  &    119.6 &  285.50 &  47.21\\
-		  2:1  &    97.11 &  263.51 &  74.19\\
-		  4:1  &    40.13 &  206.38 & 100.16\\
-		  8:1  &    41.73 &  169.02 & 137.34\\
-		 16:1  &    56.69 &  153.32 & 152.71\\
-		 32:1  &    56.69 &  145.56 & 162.09\\
-		 64:1  &   163.48 &   10.61 & 272.36\\
-	  % 128:1  &   163.82 &   10.29 & 276.71\\
+		  1:1  &    119.6 &  285.5 &  47.2\\
+		  2:1  &    97.1 &  263.5 &  74.2\\
+		  4:1  &    40.1 &  206.4 & 100.2\\
+		  8:1  &    41.7 &  169.0 & 137.3\\
+		 16:1  &    56.7 &  153.3 & 152.72\\
+		 32:1  &    56.7 &  145.6 & 162.1\\
+		 64:1  &   163.5 &   10.6 & 272.4\\
+	  % 128:1  &   163.80 &   10.3 & 276.7\\
 
@@
 
-As an alternative to the super position of the all touched true and false raster for the same resolution, the all touched false raster was down sampled to 10~m, 25~m, 50~m and 100~m with bi-linear interpolation.
-With this method smaller structures still can be seen in the cost raster, although the resolution is reduced.
-The distances of the paths that are computed from the bi-linear down sampled  5~m resolution raster (all touched false) is from table~\ref{tab:5} shows, that a low distance to path of the 5~m resolution raster was only obtained when down sampling to a resolution not to low.
-Only when down sampling to a resolution of 10~m, is the computed path considerable close to the path from the higher resolution raster.
-Is holds true, but to a much reduced extend to the path computed from the raster, that was down sampled to a resolution of 25~m.
+As an alternative to the superposition of the all touched true and false rasters for the same resolution, the all touched false raster is down-sampling to 10~m, 25~m, 50~m and 100~m (with bi-linear) interpolation.
+With this method, smaller structures can still be fully seen in the cost raster, although the resolution is reduced.
+The distances of the paths that are computed from the bi-linear down sampled raster to the path of the original 5~m resolution (all touched false) shows (see table~\ref{tab:5}), that only down-sampling to a resolution of 10~m, produces a path that is relatively close the high resolution path.
+
 The opposite is true for the lower resolution raster which is more similar to paths computed from the all touched true cost raster.
-For every path from a down sampled raster is more similar to a path computed from an all touched true raster, than an all touched false raster, although the all touched false raster of the 5 m resolution was used for down sampling.
+Every path from a down sampled raster is more similar to a path computed from an all touched true raster, than an all touched false raster, although the all touched false raster of the 5 m resolution was used for down sampling.\todo{interpretation all touched true and downsampling, both show all details, but downsampling, more similar to original geometry}
 
@@
 
-		10  & 75980.62 &   59.34 &  219.35 & 143.63\\
-		25  & 70205.27 &  385.79 &  558.12 & 432.77\\
-		50  & 69217.86 &  730.82 &  693.42 & 255.70\\
-		100 & 66667.87 &  1681.3 & 1605.63 & 400.55\\
+		10  & 75980.6 &    59.3 &  219.4 & 143.6\\
+		25  & 70205.3 &   385.8 &  558.1 & 432.8\\
+		50  & 69217.9 &   730.8 &  693.4 & 255.7\\
+		100 & 66667.9 &  1681.3 & 1605.6 & 400.6\\
 
@@
 
-Construct a polygon form the two lines.
-Buffer the polygon with twice the max minimum path distance.
+Construct a polygon from the two Least Cost Paths (all touched true and all touched false of the same resolution).
+Buffer the polygon with twice the maximum  minimum path distance  (see equation~\ref{eq:2}).
+\begin{equation}
+	\label{eq:2}
+	d_{max} = max(\sum_{i=1}^{n} d_{min}(p_i, L_2)) \Bigr\vert p_i \in L_1
+\end{equation}
 
-This enabled the possibility  to run a 2.5~m resolution cost raster and clip it to the extent of the polygon. 
-This clipped 2.5~m raster for all touched true did only slightly change the path. 
-While the all touched False raster, leads to a totally new segment at the end of the path. 
-Due to the small resolution, that a small path with the extent of about 2.5 m width became passable.
-This small path is a power line next to road between which build a cone in a protected landscape area.
-The street and the protected landscape area are both restricted areas, while the power line is preferential.
-This showing, that the way the cost raster is created in the first place can play a vital role, in the end result.
+This provided the possibility  to run a 2.5~m resolution cost raster and clip it to the extent of the polygon. 
+This clipped 2.5~m raster for all touched true changed the path only slightly change.
+
+The all touched False raster, on the other hand, leads to a completely new previously unused subroute at the end of the path.
+Due to the low resolution a small path became passable.
+This small path is a power line next to road between protected landscape areas.
+The road and the protected landscape area are both \textit{restricted} areas, while the power line is \textit{preferred}.
+This shows, that the way the cost raster is created in the first place can play a crucial role, in the end result.
 So that a nuance, can cause a detour.
-When this behaviour occurs, the polygon can not grantee to include the least cost path.
-Therefore this polygon should be overlapped with a polygon around the shortest path.
+When this behaviour occurs, the polygon may not include the Least Cost Path.
+\todo{move: discussion}This polygon should therefore be overlapped with a polygon around the shortest path.
 
@@
 \subsubsection{Examine the proposed solutions}
-To broaden the view and verify the current result, four different routes should be found with the presented strategies applied.
-Two paths form the starting point to two new points in the south east of the investigated area and two paths from the north, north east of the investigates area to the ending point.
+To broaden the view and verify the result, e.g.\ start and end points, four different routes should be found using the above strategies.
+Two routes should be found form the start point to two new points in the south east of the investigated area and two routes should be found from the north and north east of the study area to the end point.
 
-For three of the four routes the least cost path computation from the clipped raster, was able to compute the expected result.
-For the forth path, the
-least cost path from the 5~m resolution raster was out side of the buffer around the least cost path from the 50~m resolution raster.
-The speed up from the clipping of the higher resolution raster depends on the amount of pixels that, have been clipped. 
+For three of the four routes, the Least Cost Path estimation from the clipped raster, was able to calculated exactly the same result.
+For the fourth path, the Least Cost Path from the 5~m resolution raster was clipped out by the buffer around the 50~m resolution paths.
+The speed up from the clipping of the higher resolution raster depends on the number of pixels that, have been clipped.
+
 Bi-linear down sampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
-The resolution corrected aggregated costs of the least cost path from the down sampled raster are higher, than those, from the higher 5~m resolution raster and the medium 10~m resolution raster.
-In addition, is the distance, from this path to the path from the high resolution raster much higher, than the distance of path from the original 10~m resolution raster to the path from the 5~m resolution raster (see table~\ref{tab:6}).
-\todo{interpretation: Why result from down sampling so different? Don't know.}
+The aggregated cost per resolution of the Least Cost Path from the down-sampled raster is higher, than that from the higher resolution 5~m raster and the normal medium resolution 10~m raster.
+In addition, the distance from these paths to the high resolution path  is greater, than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
+\todo{interpretation: Why is the down-sampling result so different? Don't know.}
 
@@
 	\centering
-	\begin{tabular}{ r  r  r  r  r}
+	\begin{tabular}{ r @{\hspace*{3mm}}  r  r  r  r}
 		Route & Method & $length /m$ & $costs_{al=f}$ & $d_{mean}$ /m \\
 		\hline
-		P1-E & 5~m 			& 107889.58 & 208547.79 &        \\
-		 	 & Clipped 		& 107889.58 & 208547.79 &   0.0  \\
-		 	 & Down			&  96754.20 & 212910.95 & 628.05 \\
-		 	 & 10~m 		& 107232.94 & 203010.15 & 103.49 \\
+		P1-E & 5~m 			& 107889.6 & 208547.8 &        \\
+		 	 & Clipped 		& 107889.6 & 208547.8 &   0.0  \\
+		 	 & Down			&  96754.2 & 212911.0 & 628.1 \\
+		 	 & 10~m 		& 107232.9 & 203010.2 & 103.5 \\
 		\hline
-		P2-E & 5~m 			& 103706.41 & 155567.86 &        \\
-		 	 & Clipped 		& 103706.41 & 155567.86 &   0.00 \\
-		 	 & Down 	    &  92403.29 & 158238.58	& 639.88 \\
-		 	 & 10~m 		& 104249.94 & 149899.72 & 177.68 \\
+		P2-E & 5~m 			& 103706.4 & 155567.9 &        \\
+		 	 & Clipped 		& 103706.4 & 155567.9 &   0.0 \\
+		 	 & Down 	    &  92403.3 & 158238.6	& 639.9 \\
+		 	 & 10~m 		& 104249.9 & 149899.7 & 177.7 \\
 		\hline
-		S-P3 & 5~m 			& 102187.09 & 34503.81 	&         \\
-			 & Clipped 		&  90377.39 & 37926.14 	& 4465.37 \\
-			 & Down 	    &  94125.55 & 37574.93 	&  742.41 \\
-			 & 10~m 		& 102461.58 & 32445.96 	&   81.18 \\
+		S-P3 & 5~m 			& 102187.1 & 34503.8 	&         \\
+			 & Clipped 		&  90377.1 & 37926.1 	& 4465.4 \\
+			 & Down 	    &  94125.6 & 37574.9 	&  742.4 \\
+			 & 10~m 		& 102461.6 & 32446.0 	&   81.2 \\
 		\hline
-		S-P4 & 5~m 			& 96449.22 	& 33865.53 	&  		\\
-		 	 & Clipped 		& 96449.22 	& 33865.53 	& 0.00 	\\
-			 & Down 		& 87861.12 	& 36462.72 	& 796.40\\
-			 & 10~m 		& 96739.47 	& 31899.31	& 83.46 \\
+		S-P4 & 5~m 			& 96449.2 	& 33865.5 	&  		\\
+		 	 & Clipped 		& 96449.2 	& 33865.5 	& 0.0 	\\
+			 & Down 		& 87861.1 	& 36462.7 	& 796.4\\
+			 & 10~m 		& 96739.5 	& 31899.3	& 83.5 \\
 

### 2023-03-12 paper/methods.tex
style, grammar format, figures
moving related work into introduction
moving bits around in the chapters

@@
 % !TeX spellcheck = en_GB
+
+The least cost path algorithm is used to plan a possible route for a power line.
+The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
+The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
+This makes it possible to find routes on graphs that are not predefined, such as road networks.
+The weights of the edges are the local costs of travel from one pixel center to the neighbouring pixel center.
+The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate to transverse a given land usage.
+The Least Cost Path algorithm consists of at least two successive steps.
+1) The first step is to aggregate the costs of travelling from the starting point to a given set of end points.
+This step generates the aggregated cost raster of travelling from the starting point to any point of the cost raster.
+2) In the next step the backtracking, the route of the actual Least Cost Path is calculated.
+For each end point the path via the lowest cost neighbour is taken until the starting point is reached.
+
+Some implementations switch the role of start and ending points, so that either many start points and one single end point, or one single end point and many end points can be used.
+In some implementation there is an extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
+path to the starting point as integer values.
 
@@
 The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
-'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0.
+'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0, but refactored to optionally export the aggregated costs.
 
-The different layers of the different entities are optionally filtered, buffered and then rasterised.
-Filtering the layers of the files for special attributes enables further differentiation.
-For examples, it possible to differentiate between heath and uncultivated land in the land cover.
-Adding a buffer can be used either used to convert a line objects such as power grids and streets into a polygon with the
-correct physical width, or to add minimum distance from an existing of planed area to the new power grid.
-Each of theses rasterised layers are given a weight, or cost which expresses the cost of using land covered by this layer.
-The costs of all layers is aggregated with the maximum function.
-Thus, an area covered by several layers is uniformly used with the highest cost.
+To computed intermediate cost cost raster the different layers of the different entities are optionally filtered, buffered and then rasterised.
+Filtering the layers of the vector files for special attributes enables further differentiation.
+For examples, it is possible to differentiate between heath and uncultivated land in the land cover.
+Adding a buffer can be used either to convert a line object such as power lines and roads into a polygon (with the
+correct physical width), or to add minimum distance from an existing of planed area to the new power line.
+Each of theses intermediate rasters are given a weight (cost) which expresses the cost of using land covered by this layer.
+In the final cost raster costs of all intermediate rasters are aggregated with the maximum function.
+Thus, an area covered by several layers is uniformly associated with the highest cost.
 Any place in the study area, that is not covered by any layer and thus does not yet have a weight, is given the default cost.
 
-The costs have been grouped into different levels~(see table~\ref{tab:1}), starting from \textit{preferential} areas with
-very low costs, via \textit{no restriction}s, which is the default, used when no other layers are covering the local area,
-to \textit{restricted}, \textit{strongly restricted} and \textit{prohibited} areas with high costs.
+The costs have been grouped into five different levels~(see table~\ref{tab:1}), starting from \textit{Preferential} areas with
+very low costs, via \textit{No restriction}s, which is the default, used when no other layers are covering the local area,
+to \textit{Restricted}, \textit{Strongly Restricted} and \textit{Prohibited} areas with high costs.
 These higher costs resemble the degree to which a place with this cost should be avoided, while routing the path.
@@
 Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
-the weight that resembles the costs for these types of areas, has to be especially high.\\
+the weight that resembles the costs for these types of areas, is set especially high.\\
 
@@
 Both rasterisation techniques differ in the selection of the pixel, that describe the original polygon.
-The pixels can be selected , if either the centre of that pixel is overlapped
+The pixels can be selected, if either the centre of that pixel is overlapped
 by the geo-object, or any part of the pixel is overlaid.
-The version with any part overlapped is called all touched True.
-The version where an overlapped pixel centre is require is called all touched False.
-All touched False is considered the default (see figure~\ref{fig:alltouched}).
+Setting all touched to True, implies the version with any part of the pixel selected.
+The version where an overlapped pixel centre is require is setting the parameter all touched to False.
+All touched set to False is considered the default (see figure~\ref{fig:alltouched}).
 \begin{figure}[!ht]
@@
 		\begin{tikzpicture}
-			\node[draw] at (0,6.5) {\Huge all\_touched=false};
+			\node[draw] at (0,6.5) {\Huge all\_touched=False};
 			\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
@@
 		\begin{tikzpicture}
-			\node[draw] at (0,6.5) {\Huge all\_touched=true};
+			\node[draw] at (0,6.5) {\Huge all\_touched=True};
 			\draw[step=1cm,gray,very thin] (-1.9,-1.9) grid (5.9,5.9);
@@
 	}%
-	\caption{Graphical example for rastering a vector (left blue), to a raster (right blue) with either all-touched False (above), or all touched True (below).}
+	\caption{Graphical example for rasterise a vector (left blue), to a raster (right blue) with either all touched set to False (above), or True (below).}
 	\label{fig:alltouched}
@@
 		Prohibited 			& 500					& \makecell[lt]{Conversation areas as\\ National Parks, Buildings} \\
-		strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
+		Strongly Restricted & 10 					& \makecell[lt]{Conversation areas as Bird Reserve} \\
 		Restricted 			& 5						& \makecell[lt]{Protected Landscape Area,\\ Industrial Areas, motorway, railway} \\
@@
 
-The complete list of layers and their processing, can be found in \textbf{Supplement S1}.
+The complete list of layers and the applied processing steps, can be found in \textbf{Supplement S1}.
 
-All three steps of the generation of the least cost path: generation of the cost raster, aggregation and backtracking is shown with an example for a starting cost raster of 50~m resolution and a rasterisation with all touched False (see \ref{fig:costs2path}).
+All three steps of the generation of the least cost path: generation of the cost raster, aggregation and backtracking is shown with an example for a cost raster of 50~m resolution and all touched set to False (see \ref{fig:costs2path}).
 
 The chosen implementation applies early stopping.
-Therefore, the costs for points that are not needed to try to connect to the end point are not aggregated.
+Therefore, the costs for points that are not needed to try to connect to the end point are not aggregated (see figure \ref{fig:aggregation}).
 After finding an aggregated cost for every end point, the aggregation stops and the backtracking starts.
@@
 	\enskip
-	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.30\linewidth]{./images/AggregatedCosts_cut.png} }}%
+	\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.30\linewidth]{./images/AggregatedCosts_cut.png} \label{fig:aggregation}}}%
 	\enskip
@@
 
-	\caption{Figures of Cost raster and the resulting aggregated Costs and the least cost path for a resolution of 50~m.}
+	\caption{Figures of the cost raster and the resulting aggregated costs and the Least Cost Path for a resolution of 50~m, all touched set to False.}
 	\label{fig:costs2path}
@@
 %% Hence a point with a lower cost is evaluated earlier as a starting position to evaluate the cost of its neighbours, than a point with higher cost.
+
+For low resolution rasterisation with all touched set to True, will show every detail, but the objects are enlarged.
+When all touched is to False the object only appears, when the is situated at the pixel centre.
+Thus, this might be used as surrogate, that expresses the likelihood of the object to be sampled and correlates with the object size  compared to the pixel size.
+At high resolution the set all touched to True still overestimates the object size, but the extent is limited.
+Setting all touched to False will include all objects for high resolution.
+This setting is most realistic, because the over- and underestimation of the object size is limited to half a pixel size in every direction.
+The best method should to use the percentage of the pixel coverage by the object as weight, which is not possible.
+As an alternative switching between setting all touched True and False may result in a better assessment of the true costs.
+When superimposing the resulting costs rasters these map will include both aspects of the correctness: showing every detail and statistically distribute better the real cost.
+Another method to achieve the same is to downsample high resolution raster.

### 2023-03-12 paper/paper-CableRoute.tex
style, grammar format, figures
moving related work into introduction
moving bits around in the chapters

@@
 	\begin{abstract}
-		The least cost paths for a power line from a power station at the cost to a power station further inland are calculated.
-		Down-sampling and clipping are investigated as techniques to speed up the calculation.
-		The path from the clipped cost raster is often the same as the original path.
+		Least Cost Paths for a power line from a power station at the cost to a power station further inland are calculated.
+		Downsampling and clipping are investigated as techniques to speed up the calculation.
+		The paths from the clipped cost raster often are the same as the original paths.
 		The speed-up of the clipping can not be known in advance.
@@
 
-	The German power system is adding a lot of renewable energy sources in the north, because wind energy reaches its highest efficiency with high wind speeds.
+	The German power system is adding renewable energy sources in the north, where wind energy plants reach their highest efficiency, due to higher wind speeds.
 	At the same time old power plants e.g.\ nuclear, hard coal and lignite are being phased out~\cite{eser-frey_analyzing_2012}.
@@
 	by the local population have to be taken in consideration.
-	Also technical aspects, as the effects on the grid stability are further points take into consideration~\cite{schafer_understanding_2022}.
+	Also technical aspects, as the effects on the grid stability are further points to take into consideration~\cite{schafer_understanding_2022}.
 
@@
 
-	Other methods such as \acrfull{MCDA} consider further aspects as stakeholder, experts, public and decision maker~\cite{bertsch_participatory_2016} (see section~\ref{sec:related-works}).
+	The different models the power systems, that are used in the expert system range from not modelling the grid at all to models that include Kirchhoffs laws~\cite{bertsch_participatory_2016}.
+	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
+	Some models not only consider the grid of one nation, but consider neighbouring states or the whole European power system~\cite{dietrich_will_2010}.	
+	Older power grid models are e.g.\ optimising the well fare, but economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an increasingly important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
 
-	The least cost path algorithm is used to plan a possible route for a power line.
-	The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
-	The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
-	This makes it possible to find routes on graphs that are not predefined, such as road networks.
-	The weights of the edges are the local costs of travel from one pixel center to the neighbouring pixel center.
-	The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate for the land usage of a given class.
-	The Least Cost Path algorithm consists of at least two successive steps.
-	1) The first step is to aggregate the costs of travelling from the starting point to a given set of end points.
-	This step generates the aggregated cost raster of travelling from the starting point to any point of the cost raster.
-	2) In the next step the backtracking, the route of the actual Least Cost Path is calculated.
-	For each end point the path via the lowest cost neighbour is taken until the starting point is reached.
+	In contrast to a GIS analysis, these other factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
+	Stakeholders as decision makers can be included and combined with an expert system~\cite{bertsch_participatory_2016}.
 
-	Some implementations switch the role of start and ending points, so that either many start points and one single end point, or one single end point and many end points can be used.
-	In some implementation there is an extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
-	path to the starting point as integer values.
+	The uncertainties associated with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are investigated by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
+	Both inter- and intra-criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.
 
@@
 	\input{discussion}
-
-	\section{Related Works}\label{sec:related-works}
-
-	The different models of the power system, that are used in the expert system range from not modelling the grid at all to models that include Kirchhoffs laws~\cite{bertsch_participatory_2016}.
-	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
-	Some models consider not only the grid of one nation, but consider neighbouring states or the whole European power system~\cite{dietrich_will_2010}.	
-	Older national power grid models are e.g.\ are optimising the well fare, but economic modelling is not sufficient for planning modern routes, as the aspect of environmental sustainability, security of supply and the public acceptance play an increasingly important role~\cite{hauff_gesellschaftliche_2011} in modern planning.
-
-	In contrast to a GIS analysis, these other factors can be studied done with a \acrfull{MCDA}~\cite{bertsch_participatory_2016}.
-	Stakeholders as decision makers can be included and combined with an expert system~\cite{bertsch_participatory_2016}.
-
-	The uncertainties associated with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are investigated by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
-	Both inter- and intra-criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.
 

### 2023-03-12 paper/results.tex
style, grammar format, figures
moving related work into introduction
moving bits around in the chapters

@@
 The Least Cost Path is estimated from this set of rasters.
-In the last step the Least Cost Path is computed from the lower resolution rasters and compared with the Least Cost Paths computed from a high resolution raster.
+In the last step the Least Cost Path is computed from the medium resolution rasters and compared with the Least Cost Paths computed from a high resolution raster.
 
@@
 %%The surrounding of the study area, uses a no-data value and will not be use for the calculation of the Least Cost Path.
-The different cost layers in the study area are aggregated by the maximum function.
-If the resolution is higher than the object size, then the effect of the rasterisation with all touched True or False is limited.
-For all touched True any part of the pixel that is covered by the object, will attribute that whole pixel to the object.
+The different intermediate cost rasters in the study area are aggregated by the maximum function.
+If the resolution is higher than the object size, then the effect of setting all touched to True or False is limited.
+If all touched is set to True and any part of the pixel that is covered by the object, the whole pixel is attributed to the object.
 This makes the object appear a halve a pixel size larger in all directs.
@@
 	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alF_v2_small.png} }}%
-	\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 5~m.}
+	\caption{Part of the cost raster. Contrasting the for different settings of all touched at a resolution of 5~m.}
 	\label{fig:costs_5m}
@@
 	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alF_v2_small.png} }}%
-	\caption{Figures of the cost raster. Contrasting the for different values in all touched at a resolution of 100~m.}
+	\caption{Part of the cost raster. Contrasting the different settings for all touched at a resolution of 100~m.}
 	\label{fig:costs_100m}
@@
 	\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m_v2_small.png} }}%
-	\caption{Figures of the least cost paths.
+	\caption{Figures of the Least Cost Paths.
 Contrasting the paths for different resolutions.
-Paths computed from all touched false raster are indicated by dashed lines.
-Results from all touched True are indicuated by continuous lines.
+Paths with all touched set to False are indicated by dashed lines and True are indicated by continuous lines.
 Higher resolutions are indicated by the color green, lower resolutions by the color red.
@@
 
-	\caption{Figures of the least cost paths.
-Contrasting the changes of the least cost paths for the different results, depending on the parameter all\_touched.
-Paths computed from all touched false raster are indicated by dashed lines.
-Results from all touched True are indicuated by continuous lines.
+	\caption{Figures of the Least Cost Paths.
+Contrasting the changes of the least cost paths for the different results, depending on the parameter all touched.
+Paths setting all touched to False are indicated by dashed lines and True are indicated by continuous lines.
 Higher resolutions are indicated by the color green, lower resolutions by the color red.

### 2023-03-12 paper/discussion.tex
grammar, style, typos
understanding

@@
 Similarly, the use of main memory increases. 
-This again limits the number data points, that can be processed and resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls have been used for the aggregation of higher resolution.
+This again limits the number data points, that can be processed and resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls has been used for the aggregation of higher resolution.
 
@@
 These methods are used as surrogates for the more complex calculation of the Least Cost Path with the higher resolution.
-In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all touched false and true raster where averaged in different ratios, to compute the optimum weighted cost raster.
+In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all touched set to False and True raster were averaged in different ratios, to compute the optimum weighted cost raster.
 While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
@@
 
-This early stopping may result in suboptimal paths around the end points for some edge cases, where the connection via another neighbour might be more optimal.
+This shows, that the way the cost raster is created in the first place can play a crucial role, in the end result.
+So that a nuance, can cause a detour.
+When this behaviour occurs, the polygon may not include the Least Cost Path.
+This polygon should therefore be overlapped with a polygon around the shortest path.
+For the set of control paths downsampling could not outperform the original medium resolution rasters.
 
-The set of rules that are used create the cost raster, includes a rule to create buffer around buildings which is set to the level \textit{prohibited} areas.
-In all touched false rasters.
-The resolution of the medium level raster needs to be high enough to show every detail.
+Early stopping may result in suboptimal paths around the end points for some edge cases, where the connection via another neighbour might be more optimal.
+
+The set of rules, that are used to create the intermediate cost raster, includes a rule to create buffer around buildings which is set to the level \textit{prohibited} areas.
+In all touched set to False rasters, the resolution of the medium level raster needs to be high enough to show every detail.
 At least in the magnitude of the minimum object size plus twice its buffer.
-This is true for the 10~m resolution raster and less true for the 25~m resolution raster that misses some details for roads for all touched false raster.
+This is true for the 10~m resolution raster and less true for the 25~m resolution raster, that misses some details for roads for all touched false raster.
 Other details such as rivers and houses, are already included in the lower resolution raster, due to larger buffers.
+The raster with all~touched set to False might miss some objects, but the missing chance is propotional to the object size and the extent of overestimation is reduced.
 The Least Cost Path algorithm searches for an optimal path as a line.
@@
 Therefore, the used resolution should not be smaller than, the width of the object that should be placed.
-This can be avoided by downsampling, but by weighting the medium resolution all touched true or false rasters.
+This can not be avoided by downsampling, but by weighting the medium resolution all~touched True and False rasters.
 
@@
 If multiple paths are calculated from a single raster, the speed-up benefit is reduced.
-Especially, pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective.
-As the number of paths increases, fewer pixels are clipped.
+Especially, pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective, because as the number of paths increases, fewer pixels are clipped.
 The Least Cost Path algorithm does only select the single most cost-effective path.
@@
 
-In this work, the intermediate cost raster layers are aggregated using the maximum function. 
-Another possible aggregation function is the sum or average.
+In this work, the intermediate cost raster layer are aggregated using the maximum function. 
+Other possible aggregation functions are the sum and average functions.
 Each aggregation function can be justified, by a different interpretation of the cost and its scale.
 
-When the \textit{prohibited} level is used as the highest level, then summing the two highest levels would result in a new highest level. 
+When the \textit{Prohibited} level is used as the highest level, then summing the two highest levels would result in a new highest level. 
 Also, the maximum function does not interfere with the nodata value. 
-Although this can be done by a nansum- / nanmean-function, if the nodata value is set to \textit{not a number} during the aggregation.
+This can be done by a nansum- / nanmean-function, if the nodata value is set to \textit{not a number} during the aggregation.
 The disadvantage of aggregation with the maximum function is, that this aggregation is unable to distinguish between nuances of different overlapping intermediate costs.
@@
 As the default level increases, the effect would probably be reduced for low resolution rasters.
-For high resolution raster, the effect would still be present, because the fact, that the raster pixel centre is used for sampling, reflects the original geometry better.
+For low resolution raster, the effect would still be present, because the fact, that the raster pixel centre is used for sampling, reflects the original geometry better.
 
-This effect of the similar aggregated costs per resolution could also be seen in the test paths, even when the paths varied greatly with the change in resolution or algorithm.
+This effect of the similar aggregated costs per resolution could also be seen in the test paths, even when the paths varied greatly.
 This could be an indication of an even spatial distribution of the costs.
 
-The all touched true cost raster shows every detail, but the sampling with all touched true increases the size of the features.
-The fact that the aggregated costs per resolution for all touched true rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the values of the costs are not uniformly distributed, but that the high costs are much more frequent, because the costs are exponential scaled.
+The all~touched set to True cost raster shows every detail, but the sampling with all~touched set to True increases the size of the features.
+The fact that the aggregated costs per resolution for all touched true rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, because the costs are exponential scaled.

### 2023-03-12 paper/methods.tex
grammar, style, typos
understanding

@@
 
-The least cost path algorithm is used to plan a possible route for a power line.
-The least cost path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
+The Least Cost Path algorithm is used to plan a possible route for a power line.
+The Least Cost Path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
 The vertices of the graph are the pixel centres, that are connected to the eight neighbouring pixel centres via the edges.
 This makes it possible to find routes on graphs that are not predefined, such as road networks.
-The weights of the edges are the local costs of travel from one pixel center to the neighbouring pixel center.
+The weights of the edges are the local costs of transit from one pixel center to the neighbouring pixel center.
 The costs can be physical costs, such as the local slope, but also can be composed of other factors as the acceptance rate to transverse a given land usage.
@@
 
-Some implementations switch the role of start and ending points, so that either many start points and one single end point, or one single end point and many end points can be used.
-In some implementation there is an extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
+Some implementations switch the roles of start and ending points, so that either many start points and one single end point, or one single end point and many end points can be used.
+In some implementations there is an extra step between 1) and 2) that calculates cost-weight direction raster, that encodes the direction of the shortest
 path to the starting point as integer values.
@@
 
-PyWPS~\cite{noauthor_welcome_2016} is used to provide the least cost path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
-As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster and receives the resulting least cost path.
+PyWPS~\cite{noauthor_welcome_2016} is used to provide the Least Cost Path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
+As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster and receives the resulting Least Cost Path.
 The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
@@
 
-To computed intermediate cost cost raster the different layers of the different entities are optionally filtered, buffered and then rasterised.
+In order to compute intermediate cost raster the different vector layers of the different entities are optionally filtered, buffered and then rasterised.
 Filtering the layers of the vector files for special attributes enables further differentiation.
-For examples, it is possible to differentiate between heath and uncultivated land in the land cover.
-Adding a buffer can be used either to convert a line object such as power lines and roads into a polygon (with the
+For example, it is possible to differentiate between heath and uncultivated land in the land cover.
+Adding a buffer can be used either to convert a line object such as a power line and a road into a polygon (with the
 correct physical width), or to add minimum distance from an existing of planed area to the new power line.
@@
 to \textit{Restricted}, \textit{Strongly Restricted} and \textit{Prohibited} areas with high costs.
-These higher costs resemble the degree to which a place with this cost should be avoided, while routing the path.
-The ratio of the higher costs to the lower costs directly translates into the additional detour in pixels
+These higher costs represent the degree to which a place with this cost should be avoided, while routing the path.
+The ratio of the higher costs to the lower costs equals the detour
 the algorithm is willing to go.
 Thus, as \textit{prohibited} areas describe a legal obligation, not to use these areas or only to the utmost minimum,
-the weight that resembles the costs for these types of areas, is set especially high.\\
+the weight that resembles the costs for these types of areas is set especially high.\\
 
@@
 Both rasterisation techniques differ in the selection of the pixel, that describe the original polygon.
-The pixels can be selected, if either the centre of that pixel is overlapped
+A pixels will be selected, if either the centre of that pixel is overlapped
 by the geo-object, or any part of the pixel is overlaid.
-Setting all touched to True, implies the version with any part of the pixel selected.
-The version where an overlapped pixel centre is require is setting the parameter all touched to False.
-All touched set to False is considered the default (see figure~\ref{fig:alltouched}).
+Setting all~touched to True implies the version with any part of the pixel selected.
+The version, where an overlapped pixel centre is required, is setting the parameter all~touched to False.
+All~touched set to False is considered the default (see figure~\ref{fig:alltouched}).
 \begin{figure}[!ht]
@@
 	}%
-	\caption{Graphical example for rasterise a vector (left blue), to a raster (right blue) with either all touched set to False (above), or True (below).}
+	\caption{Graphical example the rasterisation of a vector (left blue), to a raster (right blue) with either all~touched set to False (above), or True (below).}
 	\label{fig:alltouched}
@@
 
-The complete list of layers and the applied processing steps, can be found in \textbf{Supplement S1}.
+The complete list of layers and the applied processing steps can be found in \textbf{Supplement S1}.
 
-All three steps of the generation of the least cost path: generation of the cost raster, aggregation and backtracking is shown with an example for a cost raster of 50~m resolution and all touched set to False (see \ref{fig:costs2path}).
+All three steps of the generation of the Least Cost Path: generation of the cost raster, aggregation and backtracking is shown with an example for a cost raster of 50~m resolution and all~touched set to False (see \ref{fig:costs2path}).
 
@@
 After finding an aggregated cost for every end point, the aggregation stops and the backtracking starts.
-Because the path ends at a power transformer, which is a building type, the paths ends at in a \textit{Prohibited} area.
+Because the path ends at a power transformer, which is a building type, the paths end at in a \textit{Prohibited} area.
 Therefore, areas even further away from the starting point have been explored first.
@@
 
-	\caption{Figures of the cost raster and the resulting aggregated costs and the Least Cost Path for a resolution of 50~m, all touched set to False.}
+	\caption{Figures of the cost raster and the resulting aggregated costs and the Least Cost Path for a resolution of 50~m, all~touched set to False.}
 	\label{fig:costs2path}
@@
 
-For low resolution rasterisation with all touched set to True, will show every detail, but the objects are enlarged.
-When all touched is to False the object only appears, when the is situated at the pixel centre.
+For low resolution rasterisation, with all~touched set to True will show every detail, but the objects are enlarged.
+When all~touched is to False the object only appears, when the is situated at the pixel centre.
 Thus, this might be used as surrogate, that expresses the likelihood of the object to be sampled and correlates with the object size  compared to the pixel size.
-At high resolution the set all touched to True still overestimates the object size, but the extent is limited.
-Setting all touched to False will include all objects for high resolution.
+At high resolution the set all~touched to True still overestimates the object size, but the extent is limited.
+Setting all~touched to False will include all objects for high resolution.
 This setting is most realistic, because the over- and underestimation of the object size is limited to half a pixel size in every direction.
-The best method should to use the percentage of the pixel coverage by the object as weight, which is not possible.
-As an alternative switching between setting all touched True and False may result in a better assessment of the true costs.
-When superimposing the resulting costs rasters these map will include both aspects of the correctness: showing every detail and statistically distribute better the real cost.
+The best method should be to use the percentage of the pixel coverage by the object as weight, which is not possible.
+As an alternative, switching between setting all~touched True and False may result in a better assessment of the true costs.
+When superimposing the resulting cost raster, these map will include both aspects of the correctness: showing every detail and statistically distribute better the real cost.
 Another method to achieve the same is to downsample high resolution raster.

### 2023-03-12 paper/paper-CableRoute.tex
grammar, style, typos
understanding

@@
 	\begin{abstract}
-		Least Cost Paths for a power line from a power station at the cost to a power station further inland are calculated.
+		Least Cost Paths for a power line from a power station at the shore line to a power station further inland are calculated.
 		Downsampling and clipping are investigated as techniques to speed up the calculation.
@@
 	%% the work being presented. Separate the keywords with commas.
-	\keywords{GIS, least cost path}
+	\keywords{GIS, Least Cost Path}
 
@@
 	However, it is not enough to find the shortest route when building new power lines.
-	Other parameters have to be taken into consideration.
-	As the steepness of a road or the soil, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
+	Other parameters have to be taken into consideration, as the steepness of a road or the soil for example, play an important role for the building cost of a road or pipeline~\cite{suleiman_optimal_2015}.
 	When planing the additional routes for a power grid, further aspects such as legal regulations and acceptance
-	by the local population have to be taken in consideration.
+	by the local population have to kept in mind.
 	Also technical aspects, as the effects on the grid stability are further points to take into consideration~\cite{schafer_understanding_2022}.
 
-	Due to the increasing demand for renewable energy wind offshore wind turbines supply 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.
+	Due to the increasing demand for renewable energy wind offshore wind turbines supplied 5.5~\% annual percentage in 2020 of the German energy mix~\cite{noauthor_nettostromerzeugung_2021}.
 
-	The different models the power systems, that are used in the expert system range from not modelling the grid at all to models that include Kirchhoffs laws~\cite{bertsch_participatory_2016}.
-	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid and the actual grid data are confidential~\cite{bertsch_participatory_2016}.
+	 The physical modulation in the power systems models, range from not modelling the grid at all to models  The use of Kirchhoffs laws~\cite{bertsch_participatory_2016}.
+	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid.
+	Besides, the actual grid data are confidential~\cite{bertsch_participatory_2016}.
 	Some models not only consider the grid of one nation, but consider neighbouring states or the whole European power system~\cite{dietrich_will_2010}.	
@@
 
-	The uncertainties associated with \acrshort{MCDA} data uncertainties, preferential uncertainties and model uncertainties are investigated by a with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
+	The uncertainties associated with \acrshort{MCDA} are data uncertainties, preferential uncertainties and model uncertainties are investigated with a sensitivity analysis~\cite{bertsch_participatory_2016} or simulation~\cite{butler_simulation_1997}.
 	Both inter- and intra-criteria preferential uncertainties~\cite{bertsch_sensitivity_2007} can be considered.
@@
 	The cost of computing of the Least Cost Path scales with the square of the resolution.
-	The difference between the aggregated costs of the paths per resolution computed from the all touched false raster and the all touched true raster only shows a linear decrease.
+	The difference between the aggregated costs of the paths per resolution computed from the all~touched set to False raster True raster only shows a linear decrease.
 	Therefore, the gain in accuracy per compute time decreases.
-	The presented methods have attempted to circumvent this bottleneck.
-	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the least cost path from the high resolution raster.
+	The presented methods attempt to circumvent this bottleneck.
+	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the Least Cost Path from the high resolution raster.
 	The paths can change significantly even for a small change in total cost.
 	Therefore, to develop an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy, to offer alternative paths and at the same time show variations between these paths.
-	An alternative for future methods, would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
+	An alternative would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
 	This again increases the computation time.

### 2023-03-12 paper/results.tex
grammar, style, typos
understanding

@@
 The different intermediate cost rasters in the study area are aggregated by the maximum function.
-If the resolution is higher than the object size, then the effect of setting all touched to True or False is limited.
-If all touched is set to True and any part of the pixel that is covered by the object, the whole pixel is attributed to the object.
-This makes the object appear a halve a pixel size larger in all directs.
+If the resolution is higher than the object size, then the effect of setting all~touched to True or False is limited.
+If all~touched is set to True and any part of the pixel that is covered by the object, the whole pixel is attributed to the object.
+This makes the object appear larger.
 As can be seen in figure~\ref{fig:costs_5m}, which shows a detailed view of the costs for the village of Beverstedt.
- %% Due to the maximum aggregation of the costs, the average cost of the raster of all touched true will be over estimated.
-The rasterisation with all touched false will be a better description of the real size of the object, for high resolution.
+ %% Due to the maximum aggregation of the costs, the average cost of the raster of all~touched true will be over estimated.
+To set all~touched to False is a better description of the real size of the object, for high resolution.
+
 \begin{figure}
@@
 	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_5m_alF_v2_small.png} }}%
-	\caption{Part of the cost raster. Contrasting the for different settings of all touched at a resolution of 5~m.}
+	\caption{Part of the cost raster. Contrasting the for different settings of all~touched at a resolution of 5~m.}
 	\label{fig:costs_5m}
@@
 
-In contrast, if the resolution is smaller, than the object-size this behaviour changes.
-In general, while the area, of the pixel increases for all touched True, the area for which the cost is overestimated also increases.
-As the pixel size for all touched  False increases, not only does the overestimated area increases, but in addition the object is less probable situated in the centre of the pixel. 
-Therefore, object is sampled at random.
-Hence, all touched false leads to a loss of information for smaller objects.
-\todo{Put this part into Discussion?} Since the default cost is much smaller than the average cost, this method underestimates the cost.
-The figure~\ref{fig:costs_100m} shows, that for the resolution of 100 m, that larger objects are still included in the map, but appear to be larger.
-But smaller objects, such as roads, are not fully or randomly included.
+In contrast, if the resolution is smaller all~touched set to False leads to a loss of information for smaller objects.
+Since the default cost is much smaller than the average cost, this method underestimates the cost.
+The figure~\ref{fig:costs_100m} shows, that for the resolution of 100 m, that larger objects are still included in the map, but smaller objects, such as roads, are partially included.
 \begin{figure}
@@
 	\subfloat[\centering All touched: False.]{{\includegraphics[width=.45\linewidth]{./images/CostRaster_100m_alF_v2_small.png} }}%
-	\caption{Part of the cost raster. Contrasting the different settings for all touched at a resolution of 100~m.}
+	\caption{Part of the cost raster. Contrasting the different settings for all~touched at a resolution of 100~m.}
 	\label{fig:costs_100m}
 \end{figure}
-\todo{Put this part into Discussion?}Thus, with having larger areas and more objects, with higher costs, all touched True rasterisation might more likely lead to longer roots and more likely to block the  direct spatial path.
 
 \subsection{Least Cost Paths}\label{subsec:least-cost-paths}
-For each resolution the Least Cost Paths were estimated from the all\_touched False and all\_touched True rasters.
+For each resolution the Least Cost Paths were estimated with the all~touched set to False and True.
 
-For the study, a start point were chosen at a transformer about 6~km north of the container terminal Bremerhaven and an end point at a transformer in the southeast of the Osterholz county.
+For the study, a start point was chosen at a transformer about 6~km north of the container terminal Bremerhaven and an end point at a transformer in the southeast of the Osterholz county.
 
-The distance of Least Cost Paths of the all touched false to the paths of the all touched true raster is calculated by the mean minimum distance.
-For each vertex $P_i$ in the path $L_1$ the minimum distance between the vertex and the path $L_2$
+The distance between the paths is calculated by the mean minimum distance.
+For each vertex $P_i$ in the path $L_1$ the minimum distance between the vertex $P_i$ and the path $L_2$
 is calculated and then the minimum distances are averaged (see equation~\ref{eq:1}).
@@
 	\label{eq:1}
-	d_{mean} = \frac{1}{|L_1|} \sum_{i=1}^{n} d_{min}(p_i, L_2) \Bigr\vert p_i \in L_1
+	d_{mean} = \frac{1}{|L_1|} \sum_{i=1}^{n} d_{min}(P_i, L_2) \Bigr\vert P_i \in L_1
 \end{equation}
 This equation is used, to measure the degree of similarity between the paths.
+The distance are measured at the same resolution (different setting for all~touched), and same setting of all~touched (different resolution).
 
-Table~\ref{tab:2} shows the distance between the two paths decreases
+Table~\ref{tab:2} shows, that the distance between two paths decreases
 with increasing resolution.
@@
 
-At the same time, the differences in the aggregated costs per resolution remain almost constant.
-On one hand it can be seen, that the all\_touched False underestimates the costs and that this tendency scales
-linearly with the resolution.
-On the other hand, the all\_touched True least cost overestimates the aggregated costs on a linear scale of
-the resolution.
+At the same time, the differences in the aggregated costs remain almost constant.
+Thus the difference between the aggregated costs per resolution decreases.
+Setting all~touched to False underestimates and to True overestimates the costs.
 
@@
 Contrasting the paths for different resolutions.
-Paths with all touched set to False are indicated by dashed lines and True are indicated by continuous lines.
+Paths with all~touched set to False are indicated by dashed lines and True are indicated by continuous lines.
 Higher resolutions are indicated by the color green, lower resolutions by the color red.
@@
 
-	\subfloat[\centering all touched false.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F_v2_small.png} }}%
+	\subfloat[\centering all touched False.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_F_v2_small.png} }}%
 	\qquad
-	\subfloat[\centering all touched true.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T_v2_small.png} }}%
+	\subfloat[\centering all touched True.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_al_T_v2_small.png} }}%
 
 	\caption{Figures of the Least Cost Paths.
-Contrasting the changes of the least cost paths for the different results, depending on the parameter all touched.
-Paths setting all touched to False are indicated by dashed lines and True are indicated by continuous lines.
+Contrasting the changes of the least cost paths for the different results, depending on the parameter all~touched.
+Paths setting all~touched to False are indicated by dashed lines and True are indicated by continuous lines.
 Higher resolutions are indicated by the color green, lower resolutions by the color red.
@@
 \begin{table*}[t]
-	\caption{Least cost paths as length for the different resolution of the raster, including the mean minimum distance and the maximum minimum distance and the agg.
+	\caption{Least Cost Paths as length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$) and the maximum minimum distance~($d_{max}$) and the agg.
 costs.
@@
 costs and the agg.
-costs per resolution are given.}
+costs per resolution are computed.}
 	\label{tab:2}
@@
 
-When estimating distance between the Least Costs Paths from all\_touched True rasterisation and all\_touched False at the same resolution 
+When estimating the distance between the Least Costs Paths from all~touched set to True and all~touched set to False at the same resolution 
 the mean minimum distance between the 100~m resolution paths is 1946.41~m and between the 5~m resolution paths is 126.04~m.
-The similarity for the all touched False paths is higher, than for the all touched true raster.
-The distance for the 100~m path to the 5~m resolution path is 243.42~m for all touched false and 2109.44~m for all touched true.
+The similarity between the all~touched set to False paths is higher, than for setting to True.
+The distance for the 100~m path to the 5~m resolution path is 243.42~m for all~touched false and 2109.44~m for all~touched true.
 
-When comparing the similarity between the all touched false paths themselves and the similarity between the all touched false rasters to the all touched true rasters of the same resolution: The similarity in between the all all\_touched False paths is higher than, than for most paths of same resolution.
+When cross comparing the similarity between all paths set with all~touched set to False and the similarity in between the same resolution, the similarity in between the all~touched set to False paths is higher than, than for most paths of same resolution.
 Namely except for the highest resolution.
@@
 This behaviour is shown in figure~\ref{fig:paths_alltouched}.
-On a more detailed level, it can be seen, that the paths of all\_touched False also converge directly to the all touched True paths, but the extent is smaller.
-The length of the paths only differs by a maximum of about 10 \%.
-On the other hand, side the length of the paths can increase, because with more vertices are used with higher resolution.
+On a more detailed level, it can be seen, that the paths of all~touched False also converge directly to the all~touched set to True paths, but the extent is smaller.
 
-The zonal stat (see table~\ref{tab:3}) for a buffer of 100~m (5~m) around the path has been used, to estimate the
-percentage of each costs levels around the path.
-When using all\_touched True rasterisation at higher resolution, the tendency is to use a higher percentage of the
-\textit{Preferential Level} and less of the\textit{NoRestriction} Level.
-The ratio of the 100~m buffered Least Cost Path, strongly shifted  to Levels lower costs.
+The zonal stat (see table~\ref{tab:3}) for a buffer of 100~m (5~m) around the paths has been used, to estimate the
+percentage of each costs levels around the paths.
+When using all~touched True at higher resolution, the tendency is to use a higher percentage of the
+\textit{Preferential} Level and less of the \textit{NoRestriction} Level.
+The ratio of the 100~m buffered Least Cost Path, shifted  to Levels lower costs.
 
-There is no strong tendency for the all\_touched False least cost paths.
-
-\subsection{Execution time}\label{subsec:execution-time}
-
-In theory, the execution time increases with the square of the resolution, because higher resolutions result in a higher number of pixels and thus data points the aggregated costs needs to be calculated for.
-A full logarithmic fit for several repetitions of the execution shows, that the execution time scales with power of $2.1997  \pm 0.007$ of the inverse resolution.
-
-The total execution time consists of two parts. 
-The aggregation of the costs and the back tracking of the least cost to find the path.
+There is no strong tendency for the all~touched set to False Least Cost Paths.
 
@@
 \begin{table*}[t]
-	\caption{resolution (r) of Category percentages of each least cost path for a buff of 100 m (5 m) around the least cost path.}
+	\caption{Resolution (res) of category percentages of each Least Cost Path for a buffer of 100 m (5 m) around the Least Cost Path for each raster reprojected to 5~m resolution.}
 	\label{tab:3}
@@
 
+\subsection{Execution time}\label{subsec:execution-time}
+
+In theory, the execution time increases by power of two with the resolution, because higher resolutions result in a higher number of pixels. 
+A double logarithmic fit shows, that the execution time scales with power of $2.1997  \pm 0.007$ of the inverse resolution.
+
+The total execution time consists of two parts. 
+The aggregation of the costs and the back tracking of the least cost to find the path.
+
 \subsection{Faster Processing of the Cost Path Algorithm}\label{subsec:faster-processing-of-the-cost-path-algorithm}
@@
 
-\subsubsection{Compare least cost paths, for overlay of both rasterisations}
+\subsubsection{Compare Least Cost Paths, from rasters of both all~touched settings}
 
-For the example paths shown, all touched true rasterisation overestimates the true costs and all touched true underestimates them.
+For the example paths shown, all~touched set to True overestimates ans set to False underestimates the true costs.
 
-A weighted average of the costs could therefore be a more accurate measure and make estimated medium resolution Least Cost Path more similar, to high resolution paths.
-As all example show, the weighting should favour the all touched false raster.
-\todo{move to discussion/conclussion:}The best weight should be the percentage of the pixels, which is covered by the object, but this can not be calculated in this work.
-An alternative would be to compute the cost raster at a high resolution and reproject them to a medium resolution by a (linear) interpolation of the weights.
+A weighted average of the costs could therefore may be a more accurate measure and make estimated medium resolution Least Cost Path more similar, to high resolution paths.
+As above example show, the weighting should favour all~touched to False.
 
-\todo{move to planning:}This will speed-up the aggregation.
+This will speed-up the aggregation.
 The time needed for the back tracking stays unchanged.
 
-The optimal ratio of overlaying all touched false and all touched true cost raster for 10~m resolution is estimated via similarity of the resulting Least Cost Path to the path of the original high Resolution raster.
-The mean distance of Least Cost Paths for different ratio is estimated to the path from the all touched false raster of the higher (5~m) resolution. 
-Table~\ref{tab:4} shows, that the distance decreases, with increasing ratio (1:1, 2:1, 4:1) and after this optimum is reached, increases with increasing ratio (8:1, 16:1 and so on).
-Comparing  the similarity of the paths from the rasters of the different ratios to normal paths with 10~m resolution, paths with a higher ratio of all touched true is nearer to the all touched true paths.
-Paths with a ratio in favour of all touched false are much closer to the all touched false paths.
+The optimal ratio of overlaying both all~touched for cost raster with 10~m resolution is estimated via similarity of the resulting Least Cost Path to the path of the original high resolution raster. 
+The mean distance of Least Cost Paths from rasters with different ratio is estimated to the path from the all~touched set to False raster of the higher (5~m) resolution. 
+Table~\ref{tab:4} shows, that the mean minimum distance decreases, with increasing ratio (1:1, 2:1, 4:1) and after this optimum is reached, increases with increasing ratio (8:1, 16:1 and so on).
+Comparing  the similarity of the paths of the different ratios to normal paths with 10~m resolution shows, paths with a higher ratio of all~touched set to True is nearer to the all~touched set to True paths.
+Paths with a ratio in favour of all~touched false are much closer to the all~touched set to False paths.
 
 \begin{table}[h!]
-	\caption{Length of the path computed from the overlaying of all touched false and true raster and the mean distance of the paths to the paths calculated from the all touched false 5~m resolution  and all touched false and true raster of 10~m resolution.}
+	\caption{Paths computed from the overlaying of all~touched set to False and True raster with the mean minium distance (d) of the paths to the paths calculated from the all~touched set to False 5~m resolution and all~touched set to False and True raster at 10~m resolution with the ratio (r).}
 	\label{tab:4}
@@
 
-\subsubsection{Compare least cost paths, for down sampled cost paths}
+\subsubsection{Compare Least Cost Paths, from downsampled cost raster}
 
-As an alternative to the superposition of the all touched true and false rasters for the same resolution, the all touched false raster is down-sampling to 10~m, 25~m, 50~m and 100~m (with bi-linear) interpolation.
-With this method, smaller structures can still be fully seen in the cost raster, although the resolution is reduced.
-The distances of the paths that are computed from the bi-linear down sampled raster to the path of the original 5~m resolution (all touched false) shows (see table~\ref{tab:5}), that only down-sampling to a resolution of 10~m, produces a path that is relatively close the high resolution path.
+As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all~touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (with bi-linear) interpolation.
 
-The opposite is true for the lower resolution raster which is more similar to paths computed from the all touched true cost raster.
-Every path from a down sampled raster is more similar to a path computed from an all touched true raster, than an all touched false raster, although the all touched false raster of the 5 m resolution was used for down sampling.\todo{interpretation all touched true and downsampling, both show all details, but downsampling, more similar to original geometry}
+The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all~touched False) shows, that only downsampling to a resolution of 10~m, produces a path that is relatively close the high resolution paths(see table~\ref{tab:5}).
+
+The opposite is true for the lower resolution raster which is more similar to paths computed from the all~touched true cost raster.
+Every path from a downsampled raster is more similar to a path computed from an all~touched set to True raster, than an set to False raster, although the all~touched set to False raster of the 5 m resolution was used for downsampling.\todo{interpretation all~touched true and downsampling, both show all details, but downsampling, more similar to original geometry}
 
 \begin{table}[h!]
-	\caption{Length of the path computed from the bi-linear down sampled raster and the mean distance of the paths from the down sampled raster to the paths calculated from the all touched true and all touched false raster of the same resolution as the down sampled raster.}
+	\caption{Length (l) of the path computed from the bi-linear downsampled raster and the mean minimum distance (d) of the paths from the downsampled raster to the paths calculated from the all~touched set to True and False raster of the same resolution as the downsampled raster with resolution (res).}
 	\label{tab:5}
@@
 % \subsubsection{Restrict search to a minimum sized bounding box}
-\subsubsection{Restrict search to a buffered around the least cost paths}
+\subsubsection{Restrict search to a buffer around the Least Cost Paths}
 
-Construct a polygon from the two Least Cost Paths (all touched true and all touched false of the same resolution).
+Construct a polygon from the two Least Cost Paths (all~touched true and all~touched false of the same resolution).
 Buffer the polygon with twice the maximum  minimum path distance  (see equation~\ref{eq:2}).
@@
 	\label{eq:2}
-	d_{max} = max(\sum_{i=1}^{n} d_{min}(p_i, L_2)) \Bigr\vert p_i \in L_1
+	d_{max} = max(\sum_{i=1}^{n} d_{min}(P_i, L_2)) \Bigr\vert P_i \in L_1
 \end{equation}
 
-This provided the possibility  to run a 2.5~m resolution cost raster and clip it to the extent of the polygon. 
-This clipped 2.5~m raster for all touched true changed the path only slightly change.
+This strategy resulted in the same Least Cost path as with the original high resolution raster.
 
-The all touched False raster, on the other hand, leads to a completely new previously unused subroute at the end of the path.
+This reduction of compute power provided the possibility  to run a 2.5~m.
+This clipped 2.5~m raster for all~touched set to True changed the path only slightly.
+
+The all~touched False raster, on the other hand, leads to a completely new previously unused subroute at the end of the path.
 Due to the low resolution a small path became passable.
@@
 The road and the protected landscape area are both \textit{restricted} areas, while the power line is \textit{preferred}.
-This shows, that the way the cost raster is created in the first place can play a crucial role, in the end result.
-So that a nuance, can cause a detour.
-When this behaviour occurs, the polygon may not include the Least Cost Path.
-\todo{move: discussion}This polygon should therefore be overlapped with a polygon around the shortest path.
 
@@
 
-\subsubsection{Examine the proposed solutions}
-To broaden the view and verify the result, e.g.\ start and end points, four different routes should be found using the above strategies.
-Two routes should be found form the start point to two new points in the south east of the investigated area and two routes should be found from the north and north east of the study area to the end point.
+\subsubsection{Apply the proposed solutions on other paths}
+To broaden the view and verify the result, four different routes should be found using the above strategies.
+Two routes should be found from the start point to two new points in the south east of the investigated area and two routes should be found from the north and north east of the study area to the end point.
 
 For three of the four routes, the Least Cost Path estimation from the clipped raster, was able to calculated exactly the same result.
-For the fourth path, the Least Cost Path from the 5~m resolution raster was clipped out by the buffer around the 50~m resolution paths.
-The speed up from the clipping of the higher resolution raster depends on the number of pixels that, have been clipped.
+For the fourth path, the Least Cost Path from the 5~m resolution raster was clipped by the buffer around the 50~m resolution paths.
+The speed up from the clipping of the higher resolution raster depends on the number of pixels, that have been clipped.
 
-Bi-linear down sampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
-The aggregated cost per resolution of the Least Cost Path from the down-sampled raster is higher, than that from the higher resolution 5~m raster and the normal medium resolution 10~m raster.
+Bi-linear downsampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
+The aggregated cost per resolution of the Least Cost Path from the downsampled raster is higher, than the normal medium resolution 10~m raster.
 In addition, the distance from these paths to the high resolution path  is greater, than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
-\todo{interpretation: Why is the down-sampling result so different? Don't know.}
+\todo{interpretation: Why is the downsampling result so different? Don't know.}
 
 \begin{table}[h!]
-	\caption{Length of the path, the resolution corrected costs and the mean distance to the path created from the 5~m resolution all touched false raster for the four control routes, for the reference path of constructed from the 5~m  and 10~m raster and the down 5~m to 10~m down sampled raster and the 5~m clipped raster.}
+	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched false raster for the four control routes, for the reference path constructed from the 5~m  and 10~m raster and the from 5~m to 10~m downsampled raster and the 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), start point to point3 (S-P3) and start point to point4 (S-P4).}
 	\label{tab:6}

### 2023-03-14 paper/discussion.tex
grammar, typos

@@
 At least in the magnitude of the minimum object size plus twice its buffer.
-This is true for the 10~m resolution raster and less true for the 25~m resolution raster, that misses some details for roads for all touched false raster.
+This is true for the 10~m resolution raster and less true for the 25~m resolution raster, that misses some details for roads for all touched False raster.
 Other details such as rivers and houses, are already included in the lower resolution raster, due to larger buffers.
@@
 
-The fact that all touched false rasters produce more similar results than high resolution raster, is probably due to the fact, that the default level is relatively low.
+The fact that all touched False rasters produce more similar results than high resolution raster, is probably due to the fact, that the default level is relatively low.
 As the default level increases, the effect would probably be reduced for low resolution rasters.
@@
 The all~touched set to True cost raster shows every detail, but the sampling with all~touched set to True increases the size of the features.
-The fact that the aggregated costs per resolution for all touched true rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, because the costs are exponential scaled.
+The fact that the aggregated costs per resolution for all touched True rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, because the costs are exponential scaled.

### 2023-03-14 paper/paper-CableRoute.tex
grammar, typos

@@
 
-	 The physical modulation in the power systems models, range from not modelling the grid at all to models  The use of Kirchhoffs laws~\cite{bertsch_participatory_2016}.
+	 The physical modelling in the power systems models ranges from not modelling the grid at all to models that use of Kirchhoffs laws~\cite{bertsch_participatory_2016}.
 	The calculation of a power system is complex, because changing one edge changes the flow in all parts of grid.

### 2023-03-14 paper/results.tex
grammar, typos

@@
 This makes the object appear larger.
-As can be seen in figure~\ref{fig:costs_5m}, which shows a detailed view of the costs for the village of Beverstedt.
+This can be seen in figure~\ref{fig:costs_5m}, which shows a detailed view of the costs for the village of Beverstedt.
  %% Due to the maximum aggregation of the costs, the average cost of the raster of all~touched true will be over estimated.
-To set all~touched to False is a better description of the real size of the object, for high resolution.
+To set all~touched to False is a better description of the real size of the object for high resolution.
 
@@
 
-In contrast, if the resolution is smaller all~touched set to False leads to a loss of information for smaller objects.
+In contrast, if the resolution is smaller, all~touched set to False leads to a loss of information for smaller objects.
 Since the default cost is much smaller than the average cost, this method underestimates the cost.
-The figure~\ref{fig:costs_100m} shows, that for the resolution of 100 m, that larger objects are still included in the map, but smaller objects, such as roads, are partially included.
+The figure~\ref{fig:costs_100m} shows, that for the resolution of 100 m, larger objects are still included in the map, but smaller objects, such as roads, are only partially included.
 \begin{figure}
@@
 
-For the study, a start point was chosen at a transformer about 6~km north of the container terminal Bremerhaven and an end point at a transformer in the southeast of the Osterholz county.
+For the study, a starting point was chosen at a transformer about 6~km north of the container terminal Bremerhaven and an end point at a transformer in the southeast of the Osterholz county.
 
@@
 \end{equation}
-This equation is used, to measure the degree of similarity between the paths.
-The distance are measured at the same resolution (different setting for all~touched), and same setting of all~touched (different resolution).
+This equation is used to measure the degree of similarity between the paths.
+The distances are measured at the same resolution (different setting for all~touched), and same setting of all~touched (different resolution).
 
@@
 	\subfloat[\centering resolution of 100 m.]{{\includegraphics[width=.45\linewidth]{./images/LeastCostPaths_100m_v2_small.png} }}%
-	\caption{Figures of the Least Cost Paths.
-Contrasting the paths for different resolutions.
+	\caption{Figures of the Least Cost Paths contrasting the paths for different resolutions.
 Paths with all~touched set to False are indicated by dashed lines and True are indicated by continuous lines.
@@
 
-	\caption{Figures of the Least Cost Paths.
-Contrasting the changes of the least cost paths for the different results, depending on the parameter all~touched.
+	\caption{Figures of the Least Cost Paths contrasting the changes of the Least Cost Paths for the different results, depending on the parameter all~touched.
 Paths setting all~touched to False are indicated by dashed lines and True are indicated by continuous lines.
@@
 
-When estimating the distance between the Least Costs Paths from all~touched set to True and all~touched set to False at the same resolution 
+When estimating the distance between the Least Costs Paths from all~touched set to True and all~touched set to False at the same resolution, 
 the mean minimum distance between the 100~m resolution paths is 1946.41~m and between the 5~m resolution paths is 126.04~m.
 The similarity between the all~touched set to False paths is higher, than for setting to True.
-The distance for the 100~m path to the 5~m resolution path is 243.42~m for all~touched false and 2109.44~m for all~touched true.
+The distance for the 100~m path to the 5~m resolution path is 243.42~m for all~touched False and 2109.44~m for all~touched True.
 
-When cross comparing the similarity between all paths set with all~touched set to False and the similarity in between the same resolution, the similarity in between the all~touched set to False paths is higher than, than for most paths of same resolution.
+When cross comparing the similarity between all paths set with all~touched set to False and the similarity in between the same resolution, the similarity in between the all~touched set to False paths is higher, than for most paths of the same resolution.
 Namely except for the highest resolution.
@@
 This behaviour is shown in figure~\ref{fig:paths_alltouched}.
-On a more detailed level, it can be seen, that the paths of all~touched False also converge directly to the all~touched set to True paths, but the extent is smaller.
+On a more detailed level, it can be seen, that the paths of all~touched False also converge directly to the all~touched True paths, but the extent is smaller.
 
@@
 percentage of each costs levels around the paths.
-When using all~touched True at higher resolution, the tendency is to use a higher percentage of the
+When using all~touched True at higher resolution the tendency is to use a higher percentage of the
 \textit{Preferential} Level and less of the \textit{NoRestriction} Level.
-The ratio of the 100~m buffered Least Cost Path, shifted  to Levels lower costs.
-
 There is no strong tendency for the all~touched set to False Least Cost Paths.
@@
 
-The total execution time consists of two parts. 
-The aggregation of the costs and the back tracking of the least cost to find the path.
+The total execution time consists of two parts:
+the aggregation of the costs and the back tracking of the least cost to find the path.
 
@@
 
-The first step is to optimise the computational speed, by a reduced area.
-Another method, is to improve the prediction of the medium resolution itself and thus reduce the need for a computation in higher resolution.
+The first step is to optimise the computational speed by a reduced area.
+Another method is to improve the prediction of the medium resolution itself and thus reduce the need for a computation in higher resolution.
 
-\subsubsection{Compare Least Cost Paths, from rasters of both all~touched settings}
+\subsubsection{Compare Least Cost Paths from rasters of both all~touched settings}
 
-For the example paths shown, all~touched set to True overestimates ans set to False underestimates the true costs.
+For the example paths shown, all~touched set to True overestimates ans set to False underestimates the True costs.
 
-A weighted average of the costs could therefore may be a more accurate measure and make estimated medium resolution Least Cost Path more similar, to high resolution paths.
-As above example show, the weighting should favour all~touched to False.
+A weighted average of the costs could therefore be a more accurate measure and make estimated medium resolution Least Cost Path more similar to high resolution paths.
+As above example shows the weighting should favour all~touched to False.
 
@@
 The mean distance of Least Cost Paths from rasters with different ratio is estimated to the path from the all~touched set to False raster of the higher (5~m) resolution. 
-Table~\ref{tab:4} shows, that the mean minimum distance decreases, with increasing ratio (1:1, 2:1, 4:1) and after this optimum is reached, increases with increasing ratio (8:1, 16:1 and so on).
-Comparing  the similarity of the paths of the different ratios to normal paths with 10~m resolution shows, paths with a higher ratio of all~touched set to True is nearer to the all~touched set to True paths.
-Paths with a ratio in favour of all~touched false are much closer to the all~touched set to False paths.
+Table~\ref{tab:4} shows, that the mean minimum distance decreases with increasing ratio (1:1, 2:1, 4:1) and after this optimum is reached, increases with increasing ratio (8:1, 16:1 and so on).
+Comparing  the similarity of the paths of the different ratios to normal paths with 10~m resolution shows, that paths with a higher ratio of all~touched set to True are nearer to the all~touched set to True paths.
+Paths with a ratio in favour of all~touched False are much closer to the all~touched set to False paths.
 
@@
 
-\subsubsection{Compare Least Cost Paths, from downsampled cost raster}
+\subsubsection{Compare Least Cost Paths from downsampled cost raster}
 
-As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all~touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (with bi-linear) interpolation.
+As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all~touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (bi-linear) interpolation.
 
-The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all~touched False) shows, that only downsampling to a resolution of 10~m, produces a path that is relatively close the high resolution paths(see table~\ref{tab:5}).
+The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all~touched False) shows, that only downsampling to a resolution of 10~m produces a path that is relatively close the high resolution paths(see table~\ref{tab:5}).
 
-The opposite is true for the lower resolution raster which is more similar to paths computed from the all~touched true cost raster.
-Every path from a downsampled raster is more similar to a path computed from an all~touched set to True raster, than an set to False raster, although the all~touched set to False raster of the 5 m resolution was used for downsampling.\todo{interpretation all~touched true and downsampling, both show all details, but downsampling, more similar to original geometry}
+The opposite is True for the lower resolution raster which is more similar to paths computed from the all~touched True cost raster.
+Every path from a downsampled raster is more similar to a path computed from an all~touched set to True raster, than to an all~touched set to False raster, although the all~touched set to False raster of the 5 m resolution was used for downsampling.\todo{interpretation all~touched True and downsampling, both show all details, but downsampling, more similar to original geometry}
 
@@
 
-Construct a polygon from the two Least Cost Paths (all~touched true and all~touched false of the same resolution).
+Construct a polygon from the two Least Cost Paths (all~touched True and all~touched False of the same resolution).
 Buffer the polygon with twice the maximum  minimum path distance  (see equation~\ref{eq:2}).
@@
 
-This strategy resulted in the same Least Cost path as with the original high resolution raster.
+This strategy results in the same Least Cost Path as with the original high resolution raster.
 
-This reduction of compute power provided the possibility  to run a 2.5~m.
-This clipped 2.5~m raster for all~touched set to True changed the path only slightly.
+This reduction of compute power provides the possibility  to run a 2.5~m.
+This clipped 2.5~m raster for all~touched set to True changes the path only slightly.
 
@@
 Due to the low resolution a small path became passable.
-This small path is a power line next to road between protected landscape areas.
-The road and the protected landscape area are both \textit{restricted} areas, while the power line is \textit{preferred}.
+This small path is a power line next to a road between protected landscape areas.
+The road and the protected landscape areas are both \textit{restricted} areas, while the power line is \textit{preferred}.
 
@@
 To broaden the view and verify the result, four different routes should be found using the above strategies.
-Two routes should be found from the start point to two new points in the south east of the investigated area and two routes should be found from the north and north east of the study area to the end point.
+Two routes should be found from the starting point to two new points in the south east of the investigated area and two routes should be found from the north and north east of the study area to the end point.
 
-For three of the four routes, the Least Cost Path estimation from the clipped raster, was able to calculated exactly the same result.
+For three of the four routes, the Least Cost Path estimation from the clipped raster was able to calculate exactly the same result.
 For the fourth path, the Least Cost Path from the 5~m resolution raster was clipped by the buffer around the 50~m resolution paths.
-The speed up from the clipping of the higher resolution raster depends on the number of pixels, that have been clipped.
+The speed up from the clipping of the higher resolution raster depends on the number of pixels that have been clipped.
 
-Bi-linear downsampling the of the high resolution raster to a medium resolution, did not result in any benefits compared to an original medium resolution raster.
-The aggregated cost per resolution of the Least Cost Path from the downsampled raster is higher, than the normal medium resolution 10~m raster.
-In addition, the distance from these paths to the high resolution path  is greater, than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
+Bi-linear downsampling the of the high resolution raster to a medium resolution did not result in any benefits compared to an original medium resolution raster.
+The aggregated cost per resolution of the Least Cost Path from the downsampled raster is higher than the normal medium resolution 10~m raster.
+In addition, the distance from these paths to the high resolution path  is greater than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
 \todo{interpretation: Why is the downsampling result so different? Don't know.}
@@
 \begin{table}[h!]
-	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched false raster for the four control routes, for the reference path constructed from the 5~m  and 10~m raster and the from 5~m to 10~m downsampled raster and the 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), start point to point3 (S-P3) and start point to point4 (S-P4).}
+	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched False raster for the four control routes, for the reference path constructed from the 5~m  and 10~m raster and the from 5~m to 10~m downsampled raster and the 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), start point to point3 (S-P3) and start point to point4 (S-P4).}
 	\label{tab:6}

### 2023-03-14 paper/discussion.tex
grammar, typos

@@
 These methods are used as surrogates for the more complex calculation of the Least Cost Path with the higher resolution.
-In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all touched set to False and True raster were averaged in different ratios, to compute the optimum weighted cost raster.
+In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all~touched set to False and True raster were averaged in different ratios, to compute the optimum weighted cost raster.
 While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
 This disadvantage could be reduced by normalising the costs.
+The fact that downsampling leads to paths that are closer to all~touched True, can be attributed, to the fact, that all objects are included similar to all~touched True.
 

### 2023-03-14 paper/results.tex
grammar, typos

@@
 
-The opposite is True for the lower resolution raster which is more similar to paths computed from the all~touched True cost raster.
-Every path from a downsampled raster is more similar to a path computed from an all~touched set to True raster, than to an all~touched set to False raster, although the all~touched set to False raster of the 5 m resolution was used for downsampling.\todo{interpretation all~touched True and downsampling, both show all details, but downsampling, more similar to original geometry}
+The opposite is true for the lower resolution raster which is more similar to paths computed from the all~touched True cost raster.
+Every path from a downsampled raster is more similar to a path computed from an all~touched set to True raster, than to an all~touched set to False raster, although the all~touched set to False raster of the 5 m resolution was used for downsampling.
 

### 2023-03-15 paper/discussion.tex
send paper

--- 2023-03-14 paper/discussion.tex
@@
 Similarly, the use of main memory increases. 
-This again limits the number data points, that can be processed and resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls has been used for the aggregation of higher resolution.
+This again limits the number of data points, that can be processed and the resolution, and probably causes difference in increase of computation time from power of 2 to a power of 2.2, because additional slower ram moduls has been used for the aggregation of higher resolution.
 
@@
 
-For b), two methods have been used to increase the similarity of the paths computed from the medium resolution raster, to the path of the highest resolution raster.
+For b), two methods have been used to increase the similarity of the paths computed from the medium resolution raster, to the paths of the highest resolution raster.
 These methods are used as surrogates for the more complex calculation of the Least Cost Path with the higher resolution.
-In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all touched set to False and True raster were averaged in different ratios, to compute the optimum weighted cost raster.
-While the second method of using an averaged raster, a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
+In the first method a bi-linear downsampling of the higher resolution raster has been applied and in the second method, the all~touched set to False and True rasters were averaged in different ratios, to compute the optimum weighted cost raster.
+While the second method of using an averaged raster, shows a higher similarity to the path from the highest resolution raster, the downsampling method is simpler and does not need to be optimised for the given cost.
 This disadvantage could be reduced by normalising the costs.
+The fact that downsampling leads to paths that are closer to all~touched True, can be attributed to the fact, that all objects are included, such as in to all~touched True.
 
-This shows, that the way the cost raster is created in the first place can play a crucial role, in the end result.
-So that a nuance, can cause a detour.
+This shows that the way the cost raster is created in the first place can play a crucial role in the end result.
+So, a nuance can cause a detour.
 When this behaviour occurs, the polygon may not include the Least Cost Path.
@@
 
-The set of rules, that are used to create the intermediate cost raster, includes a rule to create buffer around buildings which is set to the level \textit{prohibited} areas.
-In all touched set to False rasters, the resolution of the medium level raster needs to be high enough to show every detail.
-At least in the magnitude of the minimum object size plus twice its buffer.
+The set of rules, that are used to create the intermediate cost raster, includes a rule to create buffer around buildings which is set to the level \textit{Prohibited} areas.
+In all touched set to False rasters the resolution of the medium level raster needs to be high enough to show every detail,
+at least in the magnitude of the minimum object size plus twice its buffer.
 This is true for the 10~m resolution raster and less true for the 25~m resolution raster, that misses some details for roads for all touched False raster.
-Other details such as rivers and houses, are already included in the lower resolution raster, due to larger buffers.
+Other details such as rivers and houses are already included in the lower resolution raster, due to larger buffers.
 The raster with all~touched set to False might miss some objects, but the missing chance is propotional to the object size and the extent of overestimation is reduced.
@@
 As lines do not have a width, the route found might contain bottlenecks, that have a smaller width than the object that should be placed there.
-Therefore, the used resolution should not be smaller than, the width of the object that should be placed.
+Therefore, the used resolution should not be smaller than the width of the object that should be placed.
 This can not be avoided by downsampling, but by weighting the medium resolution all~touched True and False rasters.
 
-This paper examines the effect of computational costs and deviation of the results, is examined for a very limited set of points.
-Also, only the cost of finding the Least Cost Path from a single start point, to a single end point has been considered.
-If multiple endpoints are used, the computational cost for the aggregated cost raster has only to be paid once.
+This paper examines the effect of computational costs and deviation of the results for a limited set of points.
+Also, only the cost of finding the Least Cost Path from a single starting point to a single end point has been considered.
+If multiple endpoints are used, the computational cost for the aggregated cost raster has to be paid only once.
 
 If multiple paths are calculated from a single raster, the speed-up benefit is reduced.
-Especially, pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective, because as the number of paths increases, fewer pixels are clipped.
+Especially pre-calculation on medium resolution raster and clipping around a buffering of the resulting medium resolution paths becomes less effective, because as the number of paths increases, fewer pixels are clipped.
 The Least Cost Path algorithm does only select the single most cost-effective path.
@@
 In additional, slight variations on the costs rasters can lead to very different paths, although the costs will not change much.
-An end-user may be interested in selecting a path from a set of similar aggregated costs and applying their own evaluation criteria.
+End-users may be interested in selecting a path from a set of similar aggregated costs and applying their own evaluation criteria.
 This can be achieved a by adjusting the backtracking and return polygons, or by applying perturbation on the costs.
 
-In this work, the intermediate cost raster layer are aggregated using the maximum function. 
+In this work, the intermediate cost layers are aggregated by the maximum function. 
 Other possible aggregation functions are the sum and average functions.
-Each aggregation function can be justified, by a different interpretation of the cost and its scale.
+Each aggregation function can be justified by a different interpretation of the cost and its scale.
 
@@
 The disadvantage of aggregation with the maximum function is, that this aggregation is unable to distinguish between nuances of different overlapping intermediate costs.
-On sum or average aggregated rasters, one can distinguish between, different sublevels.
+On sum or average aggregated rasters, one can distinguish between different sublevels.
 
-The fact that all touched False rasters produce more similar results than high resolution raster, is probably due to the fact, that the default level is relatively low.
+All touched False rasters produce more similar results than all touched True raster, which is probably due to the fact, that the default level is relatively low.
 As the default level increases, the effect would probably be reduced for low resolution rasters.
-For low resolution raster, the effect would still be present, because the fact, that the raster pixel centre is used for sampling, reflects the original geometry better.
+For high resolution raster, the effect would still be present, because the use of the pixel centre for sampling, reflects the original geometry better.
 
@@
 The all~touched set to True cost raster shows every detail, but the sampling with all~touched set to True increases the size of the features.
-The fact that the aggregated costs per resolution for all touched True rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, because the costs are exponential scaled.
+The fact that the aggregated costs per resolution for all touched True rasters overestimate the costs when computing the path from a low resolution raster, might be due to the fact, that the high costs are much more frequent, as they are exponential scaled.

### 2023-03-15 paper/methods.tex
send paper

@@
 
-The Least Cost Path algorithm is used to plan a possible route for a power line.
+The Least Cost Path algorithm is used to plan a potential, cost-optimised route for a power line.
 The Least Cost Path algorithm is a Dijkstra algorithm~\cite{dijkstra_note_1959} applied on a raster map.
@@
 
-PyWPS~\cite{noauthor_welcome_2016} is used to provide the Least Cost Path algorithm as a \acrfull{wps} in combination with flask~\cite{noauthor_flask_nodate}.
-As client Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster and receives the resulting Least Cost Path.
+PyWPS~\cite{noauthor_welcome_2016} is used to provide the Least Cost Path algorithm as a \acrshort{wps} in combination with flask~\cite{noauthor_flask_nodate}.
+As client, Birdy~\cite{noauthor_birdy_nodate} connects to the \acrshort{wps}, sends the cost raster, starting point, end points and receives the resulting Least Cost Path.
 The initial implementation of the Least Cost Path algorithm is based on the implementation for the QGIS-Plugin
-'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0, but refactored to optionally export the aggregated costs.
+'Least Cost Path'~\cite{noauthor_leastcostpathdijkstra_algorithmpy_2022} in version 1.0, but refactored to optionally export the aggregated costs in a command line tool.
+On top, the \acrshort{wps} provides the complete Least Cost Path algorithm as a single capability.
 
@@
 All~touched set to False is considered the default (see figure~\ref{fig:alltouched}).
-\begin{figure}[!ht]
+\begin{figure}[!t]
 	\centering
@@
 
-\begin{table}[h!]
+\begin{table}[t!]
 	\caption{Used levels of costs, the applied numerical equivalent and example layer this cost have been used for.}

### 2023-03-15 paper/paper-CableRoute.tex
send paper

@@
 \usepackage{makecell}
+\usepackage{hyperref}
 
+\makeatletter
 \AtBeginDocument{%
+	\hypersetup{
+		pdftitle = {\@title},
+		pdfauthor = {\@author}
+	}
 	\providecommand\BibTeX{{%
 			Bib\TeX}}}
+\makeatother
 
 \setcopyright{acmcopyright}
-\copyrightyear{2022}
-\acmYear{2022}
+\copyrightyear{2023}
+\acmYear{2023}
 \acmDOI{XXXXXXX.XXXXXXX}
@@
 	\begin{abstract}
-		Least Cost Paths for a power line from a power station at the shore line to a power station further inland are calculated.
+		A lot of wind energy is produced in the north of Germany. The power has to be transported further inland. To reduce congestion in the power system, new power lines have to be installed.
+		A potential path for a power line from the northern shore is planned using the Least Cost Path algorithm.
+		The planning is implemented in Python and provided via a \acrfull{wps} with PyWPS~\cite{noauthor_welcome_2016} and flask~\cite{noauthor_flask_nodate}.
 		Downsampling and clipping are investigated as techniques to speed up the calculation.
@@
 
-	\received{14 March 2023}
+	\received{15 March 2023}
 	%%\received[revised]{12 March 2009}
@@
 
-	\section{Introduction}\label{sec:introduction}
+	\section{Introduction and Objectives}\label{sec:introduction}
 
@@
 
+	This paper aims to find potential paths for power lines by using the Least Cost Path algorithm.
+Land usage and planning data are used to estimate the costs arising by using the local area.
+The task was to provide the Least Cost Path using a \acrfull{wps}.
+	Beyond that, methods to reduce the needed compute power for finding the Least Cost Path had to be studied.
+
 	\section{Methods}\label{sec:methods}
@@
 	\section{Conclusion}\label{sec:conclusion}
-	The cost of computing of the Least Cost Path scales with the square of the resolution.
-	The difference between the aggregated costs of the paths per resolution computed from the all~touched set to False raster True raster only shows a linear decrease.
+	The computing costs of the Least Cost Path scales with the square of the resolution.
+	The difference between the aggregated costs of the paths per resolution computed from the all~touched set to False raster and the all~touched set to True raster only shows a linear decrease.
 	Therefore, the gain in accuracy per compute time decreases.
 	The presented methods attempt to circumvent this bottleneck.
-	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced, without compromising the run time for worse distances to the Least Cost Path from the high resolution raster.
+	When these strategies are applied to a medium resolution raster, the compute time for high resolution results can successfully be reduced without compromising the run time for worse distances to the Least Cost Path from the high resolution raster.
 	The paths can change significantly even for a small change in total cost.
-	Therefore, to develop an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy, to offer alternative paths and at the same time show variations between these paths.
-	An alternative would be to specify a range of costs and compute the aggregation of the costs and superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
+	Therefore, developing an alternative backtracking algorithm, that generates a corridor of costs might be a good strategy to offer alternative paths and at the same time show variations between these paths.
+	An alternative would be to specify a range of costs and compute the aggregation of the costs and than superimpose these aggregations and compute the path by backtracking from the superimposed cost raster.
 	This again increases the computation time.
 
-	In an actual search for Least Cost Path search, a survey could be used as a method to estimate the costs.
+	In an actual search for the Least Cost Path of a power, a survey could be used as a method to estimate the costs.
 	If the sample size of the survey is large enough, the

### 2023-03-15 paper/results.tex
send paper

@@
 	\begin{tabular}{ r  r  r  r  r  r  r  r  r  r}
-		res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m & $d_{max}/m$ & agg.  $ cost_{al=f}$ & agg. $ cost_{al=t}$ &  $\Delta $ costs & agg. $costs_{al=f} /m$ & agg. $costs_{al=t} /m$ \\
+		res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m & $d_{max}/m$ & agg.  $ cost_{al=f}$ & agg. $ cost_{al=t}$ &  $\Delta $ costs & agg. $costs_{al=f} \times m$ & agg. $costs_{al=t} \times m$ \\
 		\hline
@@
 
-As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all~touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (bi-linear) interpolation.
+As an alternative to the superposition of the rasters of same resolution, a high resolution (5~m) all touched False raster is downsampled to 10~m, 25~m, 50~m and 100~m (bi-linear) interpolation.
 
-The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all~touched False) shows, that only downsampling to a resolution of 10~m produces a path that is relatively close the high resolution paths(see table~\ref{tab:5}).
+The distances from the paths that are computed from the bi-linear downsampled raster to the path of the original 5~m resolution (all touched False) shows, that only downsampling to a resolution of 10~m produces a path that is relatively close the high resolution paths (see table~\ref{tab:5}).
 
@@
 In addition, the distance from these paths to the high resolution path  is greater than the distance from the original 10~m resolution path to the 5~m resolution path (see table~\ref{tab:6}).
-\todo{interpretation: Why is the downsampling result so different? Don't know.}
+%%\todo{interpretation: Why is the downsampling result so different? Don't know.}
 
-\begin{table}[h!]
-	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched False raster for the four control routes, for the reference path constructed from the 5~m  and 10~m raster and the from 5~m to 10~m downsampled raster and the 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), start point to point3 (S-P3) and start point to point4 (S-P4).}
+\begin{table}[ht!]
+	\caption{Length (l) of the path, the aggregated costs per resolution and the mean minimum distance ($d_{mean}$) to the path created from the 5~m resolution all~touched False raster for the four control routes.
+For the reference path constructed from 5~m  and 10~m raster and from 5~m to 10~m downsampled raster and 5~m clipped raster for the routes point1 to end point (P1-E), point2 to end point (P2-E), starting point to point3 (S-P3) and starting point to point4 (S-P4).}
 	\label{tab:6}

### 2023-03-21 presentations/20230323_seh_final.tex
added new path for images for the presentations.
changed image paths for the 20121014_seh_kickoff.tex and 20221202_seh_midterm.tex

added new presentation 20230323_seh_final.tex with first slides.

% !TeX spellcheck = en_GB
\documentclass[usenames,dvipsnames,aspectratio=169]{beamer}
\usetheme{Szeged}
\usecolortheme{beaver}
\usepackage{tikz}
\usepackage{subfig}
\setbeamertemplate{navigation symbols}{}
\usepackage{xcolor}

\title{Web Processing - Standardized GIS Analyses for Cable Route Planning}
\author{Sebastian Heiden}
\institute{Harz University of Applied Sciences}
\date{December 02, 2022}
\begin{document}

	\begin{frame}[plain]
		\includegraphics[scale=0.21]{images/3-HSH-Logo-RGB-en.png}
		\maketitle
	\end{frame}

	\section{Review \& Topics}
	\begin{frame}{Review \& Topics}

		\begin{minipage}[t]{0.48\textwidth}
			Review
			\begin{itemize}
				\item Rasterisation
				\item Cost Raster
				\item Cost Paths
				\item PyWPS (Start)
			\end{itemize}
		\end{minipage}
		\hfill	
		\begin{minipage}[t]{0.48\textwidth}
			Topics
			\begin{itemize}
				\item Web Processing Service (Concluded)
				\item From Cost Raster to Least Cost Path
				\item Comparing Paths
				\item Speed-Up
				\begin{itemize}
					\item Clipping
					\item Down Sampling
					\item Super Position Medium Resolution
					\item Validation
				\end{itemize}

			\end{itemize}
		\end{minipage}

	\end{frame}

	\section{Web Processing Service (Concluded)}
	\begin{frame}{Web Processing Service (Concluded)}
		\begin{itemize}
		\item goals: 
		\begin{itemize}
			\item use wps as a simple, standardized way
			\item optimize search algorithm
		\end{itemize}
		\item current: 
		\begin{itemize}
			\item testing PyWPS\footnote{https://pywps.readthedocs.io/en/latest/index.html}
			\item cost path (open Dijkstra implementation - QGIS-plugin)\footnote{https://github.com/Gooong/LeastCostPath}
		\end{itemize}
	\end{itemize}
	\end{frame}

	\section{From Cost Raster to Least Cost Path}
	\begin{frame}{From Cost Raster to Least Cost Path}
		\begin{figure}
			\centering
			\subfloat[\centering Cost Raster.]{{\includegraphics[width=.30\linewidth]{images/maps/CostRasterExample_cut.png} }}%
			\enskip
			\subfloat[\centering Aggregated Cost Raster.]{{\includegraphics[width=.30\linewidth]{images/maps/AggregatedCosts_cut.png} \label{fig:aggregation}}}%
			\enskip
			\subfloat[\centering Least Cost Path.]{{\includegraphics[width=.30\linewidth]{images/maps/LeastCostPathExample_cut.png} }}%

			\caption{Figures of the cost raster and the resulting aggregated costs and the Least Cost Path for a resolution of 50~m, all~touched set to False.}
			\label{fig:costs2path}
		\end{figure}
	\end{frame}

	\section{Comparing Paths}
	\begin{frame}{Comparing: Map}
		\begin{figure}
			\centering

			\subfloat[\centering all touched False.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPaths_al_F_v2_small.png} }}%
			\qquad
			\subfloat[\centering all touched True.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPaths_al_T_v2_small.png} }}%

			\caption{Figures of the Least Cost Paths, depending on the parameter all~touched. All~touched False: dashed lines, True: continuous lines. Higher resolutions: green, lower resolutions: red.}
			\label{fig:paths_alltouched}
		\end{figure}
	\end{frame}

	\begin{frame}{Comparing: Costs}
		\begin{table}[t]
			\caption{Least Cost Paths: length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$), agg. costs and the agg. costs per resolution.} 
			\label{tab:2}
			\centering
			\begin{tabular}{ r  r  r  r  r  r   }
				res /m & $l_{al=f} /m$ & $l_{al=t} /m$ & $d_{mean}$ /m  & agg. $costs_{al=f} \times m$ & agg. $costs_{al=t} \times m$ \\
				\hline
				5 	& 76136.3	& 78002.0 &  126.0  & 93329.6 &  97584.8 \\
				10 	& 75430.1 	& 77936.6 &  277.9  & 89312.5 &  97311.8 \\
				25 	& 75422.9 	& 78422.9 &  313.8  & 83871.7 &  96816.4 \\
				50 	& 76135.0	& 70620.0 & 1140.0  & 70451.2 & 115003.7 \\
				100 & 76283.8	& 74120.7 & 1946.4  & 64051.6 & 167226.8 \\
			\end{tabular}
		\end{table}
	\end{frame}

\section{Speed-Up}
	\begin{frame}{Speed-Up: Clipping}
		\begin{figure}
			\centering

			\subfloat[\centering 50~m Resolution.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_50m_small.png} }}%
			\qquad
			\subfloat[\centering 5~m, Clipped.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_5m_false_clipped_small.png} }}%

			\caption{Figures of the Least Cost Paths, on the cost raster with 50~m resolution and the computed paths, used to clip the high resolution raster.}
			\label{fig:clipping}
		\end{figure}
	\end{frame}

	\begin{frame}{Speed-Up: Down Sampling}
	\end{frame}

	\begin{frame}{Speed-Up:Super Position Medium Resolution}
	\end{frame}

	\begin{frame}{Speed-Up:Validation}
	\end{frame}
\end{document}

### 2023-03-22 presentations/20230324_seh_final.tex
renamed final presentation: corrected the date
added new slides and images + gif

--- 2023-03-21 presentations/20230323_seh_final.tex
@@
 \usecolortheme{beaver}
+\usepackage{multimedia}
 \usepackage{tikz}
@@
 \institute{Harz University of Applied Sciences}
-\date{December 02, 2022}
+\date{March 24, 2023}
 \begin{document}
@@
 	\begin{frame}{Web Processing Service (Concluded)}
-		\begin{itemize}
-		\item goals: 
-		\begin{itemize}
-			\item use wps as a simple, standardized way
-			\item optimize search algorithm
-		\end{itemize}
-		\item current: 
-		\begin{itemize}
-			\item testing PyWPS\footnote{https://pywps.readthedocs.io/en/latest/index.html}
-			\item cost path (open Dijkstra implementation - QGIS-plugin)\footnote{https://github.com/Gooong/LeastCostPath}
-		\end{itemize}
-	\end{itemize}
+	\centering
+	\movie[externalviewer]{\includegraphics[width=0.75\textwidth]{images/wps.png}}{images/wps.gif}
+
 	\end{frame}
@@
 	\section{From Cost Raster to Least Cost Path}
+	\begin{frame}{Reminder: Sampling Resolution}
+		\begin{figure}
+		\centering
+		\subfloat[\centering 10~m]{{\includegraphics[width=.30\linewidth]{images/maps/CostRaster_10m_false.png} }}%
+		\enskip
+		\subfloat[\centering 25~m]{{\includegraphics[width=.30\linewidth]{images/maps/CostRaster_25m_false.png} }}% 
+		\enskip
+		\subfloat[\centering 50~m]{{\includegraphics[width=.30\linewidth]{images/maps/CostRaster_50m_false.png} }}%
+
+		\caption{Maps of the cost raster for all~touched set to False for different resolutions.}
+		\label{fig:CostAllTouchedFalse}
+		\end{figure}
+	\end{frame}
+
+	\begin{frame}{Reminder: Sampling Resolution}
+		\begin{figure}
+		\begin{figure}
+			\centering
+			\subfloat[\centering 10~m]{{\includegraphics[width=.30\linewidth]{images/maps/CostRaster_10m_true.png} }}%
+			\enskip
+			\subfloat[\centering 25~m]{{\includegraphics[width=.30\linewidth]{images/maps/CostRaster_25m_true.png} }}% 
+			\enskip
+			\subfloat[\centering 50~m]{{\includegraphics[width=.30\linewidth]{images/maps/CostRaster_50m_true.png} }}%
+
+			\caption{Maps of the cost raster for all~touched set to True for different resolutions.}
+			\label{fig:CostAllTouchedTrue}
+		\end{figure}
+		\end{figure}
+	\end{frame}
+
 	\begin{frame}{From Cost Raster to Least Cost Path}
@@
 
-			\subfloat[\centering 50~m Resolution.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_50m_small.png} }}%
+			\subfloat[\centering 50~m.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_50m_small.png} }}%
 			\qquad
@@
 
-	\begin{frame}{Speed-Up: Down Sampling}
-	\end{frame}
+	\begin{frame}{Speed-Up: Medium Resolution}
+		\begin{minipage}[t]{0.48\textwidth}
+		bi-linear down sampling
+		\begin{itemize}
+			\item 59.3~m average distance
+			\item closer to path from all touched True
 
-	\begin{frame}{Speed-Up:Super Position Medium Resolution}
+			\item independent from cost ratios
+		\end{itemize}
+		\end{minipage}
+		\hfill	
+		\begin{minipage}[t]{0.48\textwidth}
+			weighted average cost raster
+			\begin{itemize}
+				\item best ration 4:1 False:True
+				\item 40.1~m average distance
+				\item closer to path from all touched True
+				\item Optimum: depends on cost ratio
+			\end{itemize}
+		\end{minipage}
 	\end{frame}
@@
 	\begin{frame}{Speed-Up:Validation}
+		\begin{minipage}[t]{0.48\textwidth}
+			clipping
+			\begin{itemize}
+				\item best path: 3 times
+				\item Speed-Up: Unpredictable
+			\end{itemize}
+		\end{minipage}
+		\hfill	
+		\begin{minipage}[t]{0.48\textwidth}
+			bi-linear down sampling
+			\begin{itemize}
+				\item distance worse, than from original medium resolution
+				\item Speed-Up: square of resolution
+			\end{itemize}
+		\end{minipage}
+	\end{frame}
+
+	\begin{frame}{Conclussion AND Outlook}
+
+	\begin{minipage}[t]{0.48\textwidth}
+		Conclusion
+		\begin{itemize}
+			\item provide Least Cost Path with WPS
+			\item studied algoritmic Speed-Up
+			     \begin{itemize}
+			     	\item clipping
+			     	\item medium resolution
+			     \end{itemize}
+		\end{itemize}
+	\end{minipage}
+	\hfill	
+	\begin{minipage}[t]{0.48\textwidth}
+		Outlook
+		\begin{itemize}
+			\item non algorithmic Speed-Up
+			\item set of possible paths
+			\item stability: pertubation
+		\end{itemize}
+	\end{minipage}
 	\end{frame}

### 2023-03-24 presentations/20230325_seh_final.tex
updated presentation and date

--- 2023-03-22 presentations/20230324_seh_final.tex
@@
 \institute{Harz University of Applied Sciences}
-\date{March 24, 2023}
+\date{March 25, 2023}
 \begin{document}
@@
 
-	\section{Review \& Topics}
-	\begin{frame}{Review \& Topics}
+	\section{Recap \& Follow Up}
+
+	\begin{frame}
+		\begin{figure}
+			\centering
+			\includegraphics[height=0.70\textheight]{images/maps/powerlines_germany_turbine.png}
+			\caption{Map of the medium and high voltages power lines in Western Germany, OpenStreetMap\footnote{G.
+Boeing, „OSMnx: New methods for acquiring, constructing, analyzing, and visualizing complex street networks“, Computers, Environment and Urban Systems, Bd.
+65, S.
+126–139, Sep.
+2017.}.}
+			\label{fig:PowerLines}
+		\end{figure}
+	\end{frame}
+
+	\begin{frame}{Recap \& Follow Up}
 
 		\begin{minipage}[t]{0.48\textwidth}
-			Review
+			Recap
 			\begin{itemize}
-				\item Rasterisation
-				\item Cost Raster
-				\item Cost Paths
-				\item PyWPS (Start)
+				\item rasterisation
+				\item cost  raster
+				\item Least Cost Paths
+				\item Web Processing Service (Start)
 			\end{itemize}
@@
 		\begin{minipage}[t]{0.48\textwidth}
-			Topics
+			Follow Up
 			\begin{itemize}
 				\item Web Processing Service (Concluded)
-				\item From Cost Raster to Least Cost Path
-				\item Comparing Paths
-				\item Speed-Up
+				\item from cost raster to Least Cost Path
+				\item comparing paths
+				\item speed-up
 				\begin{itemize}
-					\item Clipping
-					\item Down Sampling
-					\item Super Position Medium Resolution
-					\item Validation
+					\item clipping
+					\item down sampling and superposition
+					\item validation
 				\end{itemize}
@@
 	\section{From Cost Raster to Least Cost Path}
-	\begin{frame}{Reminder: Sampling Resolution}
+	\begin{frame}{Recap: Sampling Resolution}
 		\begin{figure}
@@
 
-	\begin{frame}{Reminder: Sampling Resolution}
+	\begin{frame}{Recap: Sampling Resolution}
 		\begin{figure}
@@
 
-			\caption{Figures of the cost raster and the resulting aggregated costs and the Least Cost Path for a resolution of 50~m, all~touched set to False.}
+			\caption{Figures of the different steps for the Least Cost Path generation for a resolution of 50~m, all~touched set to False.}
 			\label{fig:costs2path}
@@
 
-			\caption{Figures of the Least Cost Paths, depending on the parameter all~touched. All~touched False: dashed lines, True: continuous lines. Higher resolutions: green, lower resolutions: red.}
-			\label{fig:paths_alltouched}
+			\caption{Figures of the Least Cost Paths, depending on the parameter all~touched. All~touched False: dashed lines, True: continuous lines. Gradation of resolutions: high (green) to low (red).}
+			\label{fig:pathsAllTouched}
 		\end{figure}
@@
 		\begin{table}[t]
-			\caption{Least Cost Paths: length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$), agg. costs and the agg. costs per resolution.} 
+			\caption{Least Cost Paths: length~(l) for the different resolutions~(res), including the mean minimum distance~($d_{mean}$) and the agg. costs per resolution.} 
 			\label{tab:2}
@@
 			\qquad
-			\subfloat[\centering 5~m, Clipped.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_5m_false_clipped_small.png} }}%
+			\subfloat[\centering 5~m, clipped.]{{\includegraphics[width=.45\linewidth]{images/maps/LeastCostPath_5m_false_clipped_small.png} }}%
 
-			\caption{Figures of the Least Cost Paths, on the cost raster with 50~m resolution and the computed paths, used to clip the high resolution raster.}
+			\caption{Figures of the Least Cost Paths on the cost raster with 50~m resolution, used to clip the high resolution raster.}
 			\label{fig:clipping}
@@
 				\item closer to path from all touched True
-				\item Optimum: depends on cost ratio
+				\item optimum: depends on cost ratio
 			\end{itemize}
@@
 
-	\begin{frame}{Speed-Up:Validation}
+	\begin{frame}{Speed-Up: Validation}
 		\begin{minipage}[t]{0.48\textwidth}
@@
 			\begin{itemize}
-				\item best path: 3 times
-				\item Speed-Up: Unpredictable
+				\item reproduce-able best paths:\\ 3 times out of 4
+				\item speed-up: unpredictable
 			\end{itemize}
@@
 			\begin{itemize}
-				\item distance worse, than from original medium resolution
-				\item Speed-Up: square of resolution
+				\item distance worse than from original medium resolution
+				\item speed-up: square of resolution
 			\end{itemize}
@@
 
-	\begin{frame}{Conclussion AND Outlook}
+	\begin{frame}{Conclusion and Outlook}
 
 	\begin{minipage}[t]{0.48\textwidth}
-		Conclusion
+		conclusion
 		\begin{itemize}
 			\item provide Least Cost Path with WPS
-			\item studied algoritmic Speed-Up
+			\item studied algorithmic speed-up
 			     \begin{itemize}
@@
 	\begin{minipage}[t]{0.48\textwidth}
-		Outlook
+		outlook
 		\begin{itemize}
-			\item non algorithmic Speed-Up
+			\item non algorithmic speed-up
 			\item set of possible paths
-			\item stability: pertubation
+			\item stability: perturbation
 		\end{itemize}
